### 2019-10-04 arxiv/local_spectral_density_clustering_arxiv.tex

\documentclass[11pt,twoside]{article}
\usepackage{fancyhdr}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue,bookmarks=false]{hyperref}
\usepackage{amsfonts,epsfig,graphicx}
\usepackage{afterpage}
\usepackage{amsmath,amssymb,amsthm} 
\usepackage{fullpage}
\usepackage{epsf} 
\usepackage{graphics} 
\usepackage{amsfonts,amsmath}
\usepackage[sort,numbers]{natbib} 
\usepackage{psfrag,xspace}
\usepackage{color,etoolbox}

\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-6cm}
\setlength{\textheight}{\paperheight}
\addtolength{\textheight}{-4cm}
\addtolength{\textheight}{-1.1\headheight}
\addtolength{\textheight}{-\headsep}
\addtolength{\textheight}{-\footskip}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\renewcommand{\floatpagefraction}{.8}%

\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem*{remark}{Remark}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{float}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{xcolor}
%\usepackage{xr-hyper}
%\usepackage{hyperref}
%\usepackage[reqno]{amsmath}
%\usepackage{amsfonts, amsthm, amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{bm}
\usepackage{mathtools}

%%%%%% Begin Alden

\newcommand{\diam}{\rho}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\left\{#1\right\}_{n \in \mathbb{N}}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\cut}{\mathrm{cut}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\Phibf}{\Phi_{u}}
\newcommand{\Psibf}{\Psi_{u}}
\newcommand{\taubf}{\tau_{u}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\TV}{\mathrm{TV}}

%%% Vectors
\newcommand{\pbf}{p}        % removed bold font
\newcommand{\qbf}{\mathbf{q}}
\newcommand{\ebf}[1]{{e}_{#1}}
\newcommand{\pibf}{\pi}

%%% Matrices (no bold font)
\newcommand{\Abf}{A}
\newcommand{\Xbf}{X}             % removed bold font 
\newcommand{\Wbf}{W}
\newcommand{\Lbf}{L}
\newcommand{\Dbf}{D}
\newcommand{\Ibf}[1]{I_{#1}}

%%% Probability distributions (and related items)
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Cbb}{\mathbb{C}}
\newcommand{\Ebb}{\mathbb{E}}

%%% Sets
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Cset}{\mathcal{C}}
\newcommand{\Aset}{\mathcal{A}}
\newcommand{\Asig}{\Aset_{\sigma}}
\newcommand{\Csig}{\Cset_{\sigma}}
\newcommand{\Asigr}{\Aset_{\sigma,\sigma + r}}
\newcommand{\Csigr}{\Cset_{\sigma,\sigma + r}}

%%% Graph quantities
\newcommand{\Cest}{\widehat{C}}
\newcommand{\degminpr}{\deg_{\min}'}
\newcommand{\degminwt}{\widetilde{\deg}_{\min}}
\newcommand{\degmaxwt}{\widetilde{\deg}_{\max}}
\newcommand{\degmax}{\deg_{\max}}
\newcommand{\piminwt}{\widetilde{\pi}_{\min}}
\newcommand{\piminpr}{\pibf_{\min}'}
\newcommand{\degmin}{\deg_{\min}}

%%% Operators
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dt}{\,dt}

%%% Algorithm notation
\newcommand{\ppr}{{\sc PPR}}
\newcommand{\pprspace}{{\sc PPR~}}

%%% Tilde notation for quantities over the expansion set 
\newcommand{\wn}{\widetilde{n}}
\newcommand{\wX}{\widetilde{\Xbf}}
\newcommand{\wx}{\widetilde{x}}
\newcommand{\wz}{\widetilde{z}}
\newcommand{\wbz}{\widetilde{\bf{z}}}
\newcommand{\wu}{\widetilde{u}}
\newcommand{\wPbb}{\widetilde{\Pbb}}
\newcommand{\wf}{\widetilde{f}}
\newcommand{\wDbf}{\widetilde{\Dbf}}
\newcommand{\piwt}{\widetilde{\pi}}

\newcommand{\sbcomment}[1]{{\color{red} \bf{{{{SB --- #1}}}}}}

%\newtheoremstyle{aldenthm}
%{6pt} % Space above
%{6pt} % Space below
%{\itshape} % Body font
%{} % Indent amount
%{\bfseries} % Theorem head font
%{.} % Punctuation after theorem head
%{.5em} % Space after theorem head
%{} % Theorem head spec (can be left empty, meaning `normal')

%\theoremstyle{aldenthm}
%\newtheorem{theorem}{Theorem}
%\newtheorem{definition}{Definition}
%\newtheorem{lemma}{Lemma}
%\newtheorem{proposition}{Proposition}
%\newtheorem{corollary}{Corollary}

%\newtheoremstyle{aldenrmrk}
%{6pt} % Space above
%{6pt} % Space below
%{} % Body font
%{} % Indent amount
%{\itshape} % Theorem head font
%{.} % Punctuation after theorem head
%{.5em} % Space after theorem head
%{} % Theorem head spec (can be left empty, meaning `normal')
%
%\theoremstyle{aldenrmrk}
%\newtheorem{remark}{Remark}
%%%%%% End Alden

%\title{Local Spectral Clustering of Density Upper Level Sets}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.
%
%\author{
%Alden Green \And
%Sivaraman Balakrishnan \And
%Ryan J. Tibshirani}
%
%\begin{document}
%\maketitle
%
%\vspace{0.1in}
%\begin{abstract}

\newcommand{\widgraph}[2]{\includegraphics[keepaspectratio,width=#1]{#2}}
\newcommand{\Like}{\ensuremath{\mathcal{L}}}
\newcommand{\Ball}[2]{\mathbb{B}_{#1}(#2)}
\newcommand{\Complement}[1]{\overline{#1}}

\newcommand{\xsam}{\ensuremath{x}}
\newcommand{\samind}{\ensuremath{\ell}}
\newcommand{\Xrv}{\ensuremath{X}}
\newcommand{\Event}{\ensuremath{\mathcal{E}}}
\newcommand{\Fevent}{\ensuremath{\mathcal{F}}}

\newcommand{\usedim}{\ensuremath{d}}
\newcommand{\mubold}{\ensuremath{\boldsymbol{\mu}}}
\newcommand{\lambold}{\ensuremath{\boldsymbol{\lambda}}}
\newcommand{\sep}{\ensuremath{\xi}}
\newcommand{\mixind}{\ensuremath{i}}
\newcommand{\mixtwo}{\ensuremath{j}}
\newcommand{\nummix}{\ensuremath{M}}
\newcommand{\mustar}{\ensuremath{\mu^*}}
\newcommand{\muboldstar}{\ensuremath{\mubold^*}}
\newcommand{\muboldt}{\ensuremath{\mubold^t}}

\newcommand{\numobs}{\ensuremath{n}}
\newcommand{\SamLike}{\ensuremath{\Like_\numobs}}
\newcommand{\PopLike}{\ensuremath{\Like}}
\newcommand{\Exs}{\E}
\newcommand{\thetanew}{\ensuremath{\theta^{\text{\small{new}} }}}

\newcommand{\stepsize}{\ensuremath{s}}

\newcommand{\QMAT}{\ensuremath{\mathbf{Q}}} 
\newcommand{\DMAT}{\ensuremath{\mathbf{D}}}

\newcommand{\defn}{\ensuremath{: \, =}}
\newcommand{\muboldtilde}{\ensuremath{\tilde{\mubold}}}

%%% New version of \caption puts things in smaller type, single-spaced 
%%% and indents them to set them off more from the text.
\makeatletter
\long\def\@makecaption#1#2{
        \vskip 0.8ex
        \setbox\@tempboxa\hbox{\small {\bf #1:} #2}
        \parindent 1.5em  %% How can we use the global value of this???
        \dimen0=\hsize
        \advance\dimen0 by -3em
        \ifdim \wd\@tempboxa >\dimen0
                \hbox to \hsize{
                        \parindent 0em
                        \hfil 
                        \parbox{\dimen0}{\def\baselinestretch{0.96}\small
                                {\bf #1.} #2
                                %%\unhbox\@tempboxa
                                } 
                        \hfil}
        \else \hbox to \hsize{\hfil \box\@tempboxa \hfil}
        \fi
        }
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{center} {\Large{\bf{Local Spectral Clustering of Density Upper Level Sets}}}

\vspace*{.3cm}

{\large{
\begin{center}
Alden Green ~~~~~ Sivaraman Balakrishnan~~~~~ Ryan Tibshirani\\
\vspace{.2cm}
\end{center}

\begin{tabular}{c}
Department of Statistics and Data Science \\
Carnegie Mellon University
\end{tabular}

\vspace*{.2in}

\begin{tabular}{c}
\texttt{\{ajgreen,sbalakri,ryantibs\}@andrew.cmu.edu}
\end{tabular}
}}

\vspace*{.2in}

\today
\vspace*{.2in}

\begin{abstract}
\noindent We analyze the Personalized PageRank (PPR) algorithm, a local spectral method for clustering,
which extracts clusters using locally-biased random walks around a user-specified
seed node.  In contrast to previous work, we adopt a traditional statistical
learning setup, where we obtain samples from an unknown distribution, and aim to
identify connected regions of high-density (density clusters).  We prove that
PPR, run on a neighborhood graph, extracts sufficiently salient density
clusters. We also provide empirical support for our theory.
\end{abstract}
\end{center}
\section{Introduction}
\label{sec: introduction}
In this paper, we consider the problem of clustering: splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference.  Our particular focus is on
spectral clustering methods, a family of powerful
nonparametric clustering algorithms. Roughly speaking, a spectral algorithm
first constructs a geometric graph $G$, where vertices correspond to samples,
and edges correspond to proximities between samples. The algorithm 
then estimates a feature
embedding based on (an appropriate) Laplacian matrix of the graph 
$G$, and applies a simple clustering
technique (like $k$-means clustering) in the embedded feature space.

% Spectral clustering and density clustering
%Spectral clustering algorithms, which
%originated in the early work of Fiedler \cite{fiedler73}, have seen a recent revival 
%(see for instance the recent review~\cite{XX}) focused mainly on understanding the performance of spectral clustering on the Stochastic Block Model (SBM) and closely related problems.
%A parallel line of work \cite{shi09,schiebinger15}, has considered spectral partitioning applied to a non-parametric mixture model
%where once again the clustering goal is to recover the latent labels of a sample $X_1,\ldots,X_n \sim P$. These works proceed by relating the eigenvectors of the graph Laplacian to eigenfunctions of an appropriately defined integral operator, and then using assumptions about the (non-)overlap of the mixture components to show that these eigenfunctions elucidate the latent labels.

When applied to geometric graphs built from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to increased interest in
local spectral clustering algorithms, which leverage locally-biased spectra
computed using random walks around some user-specified seed node.  A popular 
local clustering algorithm is the Personalized PageRank (PPR) algorithm, first introduced by 
\citet{haveliwala2003}, then further developed by
several others~\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013}.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, leading 
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (such as
conductance).  In this paper, we adopt a classical statistical viewpoint,
and examine what the output of local clustering on a data set reveals about the
underlying density of the samples 
$f$.  In particular, we examine the ability of PPR to recover
\emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
(a central object of interest in the statistical clustering literature, dating
back to the work of \citet{hartigan1981}).

\section{Background and Related Work}
We begin by providing some standard background on the PPR algorithm
and the density clustering setup, before turning our attention to related work
and a
detailed summary of our contributions.

\subsection{PPR on a neighborhood graph} 
We now describe the clustering
algorithm that will be our focus for the rest of the paper. Let $\Xbf = \{x_1,
\ldots, x_n\}$ be a sample drawn i.i.d.\ from a distribution $\Pbb$ on $\Rd$,
with density~$f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
\emph{$r$-neighborhood graph} of $\Xbf$, an unweighted, undirected graph with
vertices $V=\Xbf$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i -
x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $\Abf \in
\Reals^{n \times n}$ the adjacency matrix, with entries $\Abf_{uv} = 1$ if
$(u,v) \in E$ and $0$ otherwise.  We also denote by $\Dbf$ the diagonal degree
matrix, with $\Dbf_{uu} = \sum_{v \in V} \Abf_{uv}$, and by $\Ibf{}$ the $n
\times n$ identity matrix.

Next, we define the PPR vector $\pbf_v = \pbf(v,\alpha;G_{n,r})$, based on
a seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, to be 
the solution of the following linear system:
\begin{equation}
\label{eqn: ppr_vector}
\pbf_v = \alpha \ebf{v} + (1 - \alpha) \pbf \Wbf,
\end{equation}
where $\Wbf = (\Ibf{} + \Dbf^{-1}\Abf)/2$ is the lazy random walk matrix over
$G_{n,r}$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in the
$v$-th position and 0 elsewhere).

\subsection{Sweep cut.}
\label{subsection: sweep_cut}

Finally, we employ the sweep cut method to obtain a local cluster from the PPR embedding $\pbf_v$. For a level $\beta > 0$ and a target stationary probability $\pi_0 > 0$, we define a \emph{$\beta$-sweep cut} of $\pbf_v = (\pbf_v(u))_{u \in V}$
as   
\begin{equation}
\label{eqn: sweep_cuts}
S_\beta := \set{u \in V: \pbf_v(u) > \beta \pi_0}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(\frac{1}{40},\frac{1}{12})$}\footnote{The choice of a specific range such as  
\smash{$(\frac{1}{40}, \frac{1}{12})$} is standard in the analysis of PPR
algorithms, see, e.g., \citep{zhu2013}.}, we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} that has minimum normalized cut.
%\smash{$\Phi(S_{\beta^*}; G_{n,r})$}. 
For concreteness, the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

\begin{algorithm}
\caption{PPR on a neighborhood graph}
\label{alg: ppr}	
{\bfseries Input:} data $\Xbf=\{x_1,\ldots,x_n\}$, radius $r > 0$, teleportation
parameter $\alpha \in [0,1]$, seed $v \in \Xbf$, target stationary probability
$\pi_0 > 0$. \\     
{\bfseries Output:} cluster $\Cest \subseteq V$.
\begin{algorithmic}[1]
  \STATE Form the neighborhood graph $G_{n,r}$.
  \STATE Compute the PPR vector $\pbf_v=\pbf(v, \alpha; G_{n,r})$ as in \eqref{eqn: 
    ppr_vector}. 
  \STATE For \smash{$\beta \in (\frac{1}{40}, \frac{1}{12})$} compute sweep cuts 
  $S_{\beta}$ as in \eqref{eqn: sweep_cuts}.
  \STATE Return the cluster \smash{$\Cest = S_{\beta^*}$}, where  
  $$
  \beta^* = \argmin_{\beta \in (\frac{1}{40}, \frac{1}{11})}~ \Phi(S_{\beta}; G_{n,r}).
  $$
\end{algorithmic}
\end{algorithm}

\subsection{Estimation of density clusters} 
Let \smash{$\Cbb_f(\lambda)$} denote 
the connected components of the density upper level set $\{x \in \Rd: f(x) >
\lambda\}$.  For a given density cluster \smash{$\Cset \in \Cbb_f(\lambda)$}, we
call $\Cset[\Xbf] = \Cset \cap \Xbf$ the \emph{empirical density cluster}. The
size of the symmetric set difference between estimated and empirical clusters is 
a commonly used metric to quantify cluster estimation error
\citep{korostelev1993,polonik1995,rigollet2009}.

\begin{definition}[Symmetric set difference]
  \label{def: symmetric_set_diff}
  For an estimator \smash{$\Cest \subseteq \Xbf$} and set
  $\mathcal{S} \subseteq \Reals^d$, we define   
  \begin{equation}
    \label{eqn: misclassification_rate}
    \Delta(\Cest, \mathcal{S}) := \abs{\Cest \setminus \mathcal{S}[\Xbf] \cup
      \mathcal{S}[\Xbf] \setminus \Cest},
  \end{equation}
  the cardinality of the symmetric set difference between 
  \smash{$\Cest$} and $\mathcal{S} \cap \Xbf = \mathcal{S}[\Xbf]$. 
\end{definition}

However, the symmetric set difference does not measure whether \smash{$\Cest$} 
can distinguish any two distinct clusters \smash{$\Cset,\Cset' \in
  \Cbb_f(\lambda)$}. We therefore also study a second notion of cluster
estimation, first introduced by \citet{hartigan1981}, and defined
asymptotically.

\begin{definition}[Consistent density cluster estimation]
  \label{def: consistent_density_cluster_estimation}
  For an estimator \smash{$\Cest \subseteq \Xbf$} and cluster 
  \smash{$\Cset \in \Cbb_f(\lambda)$}, we say \smash{$\Cest$} is a consistent 
  estimator of $\Cset$ if for all \smash{$\Cset' \in \Cbb_f(\lambda)$} with
  $\Cset \not= \Cset'$, the following holds as $n \to \infty$: 
  \begin{equation}
    \label{eqn: consistent_density_cluster_recovery}
    \Cset[\Xbf] \subseteq \Cest \quad \text{and} \quad
    \Cest \cap \Cset'[\Xbf] = \emptyset,
  \end{equation}
  with probability tending to 1.
\end{definition}
\noindent Consistent cluster recovery roughly ensures that, for a given $\lambda$, the estimated cluster $\Cest$ contains all
points in a true
density cluster $\Cset$, and simultaneously does not contain any points in any other density cluster $\Cset'$.

With these definitions in place, our broad goal will be to understand the extent to which the PPR algorithm is able to
recover a cluster which either guarantees low symmetric set difference to a true density cluster, or which consistently estimates
a true density cluster.

%\subsection{Our Contributions}

\subsection{Related Work and Our Contributions}
In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. The ability of spectral clustering to perform community detection, in this context is well-understood. In particular, \citet{rohe2011} upper bound the fraction of nodes misclassified by a spectral algorithm for the high-dimensional (large number of blocks) SBM,  \citet{lei2015} extend these results to the sparse (low average degree) regime, and \citet{balakrishnan2011} analyze the misclassification rate when the block model exhibits some hierarchical structure. The framework we consider, where nodes correspond to data points sampled from an underlying density, and edges between nodes are formed based on geometric proximity, is quite different than the SBM, and therefore so is our analysis.

The study of spectral algorithms on neighborhood graphs has focused on establishing consistency results-- that is, asymptotic convergence of eigenvalues and eigenvectors of matrices to eigenvalues and eigenfunctions of limiting operators. \citet{koltchinskii2000} establish convergence of spectral projections of the adjacency matrix to a limiting integral operator, with similar results obtained using simplified proofs in \citet{rosasco10}. \citet{vonluxburg2008} studies convergence of eigenvectors of the Laplacian matrix for a neighborhood of fixed radius. \citet{belkin07} and \citet{garciatrillos18} extend these results to the regime $r \to 0$ as $n \to \infty$.

These results are of fundamental importance; however, the behavior of the spectra of these underlying integral operators is somewhat hard to grasp. Therefore, further work relating this spectra to the geometry of the distribution $\Pbb$ is of interest. In this spirit, \citet{shi2009} and \citet{schiebinger2015} examine the ability of  
spectral algorithms to recover the latent labels in certain geometrically well-conditioned nonparametric mixture models. Their results focus on global rather than local 
methods, and thus impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density 
clusters, which is the focus in our work. Perhaps most importantly, these works
rely on general cluster saliency conditions, which implicitly depend on many
distinct geometric aspects of the cluster $\Cset$ under consideration. We make
this dependence more explicit, and in doing so expose the role each geometric
condition plays in the clustering problem.

A substantial portion of our technical effort involves estimating graph-theoretic connectivity properties of empirical density clusters in the neighborhood graph $G_{n,r}$. In particular, we prove an upper bound on the mixing time of a random walk run only over the subset of nodes in $G_{n,r}$ which fall within the empirical density cluster $\Cset[\Xbf]$. To do so, we rely on a series of seminal works upper bounding the mixing time of \emph{geometric random walks} (see \citet{vempala2005} for a comprehensive review.)  \citet{dyer1991} initiated this study, using these random walks as a fundamental subroutine to efficiently compute volumes of high-dimensional convex bodies. These results are improved in \citep{lovasz1990,kannan97,kannan06}, who show, among other things, that the bounds on mixing time can be sharpened by avoiding so-called ``start-penalties''.(These improvements are crucial to our work, for they allow us to avoid factors of $\log n$ which would imply divergence as $n \to \infty$.) Following the lead of \citep{abbasi-yadkori2016,abbasi-yadkori2016a}, we extend these results to hold for Lipschitz deformations of convex sets. Additionally, we relate the mixing time of these (continuous-space) geometric random walks to the mixing time of random walks over (discrete) neighborhood graphs.

Finally, it is worth mentioning that density clustering and level set estimation is of course a well-studied problem. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
We emphasize that our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

A summary of our results (and an outline for this
paper) is as follows.

\begin{enumerate}
\item In Section \ref{sec: consistent_cluster_estimation_with_ppr}, we introduce
  a set of natural geometric conditions on the density cluster $\Cset$
  %formalize a measure of difficulty based on these geometric conditions, 
  and show that when Algorithm \ref{alg: ppr} is properly initialized, the size of
  the symmetric set difference of \smash{$\Cest$} and a thickened version of the
  density cluster $\Csig$ can be bounded in a meaningful way.
  % this difficulty measure.

\item We further show in Section \ref{sec:
    consistent_cluster_estimation_with_ppr} that if the density cluster 
  $\Cset$ is particularly well-conditioned, Algorithm \ref{alg: ppr}
  will consistently estimate a density cluster in the sense of
  \eqref{eqn: consistent_density_cluster_recovery}.

\item In Section \ref{sec: analysis}, we detail some of the analysis required to
  prove our main results, and expose the parts various geometric quantities play 
  in the difficulty of the clustering problem.

\item In Section \ref{sec: examples}, we show the implications of our theory for some simple generative models, and compare our upper bounds on the symmetric set difference metric to those achievable by classical density cluster estimators.

\item In Section \ref{sec: experiments}, we empirically investigate the
  tightness of our analysis, and provide examples showing how violations of our
  geometric conditions impact density cluster recovery by PPR.
\end{enumerate}

\noindent At this point we summarize one of our main takeaways: 
PPR, run on a neighborhood graph, recovers \emph{geometrically
compact} high-density clusters. Our main theoretical results, make this takeaway 
precise, and provide concrete ways of quantifying the geometric
compactness of a density cluster. To complement this, we also provide experimental
evidence that the PPR algorithm can fail to recover high-density clusters which are not
geometrically compact.

\section{Main Results}
\label{sec: consistent_cluster_estimation_with_ppr}
In this section we present our main results on the accuracy of the PPR algorithm 
for recovering density clusters. 
We begin by formally introducing a variety of geometric conditions, 
and use these to define a condition
number $\kappa(\Cset)$, which measures the difficulty PPR will have in  
estimating a density cluster $\Cset$.
With this condition in place 
our first main result (Theorem~\ref{thm: misclassification_rate}) provides a bound
on the symmetric set difference between the estimated cluster $\Cest$,
obtained by an appropriately initialized version of the PPR algorithm,
and
the target cluster $\Cset$.
Our next main result (Theorem~\ref{thm: consistent_recovery_of_density_clusters})
shows that for sufficiently well-conditioned target clusters $\Cset$, the output 
of the PPR algorithm $\Cest$
is consistent in the sense defined in Definition~\ref{def: consistent_density_cluster_estimation}.

\subsection{Preliminaries}%: Geometric Conditions and the Condition Number of a Density Cluster}
%We formalize some geometric conditions, and use these to define a condition
%number $\kappa(\Cset)$, which measures the difficulty PPR will have in  
%estimating $\Cset$. (Our theoretical guarantees for PPR will be framed in terms 
%of $\kappa(\Cset)$.)
%\paragraph{Geometric conditions on density clusters.} 
At a high level, for PPR
to be successful, the underlying density cluster must be geometrically
well-conditioned.  A basic requirement is that we need to avoid clusters which contain arbitrarily
thin bridges or spikes. As in the work of \citet{chaudhuri2010}, we consider a
thickened version of \smash{$\Cset \in \Cbb_f(\lambda)$} defined as 
$\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq \sigma}$, which 
we call the \emph{$\sigma$-expansion} of $\Cset$. Here 
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$.

\begin{enumerate}[label=(A\arabic*)]
\item
  \label{asmp: bounded_density}
  \emph{Bounded density within cluster:} There exist constants
  $0<\lambda_{\sigma}< \Lambda_{\sigma}<\infty$ such that 
  $\lambda_{\sigma} \leq \inf_{x \in \Csig} f(x) \leq \sup_{x \in \Csig} f(x)
  \leq \Lambda_{\sigma}$.

\item
  \label{asmp: cluster_separation}
  \emph{Cluster separation:}
  For all clusters $\Cset' \in \Cbb_f(\lambda)$ with $\Cset' \not= \Cset$, 
  $\dist(\Csig,\Csig') > \sigma$, where $\dist(\Csig,\Csig') := \inf_{x
    \in \Csig} \dist(x,\Csig')$.

\item 
  \label{asmp: low_noise_density}
  \emph{Low noise density:} There exist $\gamma,c_0 > 0$ such that for 
  $x \in \Rd$ with $0 < \dist(x, \Csig) \leq \sigma$,   
  \begin{align*}
  \inf_{x' \in \Csig} f(x') - f(x) \geq  c_0 \dist(x, \Csig)^{\gamma}. 
  \end{align*}
  Roughly, 
  this assumption ensures that the density decays sufficiently quickly as we move away from
  the target cluster $\Csig$, and is a standard assumption in the level-set estimation literature
  (see for instance~\cite{singh2009}).

\item
  \label{asmp: embedding}
  \emph{Lipschitz embedding:}
  There exists $g: \Reals^d \to \Reals^d$ with the following properties: 
  \begin{enumerate}
  \item we have $\Csig = g(\mathcal{K})$, for a convex set $\mathcal{K} \subseteq \Rd$
  with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in \mathcal{K}}\norm{x - y} =:
  \diam < \infty$
  \item $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, where
  $\nabla g(x)$ is the Jacobian of $g$ evaluated at $x$ and
  \item for some $L
  \geq 1$,   
  \begin{equation*}
    \norm{g(x) - g(y)} \leq L \norm{x - y} ~
    \text{for all $x,y \in \mathcal{K}$}. 
  \end{equation*}
  \end{enumerate}
  Succinctly, we assume that $\Csig$ is the image of a convex set with finite diameter 
  under a  measure preserving, Lipschitz transformation.

\item
  \label{asmp: bounded_volume}
  \emph{Bounded volume:}
  We assume that neighborhood graph radius $0 < r \leq \sigma/2d$ is chosen such that
  \begin{equation*}
    2 \int_{\Csig} \Pbb(B(x,r)) f(x) dx \leq \int_{\Rd} \Pbb(B(x,r)) f(x) dx,
  \end{equation*}
  where $B(x,r)$ is the closed ball of radius $r$ at $x$.
\end{enumerate}

To motivate these conditions, we provide a brief high-level sketch
of our analysis (we return to this more formally in Section~\ref{sec: analysis}). 
\citet{zhu2013} show for an arbitrary graph $G =
(V,E)$ and subset of vertices $S \subseteq V$, the PPR estimate \smash{$\Cest$}
of subset $S$ satisfies, for a constant $c>0$,
\begin{equation}
\label{eqn: graph_symmetric_set_difference_1}
\vol(\Cest \setminus S; G) + \vol(S \setminus \Cest; G) \leq c \bigl( \Phi(S,G)
\cdot \tau_{\infty}(G[S])\bigr),
\end{equation}
where $\Phi(S;G)$ is the normalized cut of $S$ (as defined in \eqref{eqn:
  normalized_cut}), and \smash{$\tau_{\infty}(G[S])$} is  the \emph{mixing
  time} of a random walk over the induced subgraph $G[S]$ (to be defined
precisely later, in \eqref{eqn: mixing_time}).  The left-hand side in
\eqref{eqn: graph_symmetric_set_difference_1} resembles a (degree-weighted) 
form of the symmetric set difference metric in \eqref{eqn:
  misclassification_rate}, and our main goal will be to upper bound 
  this quantity. 
  Towards this goal, as we will show in Section \ref{sec:
  analysis}, the conditions \ref{asmp: bounded_density}--\ref{asmp:
  bounded_volume} allow us to upper bound the normalized cut 
\smash{$\Phi(\Csig[\Xbf]; G_{n,r})$}, and the mixing time 
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}. In more detail, assumptions  
\ref{asmp: cluster_separation}, \ref{asmp: low_noise_density} yield an upper
bound on \smash{$\cut(\Csig[\Xbf]; G_{n,r})$}, and \ref{asmp: bounded_density}
yields a lower bound on \smash{$\vol(\Csig[\Xbf]; G_{n,r})$}; together with 
\ref{asmp: bounded_volume}, this gives an upper bound on the normalized cut.  On 
the other hand, \ref{asmp: bounded_density}, \ref{asmp: embedding} preclude
bottlenecks in the induced subgraph \smash{$G_{n,r}[\Csig[\Xbf]]$}, and combined
with the upper bound on diameter in \ref{asmp: embedding}, this leads to an upper bound on
the mixing time over this subgraph.

\textcolor{red}{Mention Arias-Castro and Little here.}

\subsubsection{Condition number} 
Motivated by \eqref{eqn:
  graph_symmetric_set_difference_1}, we will define $\kappa(\Cset)$ to be an
upper bound on the product \smash{$\Phi(\Csig[\Xbf]; G_{n,r}) \cdot
  \tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}. Following, the analysis of \citet{zhu2013}
  we see that the smaller $\kappa(\Cset)$ is, the
more success PPR will have in recovering the target cluster $\Cset$. We let $\theta := (r, \sigma,
\lambda, \lambda_{\sigma}, \Lambda_{\sigma}, \gamma, \diam, L)$ contain the
geometric parameters from Assumptions~\ref{asmp:  bounded_density}--\ref{asmp:
  bounded_volume}.

\begin{definition}[Well-conditioned density clusters]
  \label{def:well_conditioned_density_cluster}
  For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy 
  \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some
  $\theta$. Then, for universal constants $c_1, c_2, c_3 > 0$ to be specified
  later, we set 
  \begin{equation}
    \label{eqn: condition_number_1}
    \Phibf(\theta) 
    := c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
    \frac{(\lambda_{\sigma} - c_0 \frac{r^{\gamma}}{\gamma +
        1})}{\lambda_{\sigma}},~  
    \taubf(\theta) := c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2
      L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{1}{r}\right) + c_3,
  \end{equation}
  and letting \smash{$\kappa(\Cset) := \Phi_{u}(\theta) \cdot
    \tau_{u}(\theta)$}, we call $\Cset$ a 
  \emph{$\kappa$-well-conditioned density cluster}.  
\end{definition}
\noindent The condition number 
$\kappa(\Cset)$ succinctly captures the role of the various
geometric parameters. 
We note in passing that $\Phibf(\theta)$ and $\taubf(\theta)$ are exactly the upper bounds 
on \smash{$\Phi(\Csig[\Xbf]; G_{n,r})$} and
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$} that we derive in our analysis
later, in Section \ref{sec: analysis}.

\subsubsection{Well-initialized algorithm} As is typical in the local clustering
literature, our algorithmic results will be stated with respect to specific
ranges of each of the user-specified parameters. In particular, for a
well-conditioned density cluster $\Cset$ (with respect to some $\theta$), we require that the tuning parameters of Algorithm~\ref{alg: ppr} are chosen to fall within the ranges
\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{2d}, \quad 
\alpha \in [1/10, 1/9] \cdot \frac{1}{\taubf(\theta)}, \\
v \in \Csig[\Xbf]^g, \quad
\pi_0 \in [3/4,5/4] \cdot \frac{\lambda_{\sigma}}{n\Lambda_{\sigma}\Pbb(\Csig)},
\end{gathered}
\end{equation}
where \smash{$\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$} will be some large
(``good'') subset of $\Csig[\Xbf]$. In particular, abbreviating $\vol_{n,r}(S)
:= \vol(S; G_{n,r})$ for $S \subseteq \Xbf$, this good subset has large volume, i.e. 
$\vol_{n,r}(\Csig[\Xbf]^g) \geq \vol_{n,r}(\Csig[\Xbf])/2$.

\begin{definition}
If the input parameters to Algorithm \ref{alg: ppr} satisfy \eqref{eqn:
  initialization} for some well-conditioned density cluster $\Cset$, we
say the algorithm is \emph{well-initialized}. 
\end{definition}
\noindent In practice it is  not feasible to set hyperparameters based on the
underlying (unknown) density $f$. Typically, one tunes PPR over a range of
hyperparameters and selects the cluster which has the smallest normalized cut.

\subsection{Cluster Recovery in Symmetric Set Difference}
We now present our first main result: a bound on the symmetric set difference metric between the estimated cluster $\widehat{C}$ and empirical cluster $\Csig[\Xbf]$.
\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$ and let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$). If Algorithm \ref{alg: ppr} is well-initialized, there exists a universal constant $c > 0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)  \frac{\Lambda_{\sigma}}{\lambda_{\sigma}} \abs{\Csig[\Xbf]}
  \end{equation}
\end{theorem}
The proof of Theorem \ref{thm: misclassification_rate} is deferred to Appendix~\ref{sec: proof_of_misclassification_rate}.
This result shows that the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.

\subsection{Consistent Cluster Recovery}
The bound on symmetric set difference \eqref{eqn: misclassification_rate_ub} does not imply consistent density cluster estimation in the sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of
consistency requires a uniform bound over $\pbf$: for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\pbf_v(w) \leq \frac{\pi_0}{40} < \frac{\pi_0}{11} \leq \pbf_v(u),
\end{equation}
so that any sweep cut $S_{\beta}$ for $\beta \in [1/40,1/11]$ (i.e., any
sweep cut considered by Algorithm \ref{alg: ppr}) will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures that as $n \to \infty$, such a gap exists with probability
1, and hence guarantees \smash{$\Cest$} will be a consistent  
estimator. We will need one additional regularity condition, to preclude arbitrarily low degree
vertices for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

Next we give our main result on consistent cluster recovery by PPR.

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >  0$
  such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{\lambda_{\sigma}^2r^d
      \nu_d}{\Lambda_{\sigma}\Pbb(\Csig)}, 
  \end{equation}
  then the output set $\Cest \subseteq \Xbf$ is a consistent estimator for
  $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}
\noindent Some remarks are in order: 
 \begin{enumerate}  
 \item We note that the restriction on $\kappa(\Cset)$ imposed by \eqref{eqn:
    kappa_ub} results in an upper bound on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
    \Cest)$} on the order of $r^d$. In plain terms, we are able to recover a
  density cluster $\Cset$ in the sense of \eqref{eqn:
    consistent_density_cluster_recovery} only when we can guarantee a very small
  fraction of points will be misclassified. This strong condition is the price
  we pay in order to obtain the uniform bound in~\eqref{eqn: ppr_gap}. 
 \item While taking the radius of the neighborhood graph $r \to 0$ as $n \to
  \infty$ (thereby ensuring that the graph $G_{n,r}$ is sparse) is computationally
  attractive, the presence of a factor of \smash{$\log^2(1/r)/r$} in
  $\kappa(\Cset)$ prevents us from making claims about the
  behavior of PPR in this regime. Although the restriction to a kernel
  function fixed in $n$ is standard for spectral clustering theory
  \citep{schiebinger2015,vonluxburg2008}, it is an interesting question whether
  PPR exhibits some degeneracy over $r$-neighborhood graphs as $r \to 0$, or if
  this is merely looseness in our upper bounds.  
\end{enumerate}

\subsection{Approximate PPR vector} 
In practice, exactly solving the system of equations~\eqref{eqn:
  ppr_vector} to compute the PPR vector 
  may be too computationally expensive. To address this limitation,
\citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector
(aPPR), which we will denote by \smash{$\pbf^{(\varepsilon)}$}. We refer the
curious reader to \citet{andersen2006} for a formal algorithmic definition of
the aPPR vector, and limit ourselves to highlighting a few salient points: the
aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time,
while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad \pbf(u) - \varepsilon \Dbf_{uu}\leq
\pbf^{(\varepsilon)}(u) \leq \pbf(u).  
\end{equation}
For a sufficiently small choice of $\varepsilon$, the 
application of \eqref{eqn: appr_error} within the proofs of Theorems \ref{thm:
  misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters}
leads to analogous results which hold for \smash{$\pbf^{(\varepsilon)}$}. We
formally state and prove this in Corollary~\ref{cor: appr} in 
Appendix~\ref{sec:appr}.

\section{Analysis Overview}
\label{sec: analysis}

The primary technical contribution of our work is showing that the geometric
conditions \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} translate to 
meaningful bounds on the normalized cut and mixing time of $\Csig[\Xbf]$ in
$G_{n,r}$. In doing so, we elaborate on how some of the geometric conditions
introduced in Section \ref{sec: consistent_cluster_estimation_with_ppr}
contribute to the difficulty of the clustering problem.

\subsection{Upper Bound on the Normalized Cut} 
We start with a finite sample upper bound on the
normalized cut \eqref{eqn: normalized_cut} of $\Cset_\sigma[\Xbf]$. For
simplicity, we write $\Phi_{n,r}(\Csig[\Xbf]) := \Phi(\Csig[\Xbf]; G_{n,r})$.

\begin{theorem}
  \label{thm: conductance_upper_bound}
  Fix $\lambda > 0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies
  Assumptions \ref{asmp: bounded_density}--\ref{asmp: low_noise_density}, 
  \ref{asmp: bounded_volume} for some $r, \sigma, \lambda_{\sigma}, c_0, \gamma
  > 0$ (no bound on maximum density is needed). Then for any $0 < \delta < 1$,
  $\epsilon > 0$, if 
  \begin{equation}
    \label{eqn: conductance_sample_complexity}
    n \geq \frac{(2+\epsilon)^2\log(3/\delta)}{\epsilon^2}\left(\frac{25}
      {6 \lambda_{\sigma}^2\nu(\Csig) \nu_d r^d}\right)^2,
  \end{equation}
  then with probability at least $1-\delta$, for a universal constant $c_1 > 0$,
  \begin{equation}
    \label{eqn: conductance_additive_error_bound}
    \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon.
  \end{equation}
\end{theorem}
\noindent {\bf Remark: } 
Observe that the diameter $\rho$ is absent from Theorem \ref{thm:
    conductance_upper_bound}, in contrast to the condition number
  $\kappa(\Cset)$, which worsens (increases) as $\rho$ increases. This
  reflects established wisdom regarding spectral partitioning
  algorithms more generally \citep{guattery1995, hein2010}, albeit newly applied
  to the density clustering setting. It suggests that if the diameter $\rho$ is
  large, PPR may fail to recover $\Csig[\Xbf]$ even when $\Cset$ is
  sufficiently well-conditioned to ensure $\Csig[\Xbf]$ has a small normalized
  cut in $G_{n,r}$. This intuition will be supported by simulations in Section
  \ref{sec: experiments}.

\subsection{Upper Bounds on the Mixing Time} 
For $S \subseteq V$, denote by $G[S] = (S, E_S)$ the
subgraph induced by $S$ (where the edges are $E_S = E \cap (S \times S)$). Let
$\Wbf_S$ be the (lazy) random walk matrix over $G[S]$, and write  
$$
q_{v}^{(t)}(u) = e_v\Wbf_S^t e_u
$$
for the $t$-step transition probability of the lazy random walk over $G[S]$
originating at $v \in V$. Also write \smash{$\pi = (\pi(u))_{u \in S}$}
for the stationary distribution of this random walk.  (As
$\Wbf_S$ is the transition matrix of a lazy random walk, 
it is well-known that a unique stationary distribution exists and is given by 
\smash{$\pi(u) = (\Dbf_S)_{uu}/\vol(S; G[S])$}, where we write $\Dbf_S$ for the  
degree matrix of $G[S]$.) We define the \emph{mixing time} of $G[S]$ as
\begin{equation}
\label{eqn: mixing_time}
\tau_{\infty}(G[S]) = \min\set{ t: \frac{\pi(u) - q_{v}^{(t)}(u)}
  {\pi(u)} \leq \frac{1}{4}, \; \text{for $u,v \in V$}}. 
\end{equation}
Next, we give an asymptotic (in the number of samples $n$) upper bound on
$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$.

\begin{theorem}
  \label{thm: mixing_time_upper_bound}
  Fix $\lambda > 0$, and assume that \smash{$\Cset \in \Cbb_f(\lambda)$}
  satisfies Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding}
  for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, \rho, L > 0$. Then, for
  any \smash{$0 < r < \sigma/2\sqrt{d}$}, with probability 1,
  \begin{equation} 
    \label{eqn: mixing_time_upper_bound}
    \limsup_{n \to \infty}\tau_{\infty}(G_{n,r}[\Csig[\Xbf]]) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{c_4 \Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + c_3
  \end{equation}
  for $c_2,c_3,c_4 > 0$ universal constants. 
\end{theorem}

The proof of Theorem \ref{thm: mixing_time_upper_bound} relies heavily on
analogous mixing time bounds developed for the mixing time of a
continuous-space ``ball walk'' over convex sets. To the best of our knowledge, our
result is the first bound, albeit asymptotic, on the mixing time of random walks
over neighborhood graphs that is independent of $n$, the number of vertices.

\vspace{0.2cm}

\noindent {\bf Remark: } 
  The embedding assumption \ref{asmp: embedding} and Lipschitz parameter $L$
  play an important role in proving the upper bound of Theorem \ref{thm:
    mixing_time_upper_bound}. There is some interdependence between $L$ and
  $\sigma,\rho$, which might lead one to hope that \ref{asmp: embedding} is
  non-essential. However, it is not possible to eliminate condition \ref{asmp:
    embedding} without incurring an additional factor of at least
  $(\rho/\sigma)^d$ in \eqref{eqn: mixing_time_upper_bound}, achieved, for
  instance, when $\Csig$ is a dumbbell-like set consisting of two balls of
  diameter $\rho$ linked by a cylinder of radius
  $\sigma$. \citet{abbasi-yadkori2016, abbasi-yadkori2016a}  
  develop theory regarding Lipschitz deformations of convex sets, wherein it
  is observed that star-shaped sets as well as half-moon shapes of the type we
  consider in Section \ref{sec: experiments} both satisfy \ref{asmp: embedding}
  for reasonably small values of $L$.

\section{Examples}
\label{sec: examples}

We illustrate the implications of our theory on some common generative models for nonparametric clustering, and compare our theoretical bounds on PPR to those derived for more classical density cluster estimators.

\subsection{Uniform clusters plus noise, dimension $d = 1$.}
\label{subsec: uniform_clusters}

We begin with a very simple model $\Pbb$ that fits ideally into our framework. For $i = 1,\ldots,n$, let $z_i \sim \delta \textrm{Bern}(1/2)$ be a latent class assignment, and let $w_i \sim \sigma \textrm{Unif}([-\sigma, \sigma])$ where $\delta$ and $\sigma$ are positive numbers such that $\delta > 3 \sigma$ (to ensure adequate separation of the density clusters). Assume we sample $x_1, \ldots, x_n$ from the following contaminated mixture model: 
\begin{equation*}
x_i = 
\begin{cases}
z_i + w_i,& \quad \textrm{with probability $1 - \epsilon$} \\
\widetilde{w}_i, & \quad \textrm{with probability $\epsilon$}
\end{cases}
\end{equation*}
where $\widetilde{w}_i \sim \textrm{Unif}(\sigma, \delta - \sigma)$ and $\epsilon$ is a positive number between $0$ and $1$. In other words, $\Pbb = (1-\epsilon)\bigl((1/2)\textrm{Unif}[-\sigma,\sigma] + (1/2)\textrm{Unif}[\delta - \sigma,\delta + \sigma]\bigr) + \epsilon \textrm{Unif}(\sigma,\delta - \sigma)$ is a mixture of two uniforms with added background noise. We will consider $\Cset = \set{0}$, so that the $\sigma$-expansion of $\Cset$ is $\Csig = [-\sigma,\sigma]$.\footnote{Although $\Cset$ is not a density level set of $\Pbb$, one can add a small bump to the density around $0$ to make $\set{0}$ a well defined density cluster for an appropriate choice of threshold $\lambda$, without changing any of the subsequent calculations. For ease of presentation, we ignore this detail.}

We verify that each of the assumptions \ref{asmp: bounded_density}-\ref{asmp: bounded_volume} hold on $\Csig$. The maximum and minimum density over $\Csig$ are the same, $\Lambda_{\sigma} = \lambda_{\sigma} = (1 - \epsilon)/(4\sigma)$. Letting $\Csig' = [\delta - \sigma, \delta + \sigma]$ be the $\sigma$-expansion around $\Cset' = \set{\delta}$, the condition $\dist(\Csig, \Csig') > \sigma$ holds since $\delta > 3\sigma$. The low noise density assumption holds with parameter $\gamma = 0$ and $c_0 = (1 - \epsilon)/(4\sigma) - \epsilon/2\sigma$. The set $\Csig$ is convex with diameter $\rho = 2\sigma$. Finally, for any $0 < r < \sigma$ the bounded volume condition of \ref{asmp: bounded_volume} holds.

We can now give an upper bound on misclassification error that holds with probability $1$ as $n \to \infty$. To do so, we begin by computing $\Phibf(\theta)$ and $\taubf(\theta)$ as defined in \eqref{eqn: condition_number_1},
\begin{equation*}
\Phibf(\theta) = \frac{c_1 r \epsilon}{(\delta - 2\sigma)(1 - \epsilon)}, \quad \taubf(\theta) = 4 c_2 \frac{\sigma^2}{r^2} \log\left(\frac{4c_4\sigma}{(1 - \epsilon)r}\right) + c_3.
\end{equation*}
Therefore, if we set $r = \sigma/2$, the largest value permissible under the conditions of Theorems \ref{thm: conductance_upper_bound} and \ref{thm: mixing_time_upper_bound}, we obtain the condition number
\begin{equation*}
\kappa(\Cset) = \frac{4c_1c_2\sigma\epsilon}{(\delta - 2\sigma)(1 - \epsilon)}\log^2\left(\frac{8c_4}{(1 - \epsilon)}\right) + \frac{c_3c_1\epsilon\sigma}{2(\delta - 2\sigma)}
\end{equation*}

Therefore, by Theorem \ref{thm: misclassification_rate}, if Algorithm \ref{alg: ppr} is well-initialized with respect to $\Cset$, then we have that the fraction of misclassified points $\frac{\Delta(\Cest, \Csig[\Xbf])}{\abs{\Csig[\Xbf]}}$ is roughly upper bounded by $\sigma \epsilon / (\delta - 2\sigma)$ (ignoring constants and logarithmic terms). Qualitatively speaking, this exhibits the correct dependence on the mixture model parameters $\sigma$, $\epsilon$, and $\delta$. In particular, as the expected fraction of background noise $\epsilon$, or the ratio of cluster width to cluster separation $\sigma/(\delta - 2\sigma)$, decrease, so does our estimate of the symmetric set difference $\Delta(\Cest, \Csig[\Xbf])$.

\subsection{Uniform clusters plus noise, general dimension.}

To better demonstrate the flexibility of our theory, we now generalize the previous example in several respects. For an integer $d' < d$, consider the $d'$-dimensional unit cube embedded in $\Reals^d$ by letting
\begin{equation*}
[0,1]^{d'} = \{(x_1,\ldots,x_d) \in \Rd: 0 \leq x_i \leq 1 ~\textrm{for}~ i = 1,\ldots,d', x_i = 0 ~\textrm{for}~ i = d'+ 1,\ldots,d\},
\end{equation*}
and denoting the $\sigma$-expansion of $[0,1]^{d'}$ by $[0,1]_{\sigma}^{d'} = \{x \in \Rd: \dist(x,[0,1]^{d'}) < \sigma\}$. For a number of clusters $K > 2$, let $f_1,\ldots,f_K: [0,1]_{\sigma}^{d'} \to [0,1]^{d}$ be Lipschitz, measure-preserving maps, with Lipschitz constant $\norm{f_k}_{\textrm{Lip}}, ||f_k^{-1}||_{\textrm{Lip}} \leq L$ for each $k$. Write $\Csig^k$ for the image $f_k([0,1]_{\sigma}^{d'})$, and let $\Psi_k$ be the uniform distribution over $\Csig^k$. Additionally, let $\Psi_0$ be the uniform distribution over $\Cset^0 = [0,1]^d \setminus \bigcup_{k = 1}^{K} \Csig^k$. For $\epsilon \in [0,1)$, let $\Pbb$ be the $K$-component mixture model given by
\begin{equation*}
\Pbb = \frac{(1 - \epsilon)}{K} \sum_{k = 1}^{K} \Psi_k + \epsilon \Psi_0. 
\end{equation*}

Suppose we observe $x_1,\ldots,x_n$ sampled i.i.d from $\Pbb$. Letting $f$ be the density function of $\Pbb$, for sufficiently small $\epsilon > 0$, and an appropriate choice of $\lambda$, the density upper level set $\Cbb_f(\lambda)$ will contain components $\Csig^1,\ldots,\Csig^k$. Let's say the set we are interested in recovering is $\Csig^1$. Under certain regularity conditions, this framework fits smoothly into the theory developed Sections~\ref{sec: consistent_cluster_estimation_with_ppr} and \ref{sec: analysis}. We will assume that for each $k$, $\Csig^k$ is a $\sigma/2$-expansion\footnote{Note that it is possible that $\Csig^k = f_k([0,1]_{\sigma}^{d'})$ for a Lipschitz function $f_k$, and yet itself not be an $\sigma/2$-expansion of any set. As we need $\Csig^k$ to be a $\sigma$-expansion, we make this requirement explicit.} of some $\Cset^k \subset [0,1]^d$ and additionally that
\begin{equation*}
\min_{k \neq \ell} \dist(\Csig^k,\Csig^\ell) > \sigma, \min_{k}\dist(\Csig^k,\partial [0,1]^d) > \sigma, \quad \textrm{and}~ 5 < K < \frac{\sigma^{d - d'}}{2^{d' + 1}}.
\end{equation*}

Now, we verify that each of the assumptions \ref{asmp: bounded_density}-\ref{asmp: bounded_volume} hold with respect to $\Csig^1$. 
\begin{itemize}
	\item \ref{asmp: bounded_density}. In this case, the minimum and maximum density within the $\sigma/2$-expansion $\Csig^1$ are the same, $\Lambda_{\sigma} = \lambda_{\sigma} = (1 - \epsilon)/(K \nu(\Csig^1))$. (Indeed, this holds for each $\Csig^k$.) 
	\item \ref{asmp: cluster_separation}. By assumption, $\dist(\Csig^1,\Csig^k) > \sigma$ for each $k = 2,\ldots,K$. 
	\item \ref{asmp: low_noise_density}. The low noise condition is satisfied with $\gamma = 0$ and $c_0 = (1 - \epsilon)/(K \nu(\Csig^1)) - \epsilon/(\nu(\Cset^0))$. 
	\item \ref{asmp: embedding} By assumption, the embedding condition is satisfied with $\rho = 1 + \sigma < 2$, and $L = L$. 
	\item \ref{asmp: bounded_volume}. When $r < \sigma/(2d)$ we have that $\nu( B(x,r) \cap \Csig^k)\geq \frac{6}{25}\lambda_{\sigma}\nu_d r^d$ (see Lemma~\ref{lem: local_conductance}). Therefore
	\begin{align*}
	\sum_{k = 2}^{K} \int_{\Csig^k} \Pbb(B(x,r)) \,dx & \geq \frac{6}{25}\sum_{k = 2}^{K}\nu(\Csig^k) \lambda_{\sigma} \nu_d r^d \\
	& \overset{(i)}{\geq} \nu(\Csig^1) \lambda_{\sigma} \nu_d r^d \\
	&  \geq \int_{\Csig^1} \Pbb(B(x,r)) \,dx.
	\end{align*} 
	where $(i)$ follows from the fact that $\nu(\Csig^k) = \nu([0,1]_{\sigma}^{d'})$ for all $k$ and the assumption that $K \geq 5$. Therefore \ref{asmp: bounded_volume} is satisfied.
\end{itemize}

The restriction that $f$ be measure-preserving implies $\nu(\Csig^k) = \nu([0,1]_{\sigma}^{d'}) \leq 2^{d'} \sigma^{d - d'}$ for each $k$. Along with the upper bound on number of clusters $K$, we additionally have $\nu(\Cset^0) \geq (1 - K 2^{d'} \sigma^{d - d'}) \geq 1/2$. We can therefore bound $\Phibf(\theta)$ and $\taubf(\theta)$,
\begin{equation}
\label{eqn:normalized_cut_mixing_time_ub_example2}
\Phibf(\theta) \leq c \frac{K dr \epsilon}{\sigma 2^{d'} \sigma^{d - d'} (1 - \epsilon)}, \quad \Psibf(\theta) \leq c \frac{d^3 L^2}{r^2} \log^2\left(\frac{K \sigma^{d - d'}}{(1 - \epsilon)r}\right),
\end{equation}
and letting $r = \sigma/(2d)$, we obtain the following bound on the condition number $\kappa(\Cset)$:
\begin{equation}
\label{eqn:kappa_example2}
\kappa(\Cset) \leq c \frac{\epsilon \sigma^{d - d' - 2} K d^5 L^2}{(1 - \epsilon)} \log^2\left(\frac{d K \sigma^{d - d'}}{\sigma(1 - \epsilon)}\right),
\end{equation}
which likewise implies a bound on the symmetric set difference metric $\Delta(\Cest,\Csig[\Xbf])$ (assuming that Algorithm~\ref{alg: ppr} is well-initialized.)

\subsubsection{Plug-in density clustering.}
It is interesting to compare \eqref{eqn:kappa_example2} to the symmetric set difference obtained by a more classical density cluster estimator. In particular, we consider the plugin approach to density cluster estimation, which we now briefly summarize (see e.g. \cite{chaudhuri2010} for a more formal algorithmic description). To motivate this approach, note that one can view the degrees $\deg_{n,r}(x_1), \ldots, \deg_{n,r}(x_n)$ in the graph $G_{n,r}$ as estimates of the density evaluated at data $f(x_1),\ldots,f(x_n)$. Therefore, a natural way to identify points in high-density regions is to throw out points without sufficiently high degree. To identify the high-density regions $\Csig^1,\ldots,\Csig^k$, consider the threshold
\begin{equation*}
\lambda' = \frac{\epsilon}{\nu(\Cset^0)} + \frac{c_0}{2}.
\end{equation*}
For any $x_i \in \bigcup_{k = 1}^{K}\Csig^k[\Xbf]$ such that $\dist(x_i, \Cset^0) > r$, the expected degree of $x_i$ is lower bounded $\Ebb(\deg_{n,r}(x_i)) = \lambda_{\sigma} \nu_d r^d = (\lambda' + \frac{c_0}{2})\nu_dr^d$. On the other hand, for any $x_i \in \Cset^0[\Xbf]$ such that $\dist(x_i, \bigcup_{k = 1}^{K}\Csig^k[\Xbf]) > r$, the expected degree of $x_i$ is upper bounded $\Ebb(\deg_{n,r}(x_i)) \leq (\lambda' - \frac{c_0}{2})\nu_dr^d$. Let $n$ be large enough to allow for concentration of the degrees around their means. Then the set $C_n(\lambda') = \set{x_i \in \Xbf: \deg_{n,r}(x_i) > \lambda'\nu_dr^d}$ serves as a reasonable estimate of $\bigcup_{k = 1}^{K} \Cset_{\sigma}^k[\Xbf]$, in the sense that with high probability $\bigcup_{k = 1}^{K} \Cset_{\sigma - r}^k[\Xbf] \subseteq C_n(\lambda') \subseteq \bigcup_{k = 1}^{K} \Cset_{\sigma + r}^k[\Xbf]$

For a sufficiently large choice of radius $r$, we can draw the additional conclusion that the subsets $\Cset_{\sigma - r}^k[\Xbf]$ are connected in $G_{n,r}$, and therefore in the graph $G_{n,r}[C_n(\lambda')]$ as well. This implies that there are connected components $C^1, \ldots, C^K$ of $G_{n,r}[C_n(\lambda')]$ which are good estimates of $\Csig^1[\Xbf],\ldots,\Csig^K[\Xbf]$, since with high probability $\Cset_{\sigma - r}^k[\Xbf] \subseteq C^k \subseteq \Cset_{\sigma + r}^k[\Xbf]$.

Those points $x_i \in \Xbf$ which are in $\Cset_{\sigma + r}^k$ but not $\Cset_{\sigma}^k$ are those points for which $\Ebb(\deg_{n,r}(x_i))$ may be  approximately equal to $\lambda'$, and so we cannot rule out their belonging to $C^k$. The symmetric set difference $\Delta(C^k, \Csig^k[\Xbf])$ is therefore roughly equal to the cardinality of $\{x_i \in X: 0 < \dist(x_i,\Csig^k) \leq r\}$.

To compare this to \eqref{eqn:kappa_example2}, we must provide an estimate of the cardinality of $\{x_i \in X: 0 < \dist(x_i,\Csig^k) \leq r\}$. By Lemma~\ref{lem: expected_number_boundary_points}, we have the following upper bound on the expected number of points in this set,
\begin{equation*}
\Ebb \abs{\{x_i \in X: 0 < \dist(x_i,\Csig^k) \leq r\}} \leq \frac{4ndr}{\sigma}\left(\lambda_{\sigma} - c_0\right)\nu(\Csig).
\end{equation*}
Combined with the fact that $\Ebb \abs{\Csig^k[\Xbf]} = n \lambda_{\sigma} \nu(\Csig)$--and using our bounds on $\lambda_{\sigma}$ and $c_0$ from the previous section--we obtain
\begin{equation*}
\frac{\Ebb\abs{\{x_i \in X: 0 < \dist(x_i,\Csig^k) \leq r\}} }{\Ebb \abs{\Csig^k[\Xbf]}} \leq c \frac{K dr \epsilon}{\sigma 2^{d'} \sigma^{d - d'} (1 - \epsilon)},
\end{equation*}
where we recognize the right hand side as being identical to the bound on normalized cut we derived in \eqref{eqn:normalized_cut_mixing_time_ub_example2}.

Let us summarize our findings with respect to the particular density cluster $\Csig^1$. First, with high probability, only those points $x_i \in X$  such that $0 < \dist(x_i,\Csig^1) \leq r$ will be misclassified by the plug in approach. Second, the expected ratio of the number of such points, to the number of total data points that fall in $\Csig^1$, is upper bounded by the same upper bound that we derived on the normalized cut $\Phi_{n,r}(\Csig^1[\Xbf];G_{n,r})$. In other words, for a particular choice of radius $r$, the symmetric set difference $\Delta(C^1,\Csig^1[\Xbf])/\abs{\Csig^1[\Xbf]}$ is comparable to the normalized cut of $\Csig^1[\Xbf]$ in $G_{n,r}$, whereas the symmetric set difference $\Delta(C^1,\Csig^1[\Xbf])/\abs{\Csig^1[\Xbf]}$ is upper bounded by the \emph{product} of the normalized cut and the mixing time, which is a looser bound. This comparison elucidates an important distinction between estimation of density clusters via PPR, and more classical methods based on plug in estimators. The latter approach is robust to poor internal connectivity of the $\sigma$-expansion $\Csig$, whereas the former may not be.

\section{Experiments}
\label{sec: experiments}

We provide numerical experiments to investigate the tightness of our bounds on
the normalized cut and mixing time of $\Csig[\Xbf]$, and examine the performance
of PPR on the ``two moons'' dataset. We defer details of the experimental
settings to the appendix.

\begin{figure}[tb]
  \centering
  % \begin{adjustbox}{minipage=\linewidth}
  %   \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.32\textwidth]{example1plots/sample2}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.32\textwidth]{example1plots/sample1}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.32\textwidth]{example1plots/sigma_normalized_cut_plot}
  % \caption{}
  % \end{subfigure}

  % \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.32\textwidth]{example1plots/sigma_mixing_time_plot}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.32\textwidth]{example1plots/diameter_normalized_cut_plot}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.32\textwidth]{example1plots/diameter_mixing_time_plot}
  % \caption{}
  % \end{subfigure}
  \caption{\it\small Top left and top middle: samples from a geometrically
    well- and poor-conditioned cluster. The points in $\Cset$ are colored in 
    red, points in $\Csig \setminus \Cset$ are colored in yellow, and the
    remaining points in blue. Other panels: empirical normalized cut and mixing
    time, as a function of $\sigma$ or $\rho$, versus their theoretical upper
    bounds.} 
  \label{fig:fig1}
%  \end{adjustbox}
\end{figure}

\paragraph{Validating theoretical bounds.}  We investigate the tightness of
Theorems \ref{thm: conductance_upper_bound} and \ref{thm:
  mixing_time_upper_bound} via simulation. Figure \ref{fig:fig1} compares our
upper bounds with the actual empirically-computed quantities \eqref{eqn:
  normalized_cut} and \eqref{eqn: mixing_time}, as we vary the diameter $\rho$
and thickness $\sigma$ of a cluster $\Cset$. The top left and top middle panels
display the resulting empirical clusters for two different values of
$\rho,\sigma$.

The bottom left and bottom right panels assure that our mixing  
time upper bounds track closely the empirical mixing time, in both 2 and 3 
dimensions.\footnote{We rescaled all values of theoretical upper
  bounds by a constant, tto mask the effect of large universal constants
  in these bounds. Therefore only the comparison of slopes, rather than
  intercepts, is meaningful.} This provides empirical evidence that Theorem
\ref{thm: mixing_time_upper_bound} has the right dependency on both expansion
parameter $\sigma$ and diameter $\rho$. The story for the normalized cut panels
is less obvious. We remark that while, broadly speaking, the trends do not
appear to match, this gap between theory and empirical results seems largest
when $\sigma $ and $\rho$ are approximately equal. As the ratio $\rho/\sigma$
grows, the slopes of empirical and theoretical curves become more similar.

\paragraph{Empirical behavior of PPR.} In Figure \ref{fig:fig2}, to drive home
the main implications of Theorems \ref{thm: misclassification_rate} and
\ref{thm: consistent_recovery_of_density_clusters}, we show the
behavior of PPR, normalized cut, and the density clustering algorithm of
\citet{chaudhuri2010} on the well known ``two moons'' dataset (with added 2d 
Gaussian noise), considered a prototypical success story for spectral clustering
algorithms. The first column shows the empirical density clusters $\Cset[\Xbf]$
and $\Cset'[\Xbf]$ for a particular threshold $\lambda$ of the density function; the 
second column shows the cluster recovered by PPR; the third column shows the
global minimum normalized cut, computed according to the algorithm of
\citet{szlam2010}; and the last column shows a cut of the density cluster tree
estimator of \citet{chaudhuri2010}.  We can see the degrading ability of PPR to
recover density clusters as the two moons become less well-separated. Of 
particular interest is the fact that PPR fails to recover one of the moons even
when normalized cut still succeeds in doing so. Additionally, we note that the Chaudhuri-Dasgupta algorithm succeeds even when both PPR and normalized cut fail.  While our main message was
that PPR recovers geometrically well-conditioned density clusters, it would be  
interesting to establish that it \emph{only} recovers such clusters, a direction
for future work.

\begin{figure}
  \centering
  % \begin{adjustbox}{minipage=\linewidth}
  %   \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth,scale = .5]{example2plots/row1_true_density_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row1_ppr_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row1_conductance_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row1_density_cluster}
  % \caption{}
  % \end{subfigure}

  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row2_true_density_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row2_ppr_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row2_conductance_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row2_density_cluster}
  % \caption{}
  % \end{subfigure}

  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row3_true_density_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row3_ppr_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row3_conductance_cluster}
  % \caption{}
  % \end{subfigure}
  % \begin{subfigure}{.24\linewidth}
  \includegraphics[width=0.24\textwidth]{example2plots/row3_density_cluster}
  % \caption{}
  % \end{subfigure}
  \caption{\it\small True density (column 1), PPR (column 2), normalized
    cut (column 3) and estimated density (column 4) clusters for 3 different 
    simulated data sets. Seed node for PPR denoted by a black cross.} 
  \label{fig:fig2}
  %\end{adjustbox}
\end{figure}

\section{Discussion}
\label{sec: discussion}

There are an almost limitless number of ways to define what the ``right''
clustering is. In this paper, we have considered one such notion---density upper
level sets---and have detailed a set of natural geometric criteria which, when 
appropriately satisfied, translate to provable bounds on estimation of the
cluster by PPR. We do not, however, provide theory to show that (some variant of) 
our geometric
conditions are \emph{required} for successful recovery of a density level set. This
seems to be supported by our empirical results, and establishing this necessity
formally is an important direction for 
future work.

\section{Acknowledgements}
SB is grateful to Peter Bickel, Martin Wainwright and Larry Wasserman for helpful and inspiring conversations. This work was supported in part by the NSF grant DMS-1713003.

\appendix

%In this supplement, we present proofs and experimental details for the article ``Local Spectral Clustering of Density Upper Level Sets''.
\section{Proof of Theorem \ref{thm: conductance_upper_bound}}
\label{sec: proof_of_theorem_1}

For notational ease, we write
\begin{align*}
\cut_{n,r} = \cut(\Csig[\Xbf]; G_{n,r}), ~ \mu_K = \mathbb{E}(\cut_{n,r}), ~ p_K = \frac{\mu_K}{{n \choose 2}} \\
\vol_{n,r} = \vol(\Csig[\Xbf]; G_{n,r}), ~ \mu_V = \mathbb{E}(\vol_{n,r}), ~ p_V = \frac{\mu_V}{{n \choose 2}} \\
\vol_{n,r}^c = \vol(\Xbf \setminus \Csig[\Xbf]; G_{n,r}), ~ \mu_V^c = \mathbb{E}(\vol_{n,r}^c), ~ p_V^c = \frac{\mu_V^c}{{n \choose 2}}
\end{align*}
for the random variable, mean, and probability of cut size and volume, respectively.

With this notation in place, the goal of Theorem~\ref{thm: conductance_upper_bound} is roughly to show that, with probability at least $1 - \delta$, for a universal constant $c_1 > 0$ and for any $\epsilon > 0$, for a sufficiently large sample size $n$:
\begin{align*}
\Phi_{n,r}(\Csig[\Xbf]) := \frac{\cut_{n,r}}{\min\{\vol_{n,r}, \vol_{n,r}^c\}} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon.
\end{align*}

The proof of this theorem follows essentially from two technical Lemmas. 
The following Lemma relates the terms in the numerator and denominator of $\Phi_{n,r}(\Csig[\Xbf])$ to their expected values.
\begin{lemma}
	\label{lem: prob_bound_cutvol}
	For any $\delta \in (0,1]$,
	\begin{equation}
	\label{eqn: numerator_additive_bound}
	\frac{\cut_{n,r}}{{n \choose 2}} \leq p_K + \sqrt{\frac{\log(1/\delta)}{n}}, \\ 
	\frac{\vol_{n,r}}{{n \choose 2}} \geq p_V - \sqrt{\frac{\log(1/\delta)}{n}}, \\
	\frac{\vol_{n,r}^c}{{n \choose 2}} \geq p_V^c - \sqrt{\frac{\log(1/\delta)}{n}},
	\end{equation}
	each with probability at least $1 - \delta$. 
\end{lemma}
We note that as a consequence of~\ref{asmp: bounded_volume} we have that $p_V^c \geq p_V$, so it will suffice to lower bound $p_V$ (since a lower bound for $p_V^c$ follows).
The 
following result provides upper and lower bounds on the expected values $p_K$ and $p_V$ respectively:
\begin{lemma}
	\label{lem: expected_density_cut}\label{lem: expected_density_volume}
	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
	\begin{align}
	p_K &\leq \frac{4 d \nu_d r^{d+1} \lambda}{\sigma} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig), \label{eqn:claim_one} \\
	p_V &\geq \frac{12}{25} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig).\label{eqn:claim_two}
	\end{align}
\end{lemma}
%\begin{lemma}
%	\label{lem: expected_density_volume}
%	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
%	\begin{equation*}
%	p_V \geq \frac{12}{25} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig).
%	\end{equation*}
%\end{lemma}

Taking these two technical lemmas as given we can now complete the proof of the theorem. 
We fix $\delta \in (0,1]$ and let $\delta' = \delta/3$. We rewrite
\begin{equation}
\label{eqn: conductance_representation_1}
\Phi_{n,r}(\Csig[\Xbf]) = \frac{p_K + \left(\frac{\cut_{n,r}}{{n \choose 2}} - p_K\right)}{p_V + \left(\frac{\min\set{\vol_{n,r}, \vol_{n,r}^c}}{{n \choose 2}} - p_V\right)}.
\end{equation}
By a union bound we see that all three bounds in \eqref{eqn: numerator_additive_bound} hold with probability at least $1 - \delta$. Along with (\ref{eqn: conductance_representation_1}) this yields the upper bound
\begin{equation*}
\Phi_{n,r}(\Csig[\Xbf]) \leq \frac{p_K + \Err_n}{p_V - \Err_n}
\end{equation*}
for $\Err_n = \sqrt{\frac{\log(1/\delta')}{n}}$.
Now, some straightforward algebraic manipulations yield,
\begin{align*}
\frac{p_K + \Err_n}{p_V - \Err_n} & = \frac{p_K}{p_V} \left(\frac{p_V}{p_V - \Err_n}\right) + \frac{\Err_n}{p_V - \Err_n} \\
& = \frac{p_K}{p_V} + \left(\frac{p_K}{p_V} + 1\right)\frac{\Err_n}{p_V - \Err_n} \\
& \leq \frac{p_K}{p_V} + 2 \frac{\Err_n}{p_V - \Err_n}.
\end{align*}

%The proof of Theorem \ref{thm: conductance_upper_bound} is more or less given by Lemmas \ref{lem: expected_density_cut}, \ref{lem: expected_density_volume}, and \ref{lem: prob_bound_cutvol}. All that remains is some algebra, which we take care of below.

\noindent Using Lemma \ref{lem: expected_density_cut}, we have
\begin{equation*}
\frac{p_K}{p_V} \leq \frac{100rd}{12\sigma} \frac{\lambda}{\lambda_{\sigma}} \frac{\left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right)}{\lambda_{\sigma}}.
\end{equation*}
Then, by the choice of sample size in \eqref{eqn: conductance_sample_complexity}, 
\begin{equation*}
n \geq \frac{(2 + \epsilon)^2 \log\left(\frac{3}{\delta}\right)}{\epsilon^2 p_V^2}
\end{equation*}
which implies $2 \frac{\Err_n}{p_V - \Err_n} \leq \epsilon$, and this yields the desired bound on $\Phi_{n,r}(\Csig[X])$. In order to complete the proof of
the theorem it remains to prove Lemmas~\ref{lem: prob_bound_cutvol} and~\ref{lem: expected_density_cut}.

\subsection{Proof of Lemma~\ref{lem: prob_bound_cutvol}}
We first note that we can write $\cut_{n,r}$ as a double sum,
	\begin{equation}
	\label{eqn: density_cut_expansion}
	\cut_{n,r} = \sum_{i = 1}^{n} \sum_{j \neq i} \1(x_i \not\in \Csig) \1(x_j \in \Csig) \1(\norm{x_i - x_j} \leq r),
	\end{equation}
	and similarly we can write $\vol_{n,r}$ as a sum of indicator functions,
	\begin{equation}
	\label{eqn: volume_expansion}
	\vol_{n,r} = \sum_{i = 1}^{n} \sum_{j \neq i} \1(x_i \in \Csig) \1(x_j \in B(x_i, r)).
	\end{equation}
From this we see that $\cut_{n,r}$ and $\vol_{n,r}$, properly scaled, can be expressed as order-$2$ $U$-statistics,
	\begin{equation*}
	\frac{\cut_{n,r}}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_K(x_i, x_j),~~ \frac{\vol_{n,r}}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_V(x_i, x_j)
	\end{equation*}
	with kernels
	\begin{align*}
	\phi_K(x_i,x_j) & = \1(x_i \in \Csig, x_j \not\in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \in \Csig, x_i \not\in \Csig, \norm{x_i - x_j} \leq r) \\
	\phi_V(x_i,x_j) & = \1(x_i \in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \in \Csig, \norm{x_i - x_j} \leq r). 
	\end{align*}
	Similarly,
	\begin{equation*}
	\frac{\vol_{n,r}^c}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_{V^c}(x_i, x_j)
	\end{equation*}
	with kernel,
	\begin{equation*}
	\phi_{V^c}(x_i,x_j) = \1(x_i \not\in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \not\in \Csig, \norm{x_i - x_j} \leq r). 
	\end{equation*}
	Appealing to a standard concentration inequality for U-statistics (see Lemma \ref{lem: bounded_difference}) we therefore have
	\begin{equation*}
	\frac{\cut_{n,r}}{{n \choose 2}} \leq p_K + \sqrt{\frac{\log(1/\delta)}{n}},~  \frac{\vol_{n,r}}{{n \choose 2}} \geq p_V - \sqrt{\frac{\log(1/\delta)}{n}}, ~ \frac{\vol_{n,r}^c}{{n \choose 2}} \geq p_V^c - \sqrt{\frac{\log(1/\delta)}{n}}
	\end{equation*}
	each with probability at least $1 - \delta$. 
	%The claim follows in light of \ref{asmp: bounded_volume}, which implies $p_V^c \geq p_V$.

\subsection{Proof of Lemma~\ref{lem: expected_density_cut}}
We write $\Pbb(\Aset) = \int_{\Aset} f(x) dx$ for measurable $\Aset \subseteq \Rd$.
We 
let $\Csigr := \set{x: 0 < \dist(x, \Csig) < r}$, where $\Csig$ is as in Theorem \ref{thm: conductance_upper_bound}. 
Our goal will be to upper bound $p_K$ by a term that depends on the probability mass $\Pbb(\Csigr)$, 
and the bulk of our technical effort will be devoted to showing the following upper bound on $\Pbb(\Csigr)$:
%Lemma \ref{lem: expected_number_boundary_points} involves the bulk of the technical effort required to prove Theorem \ref{thm: conductance_upper_bound}; it will be necessary to bound the expected cut size of $\Csig[\Xbf]$ in $G_{n,r}$. 
\begin{lemma}
	\label{lem: expected_number_boundary_points}
	Under the conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
	\begin{equation*}
	\Pbb(\Csigr) \leq \frac{2dr}{\sigma} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig)
	\end{equation*}	
\end{lemma}
\noindent Define the \emph{uniform local conductance} $\ell_{\nu,r}(u)$ to be
\begin{equation*}
\ell_{\nu,r}(u) = \nu\bigl(\Csig \cap B(u,r)\bigr).
\end{equation*}
In order to lower bound $p_V$ we will show the following lower bound on the uniform local conductance:
\begin{lemma}
	\label{lem: local_conductance}
	Let $u \in \Csig$. Then, for any $0 < r \leq \frac{\sigma}{2\sqrt{d}}$,
	\begin{equation*}
	\ell_{\nu,r}(u) \geq \frac{6}{25} \nu_d r^d.
	\end{equation*}
\end{lemma}
Taking these two results as given we can now prove each of the two claims~\eqref{eqn:claim_one} and~\eqref{eqn:claim_two} in turn.
\paragraph{Proof of Claim~\eqref{eqn:claim_one}: } Recalling,~\eqref{eqn: density_cut_expansion} 
	and by the linearity of expectations, we obtain
	\begin{equation*}
	p_K = \frac{\mu_K}{{n \choose 2}} = 2 \cdot \Pbb(x_i \not\in \Csig, x_j \in \Csig, \norm{x_i - x_j} \leq r). \tag{for each $i,j$, $i \neq j$}
	\end{equation*}
	Writing this with respect to the density function $f$, we have
	\begin{align*}
	p_K & = 2 \int_{\Rd \setminus \Csig} f(x) \Pbb\bigl(B(x,r) \cap \Csig\bigr) \dx \\
	& = 2 \int_{\Csigr} f(x) \Pbb\bigl(B(x,r) \cap \Csig\bigr) \dx \\
	& \leq 2 \nu_d r^d \lambda  \int_{\Csigr} f(x) \dx = 2 \nu_d r^d \lambda \Pbb(\Csigr).
	\end{align*}
	where the inequality follows from \ref{asmp: cluster_separation}, which implies $f(x) \leq \lambda$ for $x \in \Csig \setminus \Cset$. Then, upper bounding the integral using Lemma \ref{lem: expected_number_boundary_points} gives the final result.

%\subsection{Proof of Lemma~\ref{Z}}
\paragraph{Proof of Claim~\eqref{eqn:claim_two}: } Recalling,~\eqref{eqn: volume_expansion}
%The proof will proceed similarly to Lemma \ref{lem: expected_density_cut}. 
	and by the linearity of expectation we obtain
	\begin{equation*}
	p_V = \frac{\mu_V}{{n \choose 2}} = 2 \cdot \Pbb(x_i \in \Csig, x_j \in B(x_i,r)). \tag{for any $i,j$, $i \neq j$. }
	\end{equation*}
	Writing this with respect to the density function $f$, we have
	\begin{align*}
	p_V & = 2 \int_{\Csig} f(x) \Pbb(B(x,r)) \dx \\
	& \geq 2 \int_{\Csig} f(x) \Pbb(B(x,r) \cap \Csig) \dx
	\end{align*}
	whence the claim then follows by Lemma \ref{lem: local_conductance}.
It remains to prove Lemmas~\ref{lem: expected_number_boundary_points} and~\ref{lem: local_conductance}, and we turn our attention to this next.
\subsection{Proof of Lemma~\ref{lem: expected_number_boundary_points}}
The proof of this Lemma relies on certain volume estimates whose statement and proof we defer to Appendix~\ref{sec: volume_estimates}.
	We partition $\Csigr$ into slices based on distance from $\Csig$ as follows: for $k \in \N$,
	\begin{equation*}
	\mathcal{T}_{i,k} = \set{x \in \Csigr: t_{i,k} < \frac{\dist(x, \Csig)}{r} \leq t_{i+1,k}}, ~~ \Csigr = \bigcup_{i = 0}^{k-1} \mathcal{T}_{i,k},
	\end{equation*}
	where $t_{i,k} = i/k$ for $i = 0, \ldots, k - 1$. As a result, for any $k \in \mathbb{N}$,
	\begin{equation}
	\label{eqn: partition_ub}
	\Pbb(\Csigr) = \int_{\Csigr} f(x) \dx = \sum_{i = 0}^{k-1} \int_{\mathcal{T}_{i,k}} f(x) \dx \leq \sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x).
	\end{equation}
	Assumptions~\ref{asmp: bounded_density} and \ref{asmp: low_noise_density} imply the upper bound
	\begin{equation*}
	\max_{x \in \mathcal{T}_{i,k}} f(x) \leq \lambda_{\sigma} - c_0(rt_{i,k})^{\gamma},
	\end{equation*}
	and writing
	\begin{equation*}
	\nu(\mathcal{T}_{i,k}) = \nu(\Csig + rt_{i+1,k}B) - \nu(\Csig + rt_{i,k}B) =: \nu_{i+1,k} - \nu_{i,k},
	\end{equation*}
	we have
	\begin{align}
	\label{eqn: telescoping_sum}
	\sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x) & \leq \sum_{i = 0}^{k-1} \biggl\{ \nu_{i+1,k} - \nu_{i,k} \biggr\} \biggl( \lambda_{\sigma} - c_0(rt_{i,k})^{\gamma} \biggr) \nonumber \\
	& = \underbrace{\sum_{i = 1}^{k} 
	\nu_{i,k} \biggl( \left[\lambda_{\sigma} - c_0(rt_{i-1,k})^{\gamma}\right] -  \left[\lambda_{\sigma} - c_0(rt_{i,k})^{\gamma}\right]\biggr)}_{:= \Sigma_k} + \underbrace{\biggl(\nu_{k,k}\left[\lambda_{\sigma} - c_0r^{\gamma}\right] - \nu_{1,k}\lambda_{\sigma} \biggr)}_{:= \xi}
	\end{align}
	where the second equality comes from rearranging terms in the sum.

	% AJG 5/14/19: This argument makes sense, correct?
	We first consider the term $\Sigma_k$. $\Cset$ has finite diameter by Assumption~\ref{asmp: bounded_density}, as otherwise $\int_{\Csig} f(x) dx = \infty$. Letting $\overline{\Cset}$ be the closure of $\Cset$, we observe that $\overline{\Csig} = \overline{\Cset} + \sigma B$, and moreover for any $\delta > 0$, $\nu(\overline{\Csig} + \delta B) = \nu(\Csig + \delta B)$ (as the boundary $\partial(\Csig + \delta B)$ is Lipschitz and therefore has measure zero). As a result, for each $t_{i,k}, i = 1, \ldots,k$ we may apply Lemma \ref{lem: expansion_volume} to $\overline{\Cset}$ and obtain
	\begin{equation}
	\label{eqn: slice_volume_bound}
	\nu_{i,k} = \nu(\Csig + rt_{i,k}B) \leq \nu(\Csig)\left(1 + \frac{rt_{i,k}}{\sigma}\right)^d
	\end{equation}
	which in turn gives
	\begin{align}
	\Sigma_k & \leq c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ rt_{i,k}}{\sigma}\right)^d \biggl( (t_{i,k})^{\gamma} - (t_{i-1,k})^{\gamma}\biggr) \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}).~~~~~~~~~~~~~~ (\text{substituting}~u_{i,k} := t_{i,k}^{\gamma}) \label{eqn: Sigmak_riemann_sum}
	\end{align}
	The expression in~\eqref{eqn: Sigmak_riemann_sum} is a Riemann sum, and taking the limit as $k \to \infty$ we obtain
	\begin{align}
	\lim_{k \to \infty} c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}) & = c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{r u^{1/\gamma}}{\sigma}\right)^{d} du \nonumber \\
	& \overset{\text{(i)}}{\leq} c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{2 d r u^{1/\gamma}}{\sigma}\right) du \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \left(1 + \gamma \frac{2 d r}{(\gamma + 1)\sigma}\right). \label{eqn: Sigmak_integral}
	\end{align}
	where $\text{(i)}$ follows from the upper bound in Lemma \ref{lem: Taylor_series} in light of the fact $r \leq \sigma / 2d$.

	An upper bound on $\xi$ follows from largely the same logic, although it does not involve integration:
	\begin{align}
	\xi & \overset{\text{(ii)}}{\leq} \nu(\Csig) \biggl\{ \left(1 + \frac{ r}{\sigma}\right)^d(\lambda_{\sigma} - c_0r^{\gamma}) - \lambda_{\sigma} \biggr\} \nonumber \\
	& \overset{\text{(iii)}}{\leq} \nu(\Csig) \biggl\{ \left(1 + \frac{2d r}{\sigma}\right)(\lambda_{\sigma} - c_0r^{\gamma}) - \lambda_{\sigma} \biggr\} = \nu(\Csig) \biggl\{ \frac{2dr}{\sigma}(\lambda_{\sigma} - c_0r^{\gamma}) - c_0 r^{\gamma} \biggr\}. \label{eqn: xi_ub}
	\end{align}
	where $\text{(ii)}$ follows from \eqref{eqn: slice_volume_bound}, and $\text{(iii)}$ from Lemma \ref{lem: Taylor_series}. As the bounds in \eqref{eqn: partition_ub} and \eqref{eqn: telescoping_sum} hold for all $k$, these along with \eqref{eqn: Sigmak_integral} and \eqref{eqn: xi_ub} imply the desired result.

%\subsection{Proof of Lemma~\ref{lem: expected_number_boundary_points}}

\subsection{Proof of Lemma~\ref{lem: local_conductance}}
The proof of this result relies on estimates of the volume of spherical caps which we defer to Appendix~\ref{sec:caps}.
	Since $u \in \Csig$ there exists $x \in \Cset$ such that $u \in B(x, \sigma)$, and as $B(x,\sigma) \subseteq \Csig$,
	\begin{equation*}
	\nu\bigl(B(u, r) \cap B(x, \sigma)\bigr) \leq \nu\bigl(B(u, r) \cap \Csig \bigr)
	\end{equation*}
	Without loss of generality, let $\norm{u - x} = \sigma$; it is not hard to see that if $\norm{u - x} < \sigma$, the volume of the overlap will only grow. Then, since $\norm{u  - x} = \sigma$, $B(u, r) \cap B(x, \sigma)$ contains a spherical cap of radius $r$ and height
	\begin{equation*}
	h = r - r^2/2\sigma = r \left( 1 - \frac{r}{2 \sigma} \right)
	\end{equation*}	
	which by Lemma \ref{lem: volume_of_spherical_cap} has volume
	\begin{equation*}
	\nu_{\text{cap}} = \frac{1}{2} \nu_d r^d I_{1 - \alpha}\left( \frac{d + 1}{2}  ,\frac{1}{2}\right)
	\end{equation*}
	with $\alpha = 1 - \frac{2rh - h^2}{r^2} = \frac{r^2}{4 \sigma^2} \leq \frac{1}{16d}$. 
	By Lemmas \ref{lem: beta_integral} (applied with $t = 1$) and \ref{lem: beta_function},
	\begin{align*}
	I_{1 - \alpha}\left( \frac{d + 1}{2}  ,\frac{1}{2}\right) & \geq 1 - \frac{\Gamma\bigl(\frac{d}{2}+ 1\bigr)}{\Gamma\bigl(\frac{d + 1}{2}\bigr) \Gamma\bigl(\frac{1}{2}\bigr)} \frac{3}{4\sqrt{d}} \\
	& \geq 1 - \frac{3}{4}\sqrt{\frac{d+2}{\pi d}} \geq 1 - \frac{3}{4}\sqrt{\frac{3}{2 \pi}}.
	\end{align*}

\subsection{Volume Estimates}
\label{sec: volume_estimates}
%\section{Proofs}
%
%Subsections \ref{sec: volume_estimates} - \ref{sec: proof_of_theorem_1} detail the proof for Theorem \ref{thm: conductance_upper_bound}. In subsections \ref{sec: mixing_time_on_graphs} and \ref{sec: proof_of_proposition_tv_mixing_time}, we establish some results regarding mixing time over general graphs. The proof of Theorem \ref{thm: mixing_time_upper_bound} is then provided in subsections \ref{sec: population_conductance_function} - \ref{sec: proof_of_theorem_mixing_time}. Section~\ref{sec: concentration} gives some general concentration results we use throughout, before we finish with the proofs of Theorems \ref{thm: misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters}, and the statement and proof of our results regarding the a\pprspace vector (Corollary \ref{cor: appr}), in subsections \ref{sec: proof_of_misclassification_rate} - \ref{sec: linalg}.

%\subsection{Volume estimates}

We begin by recalling some notation. We let $\Aset \subseteq \Reals^d$, and for $\sigma \geq 0$, write $\sigma B := B(0,\sigma) = \set{x \in \Rd: \norm{x} \leq \sigma}$ for the closed ball of radius $\sigma$ centered at the origin (and let $B^{\circ}(0,\sigma)$ denote the corresponding open ball). Let $\Asig = \Aset + \sigma B$ be the direct sum of $\Aset$ and $\sigma B$, $\Asig = \set{z = x + y: x \in \Aset, y \in \sigma B}$. Recall that we use $\nu$ for Lebesgue measure, and $\nu_d = \nu(B)$ for $B = (0,1)$.

Lemma \ref{lem: expansion_volume} provides a bound on the ratio $\nu(\Csig + r B) / \nu(\Csig)$, an important intermediate quantity in bounding the ratio $\cut(\Csig[\Xbf]; G_{n,r})/\vol(\Csig[\Xbf]; G_{n,r})$.

\begin{lemma}
	\label{lem: expansion_volume}
	If $\Aset$ is closed and bounded, then for any $\delta > 0$,
	\begin{equation}
	\label{eqn: expansion_volume}
	\nu(\Asig + \delta B) \leq \left(1 + \frac{\delta}{\sigma}\right)^d \nu(\Asig).
	\end{equation}
\end{lemma}
\begin{proof}
	We will show that for any $\epsilon > 0$, 
	\begin{equation}
	\label{eqn: ratio_of_volume}
	\frac{\nu(\Asig + \delta B)}{\nu(\Asig)} \leq \frac{(\sigma + \delta + \epsilon)^d}{\sigma^d}
	\end{equation}
	Taking the limit as $\epsilon \to 0$ results in \eqref{eqn: expansion_volume}.

	Fix $\epsilon > 0$. Our first goal is to find a finite collection $x_1, \ldots, x_N \in \Rd$ (where $N$ is a finite number that may implicitly depend on $\epsilon$) such that
	\begin{equation*}
	\bigcup_{i = 1}^{N} B(x_i, \sigma) \subseteq \Asig \subset \bigcup_{i = 1}^{N} B(x_i, \sigma + \epsilon).
	\end{equation*}
	Note that $\Asig$ is the direct sum of two compact sets, and is therefore itself compact. Moreover, for any $\epsilon > 0$,
	\begin{equation*}
	\Asig \subset \bigcup_{x \in \Aset} B^{\circ}(x,\sigma + \epsilon)
	\end{equation*}
	so by compactness there exists a finite subcover $x_1, \ldots,x_N \in \Aset$ such that
	\begin{equation}
	\label{eqn: finite_subcover}
	\Asig \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon).
	\end{equation}
	As a direct consequence of \eqref{eqn: finite_subcover}, $\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon + \delta)$, and by definition for every $x_i \in \Aset$, $B(x_i,\sigma) \in \Asig$. Summarizing our findings, we have
	\begin{equation}
	\label{eqn: finite_subcover-1}
	\bigcup_{i = 1}^{N} B(x_i,\sigma) \subseteq \Asig  ,~\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \delta + \epsilon).
	\end{equation}
\noindent	We next show a lower bound on $\nu(\Asig)$. Partition $\Asig$ using the balls $B(x_i,\sigma)$, meaning let $\Aset_{\sigma}^{(1)} := B(x_1,\sigma)$, $\Aset_{\sigma}^{(2)} := B(x_2,\sigma) \setminus B(x_1,\sigma)$, and so on, so that
	\begin{equation*}
	\Aset_{\sigma}^{\text{(i)}} := B(x_i,\sigma) \setminus \bigcup_{j = 1}^{i - 1} \Aset_{\sigma}^{(j)}. \tag{$i = 1,\ldots,N$}
	\end{equation*}
	Observe that $\bigcup_{i = 1}^{N} \Asig^{\text{(i)}} = \bigcup_{i = 1}^{N} B(x_i,\sigma)$, so by \eqref{eqn: finite_subcover} $\Asig \supseteq \bigcup_{i = 1}^{N} \Asig^{\text{(i)}}$. As $\Asig^{(1)},\ldots, \Asig^{(N)}$ are non-overlapping,
	\begin{align*}
	\nu(\Asig) & \geq \sum_{i = 1}^{N} \nu(\Asig^{\text{(i)}}) \\
	& = \sigma^d \nu_d \sum_{i = 1}^{N}  \frac{\nu(\Asig^{\text{(i)}})}{\nu(B(x_i,\sigma))}
	\end{align*}
	We turn to proving an analogous upper bound on $\nu(\Asig + \delta B)$. Let $\Aset_{\sigma + \epsilon + \delta}^{(1)} := B(x_1,\sigma + \delta + \epsilon)$ and
	\begin{equation*}
	\Aset_{\sigma + \delta + \epsilon}^{\text{(i)}} := B(x_i,\sigma + \delta + \epsilon) \setminus \bigcup_{j = 1}^{i - 1} \Aset_{\sigma + \delta + \epsilon}^{(j)}. \tag{$i = 2,\ldots,N$}
	\end{equation*}
As $\bigcup_{i = 1}^{N} \Aset_{\sigma + \delta + \epsilon}^{\text{(i)}} = \bigcup_{i = 1}^{N} B(x_i,\sigma + \delta + \epsilon)$, by \eqref{eqn: finite_subcover}
	\begin{equation*}
	\Aset_{\sigma} + \delta B \subset \bigcup_{i =1}^{N} \Aset_{\sigma + \delta + \epsilon}^{\text{(i)}}
	\end{equation*}
	and therefore 
	\begin{align*}
	\nu(\Aset_\sigma + \delta B) & \leq \sum_{i = 1}^{N} \nu\bigl(\Aset_{\sigma + \delta + \epsilon}^{\text{(i)}}\bigr) \\
	& = \sum_{i = 1}^{N} \nu_d (\sigma + \delta + \epsilon)^d \frac{\nu(\Aset_{\sigma + \delta + \epsilon}^{\text{(i)}})}{\nu(B(x_i, \sigma + \delta + \epsilon))} \\
	& \leq \nu_d (\sigma + \delta + \epsilon)^d \sum_{i = 1}^{N} \frac{\nu(\Asig^{\text{(i)}})}{\nu(B(x_i,\sigma))}
	\end{align*}
	where the last inequality follows from Lemma \ref{lem: covering}. We have shown \eqref{eqn: ratio_of_volume}, and thus the claim.
\end{proof}
\noindent We require Lemma \ref{lem: covering} to prove Lemma \ref{lem: expansion_volume}.
\begin{lemma}
	\label{lem: covering}
	For $i = 1, \ldots, N$ and  $\Aset_{\sigma}^{\text{(i)}}, \Aset_{\sigma + \delta + \epsilon}^{\text{(i)}}$ as in Lemma \ref{lem: expansion_volume},
	\begin{equation*}
	\frac{\nu(\Aset_{\sigma + \delta + \epsilon}^{\text{(i)}})}{\nu(B(x_i, \sigma + \delta + \epsilon))} \leq \frac{\nu(\Aset_{\sigma}^{\text{(i)}})}{\nu(B(x_i, \sigma))}
	\end{equation*}
\end{lemma}
\begin{proof}
	Let $\delta' := \delta + \epsilon$. It will be sufficient to show that
	\begin{equation*}
	\biggl(\Aset_{\sigma + \delta'}^{\text{(i)}} - \set{x_i}\biggr) \subseteq \left(1 + \frac{\delta'}{\sigma}\right)\cdot\biggl(\Asig^{\text{(i)}} - \set{x_i}\biggr) 
	\end{equation*}
	since then
	\begin{equation*}
	\nu(\Aset_{\sigma + \delta'}^{\text{(i)}}) \leq \left(1 + \frac{\delta'}{\sigma}\right)^d \nu(\Aset_{\sigma}^{\text{(i)}}) = \frac{\nu(B(x_i, \sigma + \delta'))}{\nu(B(x_i, \sigma))} \nu(\Aset_{\sigma}^{\text{(i)}}).
	\end{equation*}

	Assume without loss of generality that $x_i = 0$, and let $x \in \Aset_{\sigma + \delta'}^{\text{(i)}}$, meaning
	\begin{equation}
	\norm{x} \leq \sigma + \delta',~ \norm{x - x_j} > \sigma + \delta'~ \textrm{for $j = 1, \ldots, i - 1$}.
	\end{equation}
	Letting $x' = \frac{\sigma}{\sigma + \delta'} x$, since $\norm{x} \leq \sigma + \delta'$, $\norm{x'} \leq \sigma$ and therefore $x' \in B(0,\sigma)$. Additionally observe that for any $j = 1, \ldots, i - 1$, by the triangle inequality
	\begin{equation*}
	\norm{x' - x_j} \geq \norm{x - x_j} - \norm{x - x'} > \sigma + \delta' - \frac{\delta'}{\sigma + \delta'}\norm{x} \geq \sigma
	\end{equation*}
	and therefore $x' \not\in B(x_j,\sigma)$ for any $j = 1,\ldots, i - 1$. So $x' \in \Asig^{\text{(i)}}$.
\end{proof}

We will need to carefully control the volume of expansion sets using the estimate in Lemma~\ref{lem: expansion_volume}; Lemma~\ref{lem: Taylor_series} serves this purpose (see also, Lemma~23 in \cite{balakrishnan2013}). 
\begin{lemma}
	\label{lem: Taylor_series}
	For any $0 \leq x \leq 1/2d$,
	\begin{align*}
	(1 + x)^d & \leq 1 + 2\dx \\
	(1 - x)^d & \geq 1 - 2\dx.
	\end{align*}
\end{lemma}
\begin{proof}
	We take the binomial expansion of $(1 + x)^d$:
	\begin{align*}
	(1 + x)^d & = \sum_{k = 0}^{d} {d \choose k} x^k \\
	& = 1 + \dx + \dx\left(\sum_{k = 2}^d \frac{{d \choose k} x^{k - 1}}{d}\right)\\
	& \leq 1 + \dx + \dx\left(\sum_{k = 2}^d \frac{{d \choose k}}{(2d)^{k-1}d}\right) \tag{since $x \leq \frac{1}{2d}$} \\
	& \leq 1 + \dx + \dx\left(\sum_{k = 2}^d \frac{1}{2^{k - 1}}\right) \leq 1 + 2\dx.
	\end{align*}
The proof for the corresponding lower bound on $(1 - x)^d$ is symmetric.
\end{proof}

\subsection{Spherical Caps and Associated Estimates}
\label{sec:caps}
In this section, we state a result for the volume of a spherical cap and derive some 
useful upper bounds. 
\begin{lemma}
	\label{lem: volume_of_spherical_cap}
	Let $\mathrm{cap}_r(h)$ denote a spherical cap of radius $r$ and height $h$. Then, 
	\begin{equation*}
	\nu\bigl( \mathrm{cap}_r(h)  \bigr) = \frac{1}{2} \nu_d r^d I_{1 - \alpha}\left(\frac{d + 1}{2}; \frac{1}{2}\right)
	\end{equation*}
	where
	\begin{equation*}
	\alpha := 1 - \frac{2 r h - h^2}{r^2}
	\end{equation*}
	and
	\begin{equation*}
	I_{1 - \alpha}(z,w) = \frac{\Gamma(z + w)}{\Gamma(z) \Gamma(w)} \int_{0}^{1 - \alpha} u^{z - 1} (1 - u)^{w - 1} du.
	\end{equation*}
	is the cumulative distribution function of a $\mathrm{Beta}(z,w)$ distribution, evaluated at $1 - \alpha$. 
\end{lemma}

The following result provides a lower bound on the Beta integral, and the result in Lemma~\ref{lem: beta_function} provides
an upper bound on the ratio of Gamma functions. 
\begin{lemma}
	\label{lem: beta_integral}
	For any $0 \leq t \leq 1$ and $\alpha \leq \frac{t^2}{16 d}$,
	\begin{equation*}
	\int_{0}^{1 - \alpha}u^{(d-1)/2}(1 - u)^{-1/2}du \geq \frac{\Gamma\bigl(\frac{d + 1}{2}\bigr)\Gamma\bigl(\frac{1}{2}\bigr)}{ \Gamma\bigl(\frac{d}{2}+ 1\bigr)} - \frac{3t}{4\sqrt{d}}
	\end{equation*}
\end{lemma}
\begin{proof}
	We can write 
	\begin{equation*}
	\int_{0}^{1 - \alpha}u^{(d-1)/2}(1 - u)^{-1/2}du = \int_{0}^{1}u^{(d-1)/2}(1 - u)^{-1/2}du - \int_{1 - \alpha}^{1}u^{(d-1)/2}(1 - u)^{-1/2}du
	\end{equation*}
	The first integral is simply the beta function, with
	\begin{equation*}
	B(\frac{d+1}{2},\frac{1}{2}) := \frac{\Gamma\bigl(\frac{d + 1}{2}\bigr)\Gamma\bigl(\frac{1}{2}\bigr)}{ \Gamma\bigl(\frac{d}{2}+ 1\bigr)}.
	\end{equation*}
	To upper bound the second integral, we expand $(1 - u)^{-1/2}$ around $u = 1 - \alpha$, obtaining
	\begin{equation*}
	(1 - u)^{-1/2} \leq \alpha^{-1/2} + \max_{u \in (1 - \alpha, 1)} \frac{\alpha}{2} (1 - u)^{-3/2} = \frac{3}{2}\alpha^{-1/2}.
	\end{equation*}
	As a result,
	\begin{align*}
	\int_{1 - \alpha}^{1}u^{(d-1)/2}(1 - u)^{-1/2}du & \leq \frac{3}{2}\alpha^{-1/2} \int_{1 - \alpha}^{1}u^{(d-1)/2}du \\
	& = \frac{3}{d+1}\alpha^{-1/2} \left(1 - (1 - \alpha)^{(d + 1)/2}\right) \\
	& \overset{\text{(iv)}}{\leq} \frac{3}{(d+1)}\alpha^{-1/2} (\alpha(d + 1)) \\
	& =  3\alpha^{1/2}.
	\end{align*}
	where $\text{(iv)}$ follows from Lemma \ref{lem: Taylor_series}, and the fact $\alpha \leq \frac{1}{16d}$. The result follows from the condition $\alpha \leq \frac{t^2}{16d}$. 
\end{proof}

\begin{lemma}
	\label{lem: beta_function}
	\begin{equation*}
	\frac{\Gamma\bigl(\frac{d}{2}+ 1\bigr)}{\Gamma\bigl(\frac{d + 1}{2}\bigr) \Gamma\bigl(\frac{1}{2}\bigr)} \leq \sqrt{\frac{d + 2}{2\pi}}
	\end{equation*}
\end{lemma}
\noindent The proof of Lemma \ref{lem: beta_function} is straightforward and follows from the fact that $\Gamma(1/2) = \sqrt{\pi}$ and the upper bound $\Gamma(x + 1)/ \Gamma(x+s) \leq (x + 1)^{1-s}$ for $s \in [0,1]$.

%\subsection{Density-weighted cut and volume estimates}
%\label{sec: density_weighted_cut_and_volume_estimates}
%
%For notational ease, we write
%\begin{align*}
%\cut_{n,r} = \cut(\Csig[\Xbf]; G_{n,r}), ~ \mu_K = \mathbb{E}(\cut_{n,r}), ~ p_K = \frac{\mu_K}{{n \choose 2}} \\
%\vol_{n,r} = \vol(\Csig[\Xbf]; G_{n,r}), ~ \mu_V = \mathbb{E}(\vol_{n,r}), ~ p_V = \frac{\mu_V}{{n \choose 2}} \\
%\vol_{n,r}^c = \vol(\Xbf \setminus \Csig[\Xbf]; G_{n,r}), ~ \mu_V^c = \mathbb{E}(\vol_{n,r}^c), ~ p_V^c = \frac{\mu_V^c}{{n \choose 2}}
%\end{align*}
%for the random variable, mean, and probability of cut size and volume, respectively. 
%
%\begin{lemma}
%	\label{lem: expected_density_cut}
%	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
%	\begin{equation*}
%	p_K \leq \frac{4 d \nu_d r^{d+1} \lambda}{\sigma} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig)
%	\end{equation*}
%\end{lemma}
%\begin{proof}
%
%\end{proof}
%
%\begin{lemma}
%	\label{lem: expected_density_volume}
%	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
%	\begin{equation*}
%	p_V \geq \frac{12}{25} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig)
%	\end{equation*}
%\end{lemma}
%\begin{proof}
%	The proof will proceed similarly to Lemma \ref{lem: expected_density_cut}. We begin by writing $\vol_{n,r}$ as the sum of indicator functions,
%	\begin{equation}
%	\label{eqn: volume_expansion}
%	\vol_{n,r} = \sum_{i = 1}^{n} \sum_{j \neq i} \1(x_i \in \Csig) \1(x_j \in B(x_i, r))
%	\end{equation}
%	and by linearity of expectation we obtain
%	\begin{equation*}
%	p_V = \frac{\mu_V}{{n \choose 2}} = 2 \cdot \Pbb(x_i \in \Csig, x_j \in B(x_i,r)). \tag{for any $i,j$, $i \neq j$. }
%	\end{equation*}
%	Writing this with respect to the density function $f$, we have
%	\begin{align*}
%	p_V & = 2 \int_{\Csig} f(x) \Pbb(B(x,r)) \dx \\
%	& \geq 2 \int_{\Csig} f(x) \Pbb(B(x,r) \cap \Csig) \dx
%	\end{align*}
%	whence the claim then follows by Lemma \ref{lem: local_conductance}.
%\end{proof}
%
%To employ Lemmas \ref{lem: expected_density_cut} and \ref{lem: expected_density_volume} in the proof of Theorem \ref{thm: conductance_upper_bound}, we must relate the random variable
%\begin{equation*}
%\Phi_{n,r}(\Csig[\Xbf]) = \frac{\cut_{n,r}}{\min \set{\vol_{n,r}, \vol_{n,r}^c}}
%\end{equation*}
%to $p_K$ and $p_V$. 
%
%In Lemma \ref{lem: prob_bound_cutvol}, we give probabilistic bounds on the $\cut_{n,r}$, $\vol_{n,r}$ and $\vol_{n,r}^c$ in terms of $p_K$ and $p_V$. These bounds are a straightforward consequence of Lemma \ref{lem: bounded_difference}, Hoeffding's inequality for U-statistics.
%
%\begin{lemma}
%	\label{lem: prob_bound_cutvol}
%	For any $\delta \in (0,1]$,
%	\begin{equation}
%	\label{eqn: numerator_additive_bound}
%	\frac{\cut_{n,r}}{{n \choose 2}} \leq p_K + \sqrt{\frac{\log(1/\delta)}{n}},~\text{and}~ \frac{\vol_{n,r}}{{n \choose 2}}, \frac{\vol_{n,r}^c}{{n \choose 2}}  \geq p_V - \sqrt{\frac{\log(1/\delta)}{n}}.
%	\end{equation}
%	each with probability at least $1 - \delta$. 
%\end{lemma}
%
%\begin{proof}[Proof of Lemma \ref{lem: prob_bound_cutvol}]
%	From \eqref{eqn: density_cut_expansion} and \eqref{eqn: volume_expansion}, we see that $\cut_{n,r}$ and $\vol_{n,r}$, properly scaled, can be expressed as order-$2$ $U$-statistics,
%	\begin{equation*}
%	\frac{\cut_{n,r}}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_K(x_i, x_j),~~ \frac{\vol_{n,r}}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_V(x_i, x_j)
%	\end{equation*}
%	with kernels
%	\begin{align*}
%	\phi_K(x_i,x_j) & = \1(x_i \in \Csig, x_j \not\in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \in \Csig, x_i \not\in \Csig, \norm{x_i - x_j} \leq r) \\
%	\phi_V(x_i,x_j) & = \1(x_i \in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \in \Csig, \norm{x_i - x_j} \leq r). 
%	\end{align*}
%	
%	Similarly,
%	\begin{equation*}
%	\frac{\vol_{n,r}^c}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_{V^c}(x_i, x_j)
%	\end{equation*}
%	with kernel,
%	\begin{equation*}
%	\phi_{V^c}(x_i,x_j) = \1(x_i \not\in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \not\in \Csig, \norm{x_i - x_j} \leq r). 
%	\end{equation*}
%	
%	From Lemma \ref{lem: bounded_difference} we therefore have
%	\begin{equation*}
%	\frac{\cut_{n,r}}{{n \choose 2}} \leq p_K + \sqrt{\frac{\log(1/\delta)}{n}},~  \frac{\vol_{n,r}}{{n \choose 2}} \geq p_V - \sqrt{\frac{\log(1/\delta)}{n}}, ~ \frac{\vol_{n,r}^c}{{n \choose 2}} \geq p_V^c - \sqrt{\frac{\log(1/\delta)}{n}}
%	\end{equation*}
%	each with probability at least $1 - \delta$. The claim follows in light of \ref{asmp: bounded_volume}, which implies $p_V^c \geq p_V$. 
%\end{proof}
%
%\subsection{Proof of Theorem \ref{thm: conductance_upper_bound}}
%\label{sec: proof_of_theorem_1}
%
%The proof of Theorem \ref{thm: conductance_upper_bound} is more or less given by Lemmas \ref{lem: expected_density_cut}, \ref{lem: expected_density_volume}, and \ref{lem: prob_bound_cutvol}. All that remains is some algebra, which we take care of below.
%
%Fix $\delta \in (0,1]$ and let $\delta' = \delta/3$. We rewrite
%\begin{equation}
%\label{eqn: conductance_representation_1}
%\Phi_{n,r}(\Csig[\Xbf]) = \frac{p_K + \left(\frac{\cut_{n,r}}{{n \choose 2}} - p_K\right)}{p_V + \left(\frac{\min\set{\vol_{n,r}, \vol_{n,r}^c}}{{n \choose 2}} - p_V\right)}.
%\end{equation}
%Assume (\ref{eqn: numerator_additive_bound}) holds with respect to $\delta'$, keeping in mind that this will happen with probability at least $1 - \delta$. Along with (\ref{eqn: conductance_representation_1}) this means
%\begin{equation*}
%\Phi_{n,r}(\Csig[\mathbf{X}]) \leq \frac{p_K + \Err_n}{p_V - \Err_n}
%\end{equation*}
%for $\Err_n = \sqrt{\frac{\log(1/\delta')}{n}}$.
%Now, some straightforward algebraic manipulations yield
%\begin{align*}
%\frac{p_K + \Err_n}{p_V - \Err_n} & = \frac{p_K}{p_V} \left(\frac{p_V}{p_V - \Err_n}\right) + \frac{\Err_n}{p_V - \Err_n} \\
%& = \frac{p_K}{p_V} + \left(\frac{p_K}{p_V} + 1\right)\frac{\Err_n}{p_V - \Err_n} \\
%& \leq \frac{p_K}{p_V} + 2 \frac{\Err_n}{p_V - \Err_n}.
%\end{align*}
%By Lemmas \ref{lem: expected_density_cut} and \ref{lem: expected_density_volume}, we have
%\begin{equation*}
%\frac{p_K}{p_V} \leq \frac{100rd}{12\sigma} \frac{\lambda}{\lambda_{\sigma}} \frac{\left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right)}{\lambda_{\sigma}}.
%\end{equation*}
%Then, by the choice of sample size in \eqref{eqn: conductance_sample_complexity}, 
%\begin{equation*}
%n \geq \frac{(2 + \epsilon)^2 \log\left(\frac{3}{\delta}\right)}{\epsilon^2 p_V^2}
%\end{equation*}
%which implies $2 \frac{\Err_n}{p_V - \Err_n} \leq \epsilon$.

\section{Proof of Theorem \ref{thm: mixing_time_upper_bound}}

Let $\widetilde{G}_{n,r} := G_{n,r}[\Csig[\Xbf]]$; in general, we will use tilde notation to refer to quantities computed over $\Csig$ or over the induced subgraph $\widetilde{G}_{n,r}$. The goal of Theorem \ref{thm: mixing_time_upper_bound} is to show that, for universal constants $c_2, c_3,c_4 > 0$, with probability one
\begin{equation*}
\limsup_{n \to \infty} \tau_{\infty}(\widetilde{G}_{n,r}) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{c_4 \Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + c_3
\end{equation*}
We follow a two-step approach typically used to establish upper bounds on the mixing time of Markov chains.

In the first step, we relate the mixing time $\tau_{\infty}(G)$ of an arbitrary undirected graph $G = (V,E)$ to expansion properties of subsets $U \subseteq V$. We now build to a formal definition of these expansion properties. First, we recall the the \emph{cut} and \emph{volume} functionals over a graph, and introduce the \emph{degree} functional as well. For $u \in V$, $S \subseteq V$, and $S^c = V \setminus S$,
\begin{equation*}
\cut(S;G) = \sum_{u \in S} \sum_{v \in S^c} \1((u,v) \in E), \quad \deg(u;G) = \sum_{v \in V} \1((u,v) \in E), \quad \vol(S;G) = \sum_{u \in S} \deg(u;G).
\end{equation*}
Additionally, we recall the \emph{normalized cut} $\Phi(S;G)$, defined (as in \eqref{eqn: normalized_cut}) as
\begin{equation*}
\Phi(S;G) = \frac{\cut(S; G)}{\min\set{\vol(S; G),\vol(S^c; G)}}.
\end{equation*}

With these functionals in hand, we can now formally define the expansion parameters \emph{local spread} $s(G)$ and \emph{conductance} $\Phi(G)$,
\begin{equation*}
s(G) := \frac{9}{10} \cdot \min_{u \in V} \set{\deg(u; G)} \cdot \min_{u \in V} \set{\pi(v)}, \quad \Phi(G) := \min_{S \subseteq V} \Phi(S;G).
\end{equation*}

The following Proposition establishes an upper bound on the mixing time $\tau_{\infty}(G)$ in terms of the local spread $s(G)$ and conductance $\Phi(G)$.
\begin{proposition}
	\label{prop: pointwise_mixing_time}
	Assume $\min_{u \in V} \deg(u; G) \geq 10$. Then,
	\begin{equation*}
	\tau_{\infty}(G) \leq \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
	\end{equation*}
\end{proposition}

The second step in our approach is to lower bound the local spread and conductance over the neighborhood graph $\widetilde{G}_{n,r}$. In the following result we give lower bounds for both quantities, which hold asymptotically as $n \to \infty$.
\begin{proposition}
	\label{prop: local_spread_conductance}
	Under the setup and conditions of Theorem \ref{thm: mixing_time_upper_bound}, with probability one,
	\begin{equation}
	\label{eqn: local_spread}
	\liminf_{n \to \infty} s(\widetilde{G}_{n,r}) \geq \frac{\lambda_{\sigma}^2 \nu_d r^d}{20\Lambda_{\sigma}} 
	\end{equation}
	and
	\begin{equation}
	\label{eqn: conductance}
	\liminf_{n \to \infty} \Phi(\widetilde{G}_{n,r}) \geq \frac{\lambda_{\sigma}^2 r}{\Lambda_{\sigma}^2 2^{13} \rho L \sqrt{d}}.
	\end{equation}
\end{proposition}

Taking these results as given, the proof of Theorem \ref{thm: mixing_time_upper_bound} is more or less complete. We note that since $r$ is fixed, with probability one $\liminf_{n \to \infty} \min_{u \in \Csig[\Xbf]} \deg(u; \widetilde{G}_{n,r}) = \infty$ (see Lemma \ref{lem: ball_bounds_in_probability}). In particular, for sufficiently large $n$ the condition $\min_{u \in \Csig[\Xbf]} \deg(u; \widetilde{G}[\Xbf]) \geq 10$ is satisfied, again with probability one. We may therefore apply Proposition \ref{prop: pointwise_mixing_time} and obtain
\begin{align*}
\liminf_{n \to \infty} \tau_{\infty}(\widetilde{G}_{n,r}) & \leq \liminf_{n \to \infty} \left\{\frac{2}{\Phi^2(\widetilde{G}_{n,r})} \log \left(\frac{1440}{s(\widetilde{G}_{n,r})}\right)\log \left(\frac{14}{s(\widetilde{G}_{n,r})}\right)  + 3 \log \left(\frac{14}{s(\widetilde{G}_{n,r})}\right) + 3 \right\}
\\
& \leq \frac{2^{27}\Lambda_{\sigma}^4\rho^2L^2 d^3}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{28800\Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + 3 \log^2\left(\frac{280\Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + 3
\end{align*}
whereupon the theorem follows after an appropriate choice of universal constants $c_2, c_3, c_4 > 0$. In the rest of this section we proceed to proving Propositions \ref{prop: pointwise_mixing_time} and \ref{prop: local_spread_conductance}.

\subsection{Proof of Proposition \ref{prop: pointwise_mixing_time}}
\label{sec: pointwise_mixing_time}

We first generalize some notation from the main text. Let $A$ be the adjacency matrix of an undirected graph $G = (V,E)$, and $D$ the associated diagonal degree matrix. The lazy random walk over $G$ is the Markov chain with transition probabilities given by $\Wbf := \frac{\Ibf{} + \Dbf^{-1} \Abf}{2}$ and stationary distribution $\pi$ with elements $\pi_u = D_{uu}/\vol(V;G)$.  Denote the $t$-step probability distribution of this random walk, originating from a  vertex $v \in V$, as $q^{(t)}: V \times V \to [0,1]$, $q^{(t)}(v,u) = e_v \Wbf^t e_u$.

Consider the uniform distance\footnote{Note $d_{\textrm{unif}}$ is not formally a distance as it is not symmetric.} between the distributions $q_v^{(t)} := q^{(t)}(v,\cdot)$ and $\pi$, given by
\begin{equation*}
d_{\textrm{unif}}(q_v^{(t)},\pi) = \max_{u \in V} \set{\frac{\pi(u) - q_v^{(t)}(u)}{\pi(u)}}.
\end{equation*}
Our goal is to demonstrate that for a sufficiently large $t$, $d_{\textrm{unif}}(q_v^{(t)},\pi) \leq \frac{1}{4}$ for every $v \in V$ (see~\eqref{eqn: mixing_time} to recall the definition of $\tau_{\infty}(G)$). To do so, we introduce the \emph{total variation distance} between the distributions $q_v^{(t)}$ and $\pi$, given by
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} = \sum_{u \in V} \abs{q_v^{(t)}(u) - \pi(u)}.
\end{equation*}
As we will see, an upper bound on the uniform distance can be obtained from an analogous upper bound on the total variation distance. First, however, we upper bound the total variation distance $||q_v^{(t)} - \pi||_{\mathrm{TV}}$ as a function of the local spread $s(G)$, the conductance $\Phi(G)$ and the number of steps $t$.
\begin{lemma}
	\label{lem: tv_mixing_time}
	For any $v \in V$, and any $0 < a \leq 1/4$,
	\begin{align}
	\norm{q_v^{(t + 3)} - \pi}_{\mathrm{TV}} \leq \max\set{as(G), \frac{1}{8} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V} \deg(u;G)}} 
	\nonumber \\
	~~~~~~~~~~~~~~~~~~~~~ + \left(\frac{1}{1 - 2as(G)} +  \frac{1}{2 as(G)}\right) \left(1 - \frac{\Phi^2(G)}{2}\right)^t \label{eqn: tv_mixing_time}
	\end{align}
\end{lemma}

As mentioned, a bound on the total variation distance implies a corresponding bound on the uniform distance $d_{\mathrm{unif}}$, given by the following result:
\begin{lemma}
	\label{lem: tv_to_uniform_distance}
	Let $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{s(G)}{14}$. Then,
	\begin{equation*}
	d_{\mathrm{unif}}(q_v^{(t + 3)},\pi) \leq \frac{1}{4}
	\end{equation*}
\end{lemma}

\noindent Taking these Lemmas as given, we proceed that show that for
\begin{equation*}
\tau_1 = \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
\end{equation*}
the uniform distance $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$, which proves the claim of Proposition \ref{prop: pointwise_mixing_time}. Fix $a = \frac{1}{18}$, and let $\tau_0 = \frac{2}{\Phi^2(G)} \log \left(\frac{80}{a s(G)}\right)$, so that
\begin{equation*}
\left(1 - \frac{\Phi^2(G)}{2}\right)^{\tau_0} \leq \exp(- \tau_0\Phi^2(G)/2) \leq \frac{as(G)}{80}.
\end{equation*}
Recall that by assumption, $\min_{u \in V} \deg(u;G) \geq 10$. Moreover note that for $a \leq 1/4$, $2as(G) < 1/2 < 1 - 2as(G)$, since $s(G) \leq 9/10 < 1$. By Lemma \ref{lem: tv_mixing_time} we therefore obtain
\begin{equation*}
\norm{q_v^{(\tau_0 + 3)} - \pi}_{\mathrm{TV}} \leq \max \set{\frac{1}{20}, \frac{1}{8} + \frac{1}{20} + \frac{1}{20}} + \left(\frac{1}{as(G)}\right) \frac{as(G)}{80} \leq \frac{1}{4}.
\end{equation*}
It is a well known fact (see e.g.~\cite{montenegro2002}) that if $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{1}{4}$, then for any $0 < \epsilon < 1$, $\norm{q_v^{(t\log_2(1/\epsilon))} - \pi}_{\mathrm{TV}} \leq \epsilon$. Therefore, letting $\tau_1 = (\tau_0 + 3) \log(\frac{14}{s(G)})$, 
\begin{equation*}
\norm{q_v^{(\tau_1)} - \pi}_{\mathrm{TV}} \leq \frac{s(G)}{14}
\end{equation*}
and so by Lemma \ref{lem: tv_to_uniform_distance}, $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$. The proof of Proposition \ref{prop: pointwise_mixing_time} is therefore complete once we have proved Lemmas \ref{lem: tv_mixing_time} and \ref{lem: tv_to_uniform_distance}.

\subsection{Proof of Lemma \ref{lem: tv_mixing_time}}
\label{sec: mixing_time_on_graphs}

We generalize the notation of the previous subsection. For a starting distribution $z$ to be specified later, and for $t \geq 0$ an integer, let $q_z^{(t)}$ be the $t$-step probability distribution of the lazy random walk with starting distribution $z$.\footnote{We say $z$ is a starting distribution over a graph $G$ when $\textrm{supp}(z) \subseteq V$ and $\sum_{u \in V}z(u) = 1$. Then, $q_z^{(t)} = zW^t$.}

We will find it useful to introduce the \emph{Lovasz-Simonovits curve}, defined for any $t \in \mathbb{N}$ and starting distribution $z$ to be $h_z^{(t)}: [0,1] \to [0,1]$,
\begin{equation*}
h_z^{(t)}(x) = \max_{w \in \mathcal{W}_x} \left\{ \sum_{u \in V} \left(q_z^{(t)}(u) - \pi(u)\right)w(u)\right\}.
\end{equation*}
The maximum in the preceding display is over the set of weight functions $\mathcal{W}_{x}$
\begin{align*}
\mathcal{W}_x = \left\{w: V \to [0,1] \Big\vert 0 \leq w(u) \leq 1~~\forall~u,~~\text{and}~~\sum_{u \in V} w(u) \pi(u) = x\right\}.
\end{align*}
The utility of the Lovasz-Simonovits curve is demonstrated by the following observations. First, note that
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} = \sup_{S \subseteq V} \abs{q_v^{(t)}(S) - \pi(S)}
\end{equation*}
where we use the standard notation $\pi(S) := \sum_{u \in S} \pi(u)$, and likewise for $q_v^{(t)}(S)$. Moreover, observe that for any $S \subseteq V$ and any integer $t \geq 0$,
\begin{equation}
\label{eqn: lovasz_simonovits_1}
\abs{q_v^{(t)}(S) - \pi(S)} \leq \max \set{h_v^{(t)}(\pi(S)), h_v^{(t)}(1 - \pi(S))}
\end{equation}
(To see this, consider the weight functions $w(u) = \1(u \in S)$ and $w'(u) = 1 - w(u)$.) Taking the maximum on both sides of \eqref{eqn: lovasz_simonovits_1}  over all $S \subseteq V$, we have
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \max_{0 \leq x \leq 1} h_v^{(t)}(x).
\end{equation*}
Now, take $z = e_vW^3$, and trivially observe that $h_v^{(t+3)}(x) = h_z^{(t)}(x)$. To prove Lemma \ref{lem: tv_mixing_time}, it is therefore sufficient to show the desired upper bound -- the right hand side of~\eqref{eqn: tv_mixing_time}-- holds with respect to $h_z^{(t)}(x)$ for all $x \in [0,1]$, and all starting distributions $z = e_vW^3$.

Let $\mu = as(G)$, and note under the condition $a < 1/4$, $\mu < 1 - \mu$. To show the desired upper bound, we split the interval $[0,1]$ into the subinterval $[\mu, 1 - \mu]$ and the remainder $[0,\mu) \cup (1 - \mu, 1]$. Let $\ell_{\mu}(x)$ be the linear interpolator between $h_z^{(0)}(\mu)$ and $h_z^{(0)}(1 - \mu)$, 
\begin{equation*}
\ell_{\mu}(x) = \frac{1 - \mu - x}{1 - 2\mu} h_z^{(0)}(\mu) + \frac{x - \mu}{1 - 2\mu}h_z^{(0)}(1 - \mu).
\end{equation*}

The following technical Lemma collects the upper bounds which hold over $[\mu, 1 - \mu]$ and $[0,\mu) \cup (1 - \mu, 1]$, respectively:
\begin{lemma}
	\label{lem: mixing_all_sets}
	For any $\mu \leq x \leq 1 - \mu$ and any starting distribution $z = e_vW^3$,
	\begin{equation}
	\label{eqn: mixing_large_sets}
	h_z^{(t)}(x) \leq \ell_{\mu}(x) + \left(\frac{1}{1 - 2\mu} + \frac{1}{\mu}\right)\left(1 - \frac{\Phi^2(G)}{2}\right)^t.
	\end{equation}
	For any $0 \leq x < \mu$ or $1 - \mu < x \leq 1$
	\begin{equation}
	\label{eqn: mixing_small_sets}
	h_z^{(t)}(x) \leq \max\set{as(G), \frac{1}{2^{t+3}} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} }
	\end{equation}
\end{lemma}
\noindent Taking Lemma \ref{lem: mixing_all_sets} as given, we have nearly completed our proof of Lemma \ref{lem: tv_mixing_time}. Note that for any $\mu \leq x \leq 1 - \mu$,
\begin{align*}
\ell_{\mu}(x) & \leq \max\set{h_z^{(t)}(\mu),h_z^{(t)}(1 - \mu)} \\
& \leq \max\set{as(G), \frac{1}{2^{t + 3}} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} }
\end{align*}
with the latter inequality following from \eqref{eqn: mixing_small_sets}. Therefore, for any $x \in [0,1]$,
\begin{equation*}
h_z^{(t)}(x) \leq \max\set{as(G), \frac{1}{8} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} } + \left(\frac{1}{1 - 2as(G)} + \frac{1}{as(G)}\right)\left(1 - \frac{\Phi^2(G)}{2}\right)^t
\end{equation*}
which is exactly the claimed result of Lemma \ref{lem: tv_mixing_time}.

\subsection{Proof of Lemma \ref{lem: mixing_all_sets}}

We first prove \eqref{eqn: mixing_large_sets}, and then \eqref{eqn: mixing_small_sets}.
\paragraph{Proof of \eqref{eqn: mixing_large_sets}:}
The desired result is a consequence of Theorem 1.2 of \cite{lovasz1990}. To state this theorem, we introduce the notation
\begin{equation*}
C_{\mu} = \max \set{\frac{h_z^{(0)}(x) - \ell_{\mu}(x)}{\sqrt{x - \mu}}, \frac{h_z^{(0)}(x) - \ell_{\mu}(x)}{\sqrt{1 - x - \mu}}: \mu < x < 1 - \mu},
\end{equation*}
and then the theorem itself.
\begin{theorem}[Theorem 1.2 of \cite{lovasz1990}]
	\label{thm: lovasz_simonovits_1993}
	For any $\mu \leq x \leq 1 - \mu$, $z$ an arbitrary starting distribution, and an integer $t \geq 0$,
	\begin{equation}
	\label{eqn: lovasz_simonovits_1993}
	h_z^{(t)}(x) \leq \ell_{\mu}(x) + C_{\mu} \min \set{\sqrt{x - \mu}, \sqrt{1 - x - \mu}} \left(1 - \frac{\Phi^2(G)}{2}\right)^t
	\end{equation}
\end{theorem}
In order to show \eqref{eqn: mixing_large_sets}, we must therefore provide an appropriate bound on the quantity $C_{\mu}$. Precisely, we will show that for any $\mu < x < 1 - \mu$,
\begin{equation}
\label{eqn: lt_ub_1}
h_z^{(0)}(x) - \ell_{\mu}(x) \leq \max \set{\frac{h_z^{(0)}(\mu)}{1 - 2\mu} + \frac{h_z^{(0)}(\mu)}{\mu} , \frac{h_z^{(0)}(1 - \mu) }{1 - 2\mu} + 1 } \min \set{\sqrt{x - \mu},\sqrt{1 - x - \mu}}
\end{equation}
which will in turn imply
\begin{align*}
C_{\mu} & \leq \max \set{\frac{h_z^{(0)}(\mu)}{1 - 2\mu} + \frac{h_z^{(0)}(\mu)}{\mu} , \frac{h_z^{(0)}(1 - \mu) }{1 - 2\mu} + 1 } \\
& \leq \frac{1}{\mu} + \frac{1}{1 - 2\mu}.
\end{align*}
Plugging this upper bound into \eqref{eqn: lovasz_simonovits_1993}, we obtain
\begin{align*}
h_z^{(t)}(x) & \leq \ell_{\mu}(x) +  \left(\frac{1}{\mu} + \frac{1}{1 - 2\mu}\right)\min \set{\sqrt{x - \mu}, \sqrt{1 - x - \mu}} \left(1 - \frac{\Phi^2(G)}{2}\right)^t \\
& \leq \ell_{\mu}(x) +  \left(\frac{1}{\mu} + \frac{1}{1 - 2\mu}\right)\left(1 - \frac{\Phi^2(G)}{2}\right)^t
\end{align*}
which is the desired result.

It remains to show \eqref{eqn: lt_ub_1}. To do so, we make use of an equivalent representation of the Lovasz-Simonovits curve $h_z^{(t)}$. Order the elements of $V = \set{u_1, \ldots, u_N}$, where $N = \abs{V}$, such that
\begin{equation}
\label{eqn: lovasz_simonovits_curve_1}
\frac{q_z^{(t)}(u_1)}{\pi(u_1)} \geq \frac{q_z^{(t)}(u_2)}{\pi(u_2)} \geq \ldots \geq \frac{q_z^{(t)}(u_N)}{\pi(u_N)},
\end{equation}
and for each $k = 1,\ldots, N$, let $U_k = \set{u_1, \ldots, u_k}$. Then for any $x \in [0,1]$, letting $k$ satisfy $\pi(U_{k - 1}) < x < \pi(U_k)$, it can be shown that\footnote{See \cite{lovasz1990} for a formal justification.},
\begin{equation}
\label{eqn: lovasz_simonovits_curve}
h_z^{(t)}(x) = \sum_{j = 1}^{k - 1} (q_z^{(t)}(u_j) - \pi(u_j)) + \frac{x - \pi(U_{k - 1})}{\pi(u_k)} \left(q_z^{(t)}(u_k) - \pi(u_k) \right).
\end{equation}
The representation of the Lovasz-Simonovits curve given by~\eqref{eqn: lovasz_simonovits_curve} makes it clear that $h_z^{(t)}(x)$ is a piecewise linear curve, with knots at the points $x_k = \pi(U_k)$ for $k = 1,\ldots,N$, where the $k$th linear segment has slope $v_k = q_z^{(t)}(u_{k})/\pi(u_{k}) - 1$. By the ordering of~\eqref{eqn: lovasz_simonovits_curve_1}, we have that $v_1(x) > v_2(x) > \ldots > v_{N-1}(x)$, and the curve is therefore concave.

In fact, letting $v(x) = \sum_{k = 1}^{N} v_k \1(x \in [U_k, U_{k+1}) )$ be the slope of the Lovasz-Simonovits curve, for any $x \geq \mu$ it can be shown that
\begin{align}
v(x) & \leq \min_{k: u_k \leq x} v_k \nonumber  \\
& \leq \frac{h_z^{(0)}(\mu)}{\mu}. \label{eqn: lovasz_simonovits_slope}
\end{align}
From the upper bound in~\eqref{eqn: lovasz_simonovits_slope} along with the concavity of $h_z^{(0)}(x)$, we have that for any $x \geq \mu$,
\begin{equation*}
h_z^{(0)}(x) \leq h_z^{(0)}(\mu) + (x - \mu)\frac{h_z^{(0)}(\mu)}{\mu}.
\end{equation*}
Some algebra then yields that for any $x \geq \mu$,
\begin{align*}
h_z^{(0)}(x) - \ell_{\mu}(x) & \leq h_z^{(0)}(\mu) - \left(\frac{1 - \mu - x}{1 - 2\mu} h_z^{(0)}(\mu) + \frac{x - \mu}{1 - 2\mu} h_z^{(0)}(1 - \mu)\right) + \frac{h_z^{(0)}(\mu)}{\mu}(x - \mu) \\
& =  \frac{x - \mu}{1 - 2\mu}h_z^{(0)}(\mu) + \frac{h_z^{(0)}(\mu)}{\mu} (x - \mu) - \frac{x - \mu}{1 - 2\mu} h_z^{(0)}(1 - \mu) \\
& \leq \sqrt{x - \mu} \left(\frac{h_z^{(0)}(\mu)}{1 - 2\mu} + \frac{h_z^{(0)}(\mu)}{\mu} \right)
\end{align*}
On the other hand, $\ell_{\mu}(1 - \mu) = h_z^{(0)}(1 - \mu)$, and by the concavity of $h_z^{(0)}$ and \eqref{eqn: lovasz_simonovits_slope},  for $x \leq 1 - \mu$
\begin{equation*}
h_z^{(0)}(x) \leq h_z^{(0)}(1 - \mu) + (1 - x - \mu).
\end{equation*}
Similar manipulations to above give the upper bound
\begin{align*}
h_z^{(0)}(x) - \ell_{\mu}(x) \leq \sqrt{1 - \mu - x}\left(\frac{h_z^{(0)}(1 - \mu) }{1 - 2\mu} + 1\right)
\end{align*}
and \eqref{eqn: lt_ub_1} follows.

\paragraph{Proof of \eqref{eqn: mixing_small_sets}:}
We let $z = e_vW^3$ for an arbitrary $v \in V$. First, we deal with the case $x \leq as(G)$. Letting $\pi(U_{k-1}) \leq x \leq \pi(U_{k})$, we have
\begin{equation}
\label{eqn: mixing_time_small_sets_1}
h_z^{(t)}(x) \leq q_z^{(t)}(U_{k - 1}) + q_z^{(t)}(u_k)
\end{equation}
We observe a few facts about the random walk defined by $q_z^{(t)}$. By the laziness of the random walk, for $u \neq v, t \geq 1$
\begin{equation}
\label{eqn: mixing_time_small_sets_2}
q_z^{(t)}(u) \leq \frac{1}{2 \min_{u \in V}\deg(u;G)}
\end{equation}
On the other hand if $u = v$,
\begin{equation}
\label{eqn: mixing_time_small_sets_3}
q_z^{(t)}(u) \leq \frac{1}{2^{t + 3}} + \frac{1}{2\min_{u \in V}\deg(u;G)}.
\end{equation}
Therefore by \eqref{eqn: mixing_time_small_sets_1}, \eqref{eqn: mixing_time_small_sets_2}, and \eqref{eqn: mixing_time_small_sets_3}
\begin{equation*}
h_z^{(t)}(x) \leq \frac{1}{2^{t + 3}} + \frac{\abs{U_{k - 1}}}{2 \min_{u \in V}\deg(u;G)} + \frac{1}{2 \min_{u \in V}\deg(u;G)}.
\end{equation*}
Finally, since $x \leq a s(G)$ and $x \geq \pi(U_{k - 1})$,
\begin{align*}
\abs{U_{k - 1}} & \leq \frac{\pi(U_{k-1})}{\min_{u \in V} \pi(v) } \\
& \leq \frac{as(G)}{\min_{u \in V} \pi(v) } \\
& \leq \frac{9a\min_{u \in V}\deg(u;G)}{10}.
\end{align*}
and the desired result is shown.

Now, we turn to the case where $x \geq 1 - as(G)$. Noting that the slope $v(x)$ of the Lovasz-Simonovits curve $h_z^{(t)}$ is bounded below by $-1$ for all $x \in [0,1]$, by the concavity of $h_z^{(t)}$ we have
\begin{align*}
h_z^{(t)}(x) & \leq h_z^{(t)}(1) + (1 - x) \\
& = 1 - x \leq as(G).
\end{align*}
and the proof of Lemma~\ref{lem: tv_mixing_time} is complete.

\subsection{Proof of Lemma \ref{lem: tv_to_uniform_distance}}

The proof of Lemma \ref{lem: tv_to_uniform_distance} follows from straightforward algebraic manipulation. Fix $u \in V$ and let $m \geq t + 1$ be arbitrary. The stationarity of $\pi$ gives (see~(16) of \citep{morris2005})
\begin{align}
\frac{\pi(u) - q_v^{m}(u)}{\pi(u)} & = \sum_{y \in V} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \nonumber \\
& \overset{\text{(i)}}{=} \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{\pi(u)} \left(\frac{1}{2} - \pi(u)\right) \nonumber \\
& \leq \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{2 \pi(u)} \label{eqn: tv_to_uniform_distance_1}
\end{align}
where $\text{(i)}$ follows from $q^{(1)}(u,u) = \frac{1}{2}$. Then,
\begin{align}
\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) & \leq \norm{q_v^{(m-1)} - \pi}_{\mathrm{TV}} \max_{y \neq u} \abs{\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}} \nonumber \\
& \leq \norm{q_v^{(m-1)} - \pi}_{\mathrm{TV}} \max \set{1, \max_{y \neq u}\set{\frac{q^{(1)}(y,u)}{\pi(u)}}} \nonumber \\
& \overset{(ii)}{\leq} \norm{q_v^{(m-1)} - \pi}_{\mathrm{TV}} \max \set{1, \frac{1}{2 \pi(u) \min_{u' \in V} \deg(u';G) }} \nonumber \\
& \leq \norm{q_v^{(m-1)} - \pi}_{\mathrm{TV}} \frac{1}{s(G)} \label{eqn: tv_to_uniform_distance_3}
\end{align}
where $(ii)$ follows from the fact that for $y \neq u$, $q^{(1)}(y,u) \leq 1/\left(2 \min_{u \in V} \deg(u; G)\right)$. As $m - 1 \geq t$, it is well known (see e.g. \cite{lovasz1990}) that the laziness of the random walk guarantees $\norm{q_v^{(m - 1)} - \pi}_{\mathrm{TV}} \leq \norm{q_v^{(t)} - \pi}_{\mathrm{TV}}$, and therefore by \eqref{eqn: tv_to_uniform_distance_3} and the hypothesis of Lemma \ref{lem: tv_to_uniform_distance},
\begin{equation*}
\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \leq \frac{1}{14}.
\end{equation*}
Plugging this in to \eqref{eqn: tv_to_uniform_distance_1} and taking maximum on both sides, we obtain
\begin{equation}
d_{\textrm{unif}}(q_v^{(m)}, \pi) \leq \frac{1}{14} + \frac{d_{\textrm{unif}}(q_v^{(m - 1)}, \pi)}{2} \label{eqn: tv_to_uniform_distance_2}
\end{equation}
The recurrence relation of \eqref{eqn: tv_to_uniform_distance_2} along with the initial condition $d_{\textrm{unif}}(q_v^{(t)}, \pi) \leq 1$ yields
\begin{equation*}
d_{\textrm{unif}}(q_v^{(t + 1)}, \pi) \leq \frac{8}{14} \Rightarrow d_{\textrm{unif}}(q_v^{(t + 2)}, \pi) \leq \frac{5}{14} \Rightarrow  d_{\textrm{unif}}(q_v^{(t + 3)}, \pi) \leq \frac{1}{4}
\end{equation*}
and the claim is shown. We have proved Lemmas \ref{lem: tv_mixing_time} and \ref{lem: tv_to_uniform_distance}, and therefore Proposition \ref{prop: pointwise_mixing_time}.

\subsection{Proof of Proposition \ref{prop: local_spread_conductance}}

We prove~\eqref{eqn: local_spread} immediately before turning our attention to~\eqref{eqn: conductance}, which will require the bulk of our attention. First, however, we introduce some notation. Recall that $\widetilde{G}_{n,r} = G_{n,r}\bigl[\Csig[\Xbf]\bigr]$ denotes the subgraph induced by $\Csig[\Xbf]$. For $S \subseteq \Csig[\Xbf]$ and $u \in \Csig[\Xbf]$, we will write
\begin{equation*}
\widetilde{\cut}_{n,r}(S) := \cut(S; \widetilde{G}_{n,r}), \quad \widetilde{\deg}_{n,r}(u) := \deg(u; \widetilde{G}_{n,r}), \quad 
\widetilde{\vol}_{n,r}(S) := \vol(S; \widetilde{G}_{n,r})
\end{equation*}
and let $\widetilde{\pi}_{n,r}(u) = \widetilde{\deg}_{n,r}(u)/\widetilde{\vol}_{n,r}(\Csig[\Xbf])$ be the stationary distribution of the lazy random walk over $\widetilde{G}_{n,r}$. We also let $\widetilde{\Phi}_{n,r}(S) := \Phi(S; \widetilde{G}_{n,r})$ denote the normalized cut functional over $\widetilde{G}_{n,r}$.

\paragraph{Proof of \eqref{eqn: local_spread}:}

Letting $\degminwt := \min_{u \in \Csig[\Xbf]} \widetilde{\deg}_{n,r}(u)$, we rewrite $s(\widetilde{G}_{n,r}) = \frac{9 \degminwt^2}{10\widetilde{\vol}_{n,r}(\Csig[\Xbf])}$. Then by Lemma \ref{lem: ball_bounds_in_probability}, with probability one
\begin{equation*}
\liminf_{n \to \infty} s(\widetilde{G}_{n,r}) \geq \frac{36 \lambda_{\sigma} r^d \nu_d}{625 \Pbb(\Csig) \Lambda_{\sigma}} \geq \frac{\lambda_{\sigma}^2 \nu_d r^d}{20\Lambda_{\sigma}}.
\end{equation*}

\paragraph{Proof of \eqref{eqn: conductance}:}

Our goal is to show that with probability one
\begin{equation*}
\liminf_{n \to \infty} \set{\min_{S \subseteq \Csig[\Xbf]} \widetilde{\Phi}_{n,r}(S)} \geq \frac{\lambda_{\sigma}^2 r}{\Lambda_{\sigma}^2 2^{13} \rho L \sqrt{d}}.
\end{equation*}
In order to show this bound holds uniformly over all sets $S \subseteq \Csig[\Xbf]$, we will split the analysis into two cases based on the size of $S \subseteq \Csig[\Xbf]$. To do so, we introduce $\mathcal{L}(G) := \set{S \subseteq V: \pi(S), \pi(S^c) \geq s(G)}$ (where as usual $\pi$ denotes the stationary distribution of a lazy random walk over $G = (V,E)$.)

Lemma \ref{lem: graph_conductance_profile_lb_1} shows that for any subset $S \subseteq \Csig[\Xbf]$ not in $\mathcal{L}(\widetilde{G}_{n,r})$, the graph normalized cut of $S$ is at least $1/10$. In fact, this statement holds for any graph $G$.
\begin{lemma}
	\label{lem: graph_conductance_profile_lb_1}
	Let $G = (V,E)$ be an arbitrary undirected graph. Then, for non-empty subsets $S \subseteq V$, 
	\begin{equation*}
	\min_{S \not\in \mathcal{L}(G)}\Phi(S; G) \geq \frac{1}{10}.
	\end{equation*}
\end{lemma}
\begin{proof}
	The claim follows by simple manipulations:
	\begin{align*}
	\Phi(S;G) & \geq \frac{\cut(S;G)}{\vol(S;G)} \\
	& \geq \sum_{u \in S} \frac{\deg(u;G) - \abs{S}}{\vol(S;G)} \\
	& \geq \sum_{u \in S} \frac{\deg(u;G) - \pi(S)/(\min_{u \in V}\pi(u))}{\vol(S;G)} \\
	& \geq \sum_{u \in S} \frac{\deg(u;G) - \frac{9}{10} \min_{u \in V}\deg(u;G)}{\vol(S;G)} \\
	& \geq \frac{1}{10} \sum_{u \in S} \frac{\deg(u;G)}{\vol(S;G)} = \frac{1}{10}.
	\end{align*}
\end{proof}

In Lemma \ref{lem: graph_conductance_profile_lb}, we establish a uniform lower bound on the normalized cut for the remaining sets $S \in \mathcal{L}(\widetilde{G}_{n,r})$.

\begin{lemma}
	\label{lem: graph_conductance_profile_lb}
	Under the setup and conditions of Theorem \ref{thm: mixing_time_upper_bound}, with probability one
	\begin{equation}
	\label{eqn: graph_conductance_profile_lb}
	\liminf_{n \to \infty} \left\{ \min_{S \in \mathcal{L}(\widetilde{G}_{n,r})} \widetilde{\Phi}_{n,r}(S) \right\} \geq \frac{\lambda_{\sigma}^2 r}{\Lambda_{\sigma}^2 2^{13} \rho L \sqrt{d}}.
	\end{equation}
\end{lemma}

The desired upper bound on graph conductance \eqref{eqn: conductance} follows from Lemma \ref{lem: graph_conductance_profile_lb}, along with Lemma \ref{lem: graph_conductance_profile_lb_1}, in light of the fact $\frac{\lambda_{\sigma}^2 r}{\Lambda_{\sigma}^2 2^{13} \rho L \sqrt{d}} < \frac{1}{10}$. We turn now to the proof of Lemma \ref{lem: graph_conductance_profile_lb}.

\subsection{Proof of Lemma \ref{lem: graph_conductance_profile_lb}.}

The proof of Lemma~\ref{lem: graph_conductance_profile_lb} will essentially follow from a pair of technical results. The first of these will demonstrate that the functional $\widetilde{\Phi}_{n,r}(S)$ can be lower bounded by a population analogue $\widetilde{\Phi}_{\Pbb,r}(\Sset)$, for an appropriately chosen $\Sset \subseteq \Csig$; we term this latter functional the continuous normalized cut. This lower bound will hold uniformly over sets $S \subseteq \Csig[X]$. The second technical Lemma will build on known continuous space isoperimetric inequalities to lower bound the continuous normalized cut $\widetilde{\Phi}_{\Pbb,r}(\Sset)$ uniformly over sets $\Sset \subseteq \Csig$.

We will now build slowly toward a formal definition of the continuous normalized cut, before establishing a relation between it and it's discrete counterpart. Let $\Sset \subseteq \Csig$ be measurable. We introduce the \emph{$r$-ball walk}, a Markov chain over $\Csig$ with transition probability at $x \in \Csig$ given by 
\begin{equation*}
\widetilde{P}_{\Pbb,r}(x; \Sset) := \frac{\Pbb(\Sset \cap B(x,r))}{\Pbb(\Csig \cap B(x,r))}.
\end{equation*}

Denote the stationary distribution for this Markov chain by $\piwt_{\Pbb,r}$, which is defined by the relation
\begin{equation*}
\int_{\Csig} \widetilde{P}_{\Pbb,r}(x; \Sset) \,d \piwt_{\Pbb,r}(x) = \piwt_{\Pbb,r}(\Sset).
\end{equation*}
Letting the \emph{$\Pbb$-local conductance} be given by
\begin{equation*}
\ell_{\Pbb,r}(x) := \Pbb\bigl(\Csig \cap B(x,r)\bigr)
\end{equation*}
a bit of algebra verifies that
\begin{equation*}
\piwt_{\Pbb,r}(\Sset) = \frac{\int_{\Sset} \ell_{\Pbb,r}(x) \,d \Pbb(x) }{\int_{\Csig} \ell_{\Pbb,r}(x) \,d \Pbb(x)}.
\end{equation*}

We next introduce the \emph{ergodic flow}, $\widetilde{Q}_{\Pbb,r}$. Let $\Sset \cap \Sset^c = \Csig$ be a partition of $\Csig$. Then the ergodic flow between $\Sset$ and $\Sset$ is given by 
\begin{equation*}
\widetilde{Q}_{\Pbb,r}(\Sset, \Sset^c) := \int_{\Sset} \widetilde{P}_{\Pbb,r}(x; \Sset^c) \,d\piwt_{\Pbb,r}(x), 
\end{equation*}
and the \emph{($\Pbb$) continuous normalized cut} by
\begin{equation*}
\widetilde{\Phi}_{\Pbb,r}(\Sset) := \frac{\widetilde{Q}_{\Pbb,r}(\Sset, \Sset^c)}{\min \set{\piwt_{\Pbb,r}(\Sset),\piwt_{\Pbb,r}(\Sset^c)}},
\end{equation*}

As the functionals $\widetilde{\Phi}_{n,r}$ and $\widetilde{\Phi}_{\Pbb,r}$ act in the different spaces $\Csig[\Xbf]$ and $\Csig$, respectively, it is not obvious how to relate them. To do so, following the lead of \cite{garciatrillos16}, we introduce transportation maps between the space $\Csig$ and the sample points $\Csig[\Xbf]$. We note that by assumption~\ref{asmp: bounded_density}, $\Pbb(\Csig) > 0$, and therefore with probability one as $n \to \infty$, the number of sample points $\abs{\Csig[\Xbf]}$ will be non-zero as well. We may therefore define the conditional probability measures
\begin{equation*}
\widetilde{\Pbb}(\Sset) = \frac{\Pbb(\Sset)}{\Pbb(\Csig)}, ~ \widetilde{\Pbb}_{n}(\Sset) := \frac{1}{\abs{\Csig[\Xbf]}} \sum_{x_i \in \Csig[\Xbf]} \1(x_i \in \Sset).
\end{equation*} 
We then define a \emph{transportation map} between $\widetilde{\Pbb}$ and $\widetilde{\Pbb}_n$ to be any measurable map $T: \Csig \to \Csig[\Xbf]$ such that for every $S \subseteq \Csig[\Xbf]$,
\begin{equation*}
\widetilde{\Pbb}(T^{-1}(S)) = \widetilde{\Pbb}_n(S),
\end{equation*}
where $T^{-1}(S) = \set{x \in \Csig: T(x) \in S}$ is the preimage of $T$. Observe that by the definition of the transportation map $T$, for any $g \in L^1\bigl(\widetilde{\Pbb}_n\bigr)$ the following change of variables formula holds
\begin{equation*}
\int_{\Csig} g(x) \,d\widetilde{\Pbb}_{n}(x) = \int_{\Csig} g\bigl(T(x)\bigr) \,d\widetilde{\Pbb}(x)
\end{equation*}
Using the change of variables formula with an appropriate choice of $g$, after suitable rescaling we can relate $\widetilde{\cut}_{n,r}(S)$ to $\widetilde{Q}_{\Pbb,r}(T^{-1}(S), T^{-1}(S)^c)$. Similarly, we can relate $\widetilde{\vol}_{n,r}(S^c)$ to $\widetilde{\pi}_{\Pbb,r}(T^{-1}(S))$. Working along these lines, we obtain the following lower bound on $\widetilde{\Phi}_{n,r}(S)$, stated in terms of the transportation distance $\norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)}$, where $\mathrm{Id}(x) = x$ is the identity mapping over $\Csig$.

\begin{lemma}
	\label{lem: graph_to_continuous_conductance}
	Let $T:\Csig \to \Csig[\Xbf]$ be a transportation map between $\widetilde{\Pbb}$ and $\widetilde{\Pbb}_n$. uppose $\norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)} < \min \set{s(\widetilde{G}_{n,r}),r, \lambda_{\sigma}/(2^{d+1}d\Lambda_{\sigma}r)}$. Then there exists a constant $c > 0$ which may depend on $\Pbb$, but does not depend on the sample size $n$, such that for all $S \in \mathcal{L}(\widetilde{G}_{n,r})$, 
	\begin{equation}
	\label{eqn: graph_to_continuous_conductance}
	\widetilde{\Phi}_{n,r}(S) \geq \widetilde{\Phi}_{\Pbb,r}(T^{-1}(S)) - \frac{c \norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)} }{s(\widetilde{G}_{n,r}) - c \norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)}  }
	\end{equation}
\end{lemma}

Clearly, Lemma \ref{lem: graph_to_continuous_conductance} is useful only when combined with an upper bound on the transportation distance $\norm{\mathrm{Id} - T_n}_{L^{\infty}(\Pbb)}$. Proposition 5 of \cite{garciatrillos16} establishes such an upper bound, with respect to transportation maps on measures supported on open, connected and bounded domains with Lipschitz boundaries. The following result is a restatement of this Proposition with respect to the domain $\Csig$, and the measure $\widetilde{\Pbb}$. Although $\Csig$ is closed rather than open, as $\nu(\partial \Csig) = 0$ we may apply Proposition 5 of \cite{garciatrillos16} to the interior $\Csig^o$ of $\Csig$, and the desired result will hold for any arbitrary extension of $T_n$ to $\Csig$. Let $\widetilde{n} = \abs{\Csig[\Xbf]}$.

\begin{theorem}[Restatement of Proposition 5 of \cite{garciatrillos16}]
	\label{thm: stagnating_transportation_maps}
	With probability one, there exists a sequence of transportation maps $\seq{T_n}$, $T_n: \Csig \to \Csig[\Xbf]$ such that the following statement holds:
	\begin{equation}
	\label{eqn: stagnating_transportation_maps}
	\limsup_{\widetilde{n} \to \infty} \frac{\widetilde{n}^{1/d} \norm{\mathrm{Id} - T_n}_{L^{\infty}(\widetilde{\Pbb})}}{(\log \widetilde{n})^{p_d}} \leq c
	\end{equation}
	where $c$ is constant in $n$ and $p_d = 3/4$ for $d = 2$ and $1/d$ for $d \geq 3$.
\end{theorem}
We remark that $\norm{\mathrm{Id} - T_n}_{L^{\infty}(\widetilde{\Pbb})} = \norm{\mathrm{Id} - T_n}_{L^{\infty}(\Pbb)}$, and therefore the preceding theorem shows that with probability one, $\limsup_{n \to \infty} \norm{\mathrm{Id} - T_n}_{L^{\infty}(\Pbb)} = 0$. The following result then supplies the last step, a uniform lower bound on the continuous normalized cut $\widetilde{\Phi}_{\Pbb,r}(\Sset)$ for all sets $\Sset \subseteq \Csig$. Let the \emph{$\Pbb$-continuous conductance} be given by
\begin{equation*}
\widetilde{\Phi}_{\Pbb,r} := \min_{\Sset \subseteq \Csig} \widetilde{\Phi}_{\Pbb,r}(\Sset).
\end{equation*} 
\begin{lemma}
	\label{lem: nonuniform_continuous_conductance}
	Under the setup and conditions of Theorem \ref{thm: mixing_time_upper_bound}, the $\Pbb$-continuous conductance of the $r$-ball walk satisfies
	\begin{equation*}
	\widetilde{\Phi}_{\Pbb,r} > \frac{\lambda_{\sigma}^2 r}{2^{13} \Lambda_{\sigma}^2 \rho L \sqrt{d}}.
	\end{equation*}
\end{lemma}

With Lemmas \ref{lem: graph_to_continuous_conductance} and \ref{lem: nonuniform_continuous_conductance}, as well as Theorem \ref{thm: stagnating_transportation_maps}, in hand, we proceed to the proof of Lemma \ref{lem: graph_conductance_profile_lb}. Observe that for any $c > 0$ fixed in $n$, for sufficiently large $n$ with probability one $\norm{\mathrm{Id} - T_n}_{L^{\infty}(\Pbb)} \leq c$. This fact, along with the asymptotic upper bound on $s(\widetilde{G}_{n,r})$ given by~\eqref{eqn: local_spread}, guarantees that the conditions of Lemma~\ref{lem: graph_to_continuous_conductance} are satisfied with probability one as $n \to \infty$. Therefore,
\begin{align*}
\liminf_{n \to \infty} \set{\min_{S \in \mathcal{L}(\widetilde{G}_{n,r})} \widetilde{\Phi}_{n,r}(S)} & \geq \liminf_{n \to \infty} \set{\min_{S \in \mathcal{L}(\widetilde{G}_{n,r})} \widetilde{\Phi}_{\Pbb,r}(\Sset) - \frac{c \norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)} }{s(\widetilde{G}_{n,r}) - c \norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)}  }} \\
& \geq \widetilde{\Phi}_{\Pbb,r} - \liminf_{n \to \infty} \set{\frac{c \norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)}}{s(\widetilde{G}_{n,r}) - c \norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)}}} \\
& = \widetilde{\Phi}_{\Pbb,r} \tag{Theorem \ref{thm: stagnating_transportation_maps},\eqref{eqn: local_spread}, and Slutsky's Theorem.}
\end{align*}
and the bound on $\widetilde{\Phi}_{\Pbb,r}$ given by Lemma \ref{lem: nonuniform_continuous_conductance} yields the desired result. It remains to prove Lemmas \ref{lem: graph_to_continuous_conductance} and \ref{lem: nonuniform_continuous_conductance}, which we do in the following two subsections.

\subsection{Proof of Lemma \ref{lem: graph_to_continuous_conductance}}
Let $S \subseteq \Csig[\Xbf]$ be an element of $\mathcal{L}(\widetilde{G}_{n,r})$, and denote $\Sset := T^{-1}(S)$. Denote $\varDelta = \norm{\mathrm{Id} - T}_{L^{\infty}(\Pbb)}$, and let $r^{+} = r + \varDelta$ and $r^{-} = r - \varDelta$. Our goal is to lower bound
\begin{equation*}
\frac{\widetilde{\cut}_{n,r}(S,S^c)}{\widetilde{\vol}_{n,r}(S)} \geq \widetilde{\Phi}_{\Pbb,r}(\Sset) - \frac{c \varDelta}{s(\widetilde{G}_{n,r}) - c \varDelta}
\end{equation*}
for some constant $c > 0$ which may depend on the distribution $\Pbb$ but not on the sample size $n$. We first state the relationships between the discrete functionals $\widetilde{\cut}_{n,r}$ and $\widetilde{\vol}_{n,r}$, and the continuous functionals $\widetilde{Q}_{\Pbb,r}$ and $\widetilde{\pi}_{\Pbb,r}$, alluded to in the previous section.
\begin{lemma}
	\label{lem: cut_volume}
	\begin{equation}
	\label{eqn: cut}
	\frac{1}{\widetilde{n}^2}\widetilde{\cut}_{n,r}(S) \geq \frac{\int_{\Csig} \ell_{\Pbb,r^-}(x) \,d\Pbb(x)}{\Pbb(\Csig)^2} \widetilde{Q}_{\Pbb,r^-}(\Sset, \Sset^c)
	\end{equation}
	and 
	\begin{equation}
	\label{eqn: volume}
	\frac{1}{\widetilde{n}^2}\widetilde{\vol}_{n,r}(S) \leq \frac{\int_{\Csig} \ell_{\Pbb,r^+}(x) \,d\Pbb(x)}{\Pbb(\Csig)^2} \widetilde{\pi}_{\Pbb,r^+}(\Sset)
	\end{equation}
\end{lemma}
To make use of Lemma \ref{lem: cut_volume}, we provide deviation inequalities on $\int_{\Csig} \ell_{\Pbb,r^+}(x) \,d\Pbb(x)$, $\int_{\Csig} \ell_{\Pbb,r^-}(x) \,d\Pbb(x)$, $\widetilde{Q}_{\Pbb,r^-}(\Sset, \Sset^c)$, and $\widetilde{\pi}_{\Pbb,r^+}(\Sset)$ in terms of the transportation distance $\varDelta$.
\begin{lemma}
	\label{lem: deviation_transportation_distance}
	Suppose $\Delta \leq r$. Then there exist constants $c_5,c_6 \geq 0$ which may depend on $\Pbb$ but do not depend on the sample size $n$, such that for every $\Sset \subseteq \Csig$,
	\begin{align*}
	\int_{\Csig} \ell_{\Pbb,r^+}(x) \,d\Pbb(x) & \leq \int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x) + c_5 \varDelta \\
	\int_{\Csig} \ell_{\Pbb,r^-}(x) \,d\Pbb(x) & \geq \int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x) - c_5 \varDelta \\
	\widetilde{Q}_{\Pbb,r^-}(\Sset, \Sset^c) & \geq \widetilde{Q}_{\Pbb,r}(\Sset, \Sset^c) - c_6 \varDelta \\
	\widetilde{\pi}_{\Pbb,r^+}(\Sset) & \leq \widetilde{\pi}_{\Pbb,r}(\Sset) + c_6 \varDelta.
	\end{align*}
\end{lemma}

The combination of Lemmas \ref{lem: cut_volume} and \ref{lem: deviation_transportation_distance} brings us close to our goal, as demonstrated by the following manipulations:
\begin{align}
\frac{\widetilde{\cut}_{n,r}(S,S^c)}{\widetilde{\vol}_{n,r}(S)} & \geq \frac{\widetilde{Q}_{\Pbb,r^-}(\Sset,\Sset^c)}{\widetilde{\pi}_{\Pbb,r^+}(\Sset)} \frac{\int_{\Csig} \ell_{\Pbb,r^-}(x) \,d\Pbb(x)}{\int_{\Csig} \ell_{\Pbb,r^+}(x) \,d\Pbb(x)} \tag{Lemma \ref{lem: cut_volume}} \\
& \geq \frac{\widetilde{Q}_{\Pbb,r}(\Sset,\Sset^c) - c_6 \varDelta }{\widetilde{\pi}_{\Pbb,r^+}(\Sset) + c_6 \varDelta} \left(1 - \frac{2 c_5 \varDelta}{\int_{\Csig}\ell_{\Pbb,r}(x)\,d\Pbb(x)}\right) \tag{Lemma \ref{lem: deviation_transportation_distance} } \\
& \geq \frac{\widetilde{Q}_{\Pbb,r}(\Sset,\Sset^c)}{\widetilde{\pi}_{\Pbb,r}(\Sset)} - \left(\frac{2c_6}{\widetilde{\pi}_{\Pbb,r}(\Sset)} + \frac{2c_5}{\int_{\Csig}\ell_{\Pbb,r}(x)\,d\Pbb(x)}\right)\varDelta \label{eqn: proof_of_lemma_graph_to_continuous_1}
\end{align}
where the last line follows from some basic algebra. The following result relates the unknown stationary distribution $\widetilde{\pi}_{\Pbb,r}(\Sset)$ to $\widetilde{\pi}_{n,r}(S)$.
\begin{lemma}
	\label{lem: continuous_pi_lb}
	Suppose $\varDelta \leq \lambda_{\sigma}/(2^{d+1}d\Lambda_{\sigma}r)$. Then there exists a constant $c_7 > 0$ which may depend on the distribution $\Pbb$ but does not depend on the sample size $n$ such that for every $S \subseteq \Csig[\Xbf]$,
	\begin{equation*}
	\widetilde{\pi}_{\Pbb,r}(\Sset) \geq \widetilde{\pi}_{n,r}(S) - c_7 \varDelta
	\end{equation*}
\end{lemma}
Combining this result with~\eqref{eqn: proof_of_lemma_graph_to_continuous_1}, we obtain
\begin{align*}
\frac{\widetilde{\cut}_{n,r}(S,S^c)}{\widetilde{\vol}_{n,r}(S)} & \geq \widetilde{\Phi}_{\Pbb,r}(\Sset) - \left(\frac{2c_6}{\widetilde{\pi}_{n,r}(S) - c_7\varDelta} + \frac{2c_5}{\int_{\Csig}\ell_{\Pbb,r}(x)\,d\Pbb(x)}\right)\varDelta \\
& \geq \widetilde{\Phi}_{\Pbb,r}(\Sset) - \left(\frac{2c_6}{s(\widetilde{G}_{n,r}) - c_7\varDelta} + \frac{2c_5}{\int_{\Csig}\ell_{\Pbb,r}(x)\,d\Pbb(x)}\right)\varDelta
\end{align*}
where the latter inequality follows as we assumed $S \in \mathcal{L}(\widetilde{G}_{n,r})$. Choosing the constant $c$ in Lemma \ref{lem: graph_to_continuous_conductance} to be $c = \max\set{2c_5/\bigl(\int_{\Csig}\ell_{\Pbb,r}(x)\,d\Pbb(x) \bigr) + 2c_6, c_7}f$, we achieve our desired result. It remains to show Lemmas~\ref{lem: cut_volume}, \ref{lem: deviation_transportation_distance}, and \ref{lem: continuous_pi_lb}, which we now turn to.

\paragraph{Proof of Lemma~\ref{lem: cut_volume}: }

We begin with $\widetilde{\cut}_{n,r}(S)$.
\begin{align*}
\frac{1}{\widetilde{n}^2} \widetilde{\cut}_{n,r}(S) & = \frac{1}{ \widetilde{n}^2} \sum_{x_i, x_j \in \Csig[\Xbf]} \1(\norm{x_i - x_j} \leq r) \1(x_i \in S) \1(x_j \not\in S)\\
& = \iint_{\Csig \times \Csig} \1(\norm{x - y} \leq r) \1(x \in S) \1(y \not\in S) \,d\widetilde{\Pbb}_n(x) \,d\widetilde{\Pbb}_n(y) \\
& =  \iint_{\Csig \times \Csig} \1(\norm{T(x) - T(y)} \leq r) \1(T(x) \in S) \1(T(y) \not\in S)\,d\widetilde{\Pbb}(x) \,d\widetilde{\Pbb}(y) \tag{change of variables}  \\
& \geq \iint_{\Csig \times \Csig} \1(\norm{x - y} \leq r^-) \1(T(x) \in S) \1(T(y) \not\in S) \,d\widetilde{\Pbb}(x) \,d\widetilde{\Pbb}(y) \\
& = \int_{\Sset} \int_{\Sset^c \cap B(x,r^-)} 1 \,d\widetilde{\Pbb}(y) \,d\widetilde{\Pbb}(x)
\end{align*}
By definition we have $\frac{\,d\Pbb(x)}{\,d\widetilde{\Pbb}(x)} = \Pbb(\Csig)$. Therefore,
\begin{align}
\int_{\Sset} \int_{\Sset^c \cap B(x,r^-)} d\widetilde{\Pbb}(y) d\widetilde{\Pbb}(x) & = \frac{1}{\Pbb(\Csig)^2} \int_{\Sset} \int_{\Sset^c \cap B(x,r^-)} d\Pbb(y) d\Pbb(x) \nonumber \\
& = \frac{\int_{\Csig} \ell_{\Pbb,r^-}(x) d\Pbb(x)}{\Pbb(\Csig)^2} \int_{\Sset} \frac{\Pbb\bigl(\Sset^c \cap B(x,r^-)\bigr)}{\ell_{\Pbb,r^-}(x)} d\piwt_{\Pbb,r^-}(x) \nonumber \\
& = \frac{\int_{\Csig} \ell_{\Pbb,r^-}\,d\Pbb(x)}{\Pbb(\Csig)^2} \widetilde{Q}_{\Pbb,r^-}(\Sset, \Sset^c). \label{eqn: cut_bound}
\end{align}	
We obtain an upper bound on $\widetilde{\vol}_{n,r}(S)$ by similar manipulations, as follows:
\begin{align}
\frac{1}{\widetilde{n}^2} \widetilde{\vol}_{n,r}(S) & \leq \frac{1}{ \widetilde{n}^2} \sum_{x_i, x_j \in \Csig[\Xbf]} \1(\norm{x_i - x_j} \leq r) \1(x_i \in S) \nonumber \\
& = \iint_{\Csig \times \Csig} \1(\norm{x - y} \leq r) \1(x \in S) \,d\widetilde{\Pbb}_n(x) \,d\widetilde{\Pbb}_n(y) \nonumber \\
& = \iint_{\Csig \times \Csig} \1(\norm{T(x) - T(y)} \leq r) \1(T(x) \in S) \,d\widetilde{\Pbb}(x) \,d\widetilde{\Pbb}(y) \nonumber \\
& \leq \iint_{\Csig \times \Csig} \1(\norm{x - y} \leq r^+) \1(x \in S) \,d\widetilde{\Pbb}(x) \,d\widetilde{\Pbb}(y) \nonumber \\
& = \int_{\Sset} \int_{\Csig \cap B(x,r^+)} 1 \,d\widetilde{\Pbb}(y) \,d\widetilde{\Pbb}(x) \nonumber \\
& = \frac{1}{\Pbb(\Csig)^2}\int_{\Sset} \int_{\Csig \cap B(x,r^+)} 1 \,d\Pbb(y) \,d\Pbb(x) \nonumber \\
& = \frac{1}{\Pbb(\Csig)^2}\int_{\Sset} \ell_{\Pbb,r^+}(x) \,d\Pbb(x) \nonumber \\
& = \frac{\int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x)}{\Pbb(\Csig)^2} \piwt_{\Pbb,r^+}(\Sset). \label{eqn: vol_bound}
\end{align}

\paragraph{Proof of Lemma~\ref{lem: deviation_transportation_distance}: }

Consider the set
\begin{equation*}
\mathcal{R}(x) := \set{y \in \Csig: y \in B(x,r^+), y \not\in B(x,r^-)}
\end{equation*} 
for $x \in \Csig$, and observe that
\begin{equation*}
\int_{\mathcal{R}(x)} \,d\Pbb(y) \leq \Lambda_{\sigma} \nu_d \left((r + \delta)^d - (r - \varDelta)^d\right)
\end{equation*}
and therefore
\begin{equation}
\label{eqn: deviation_transportation_distance_2}
\int_{\Csig} \int_{\mathcal{R}(x)} \,d\Pbb(y) \,d\Pbb(x) \leq \Pbb(\Csig )\Lambda_{\sigma} \nu_d \left((r + \varDelta)^d - (r - \varDelta)^d\right).
\end{equation}
A first-order Taylor expansion of $(r + \varDelta)^d$ results in the upper bound $(r + \varDelta)^d \leq r^d + d\varDelta(r + \varDelta)^{d - 1}$, and similarly $(r - \varDelta)^d \geq r^d - d\varDelta(r + \varDelta)^{d - 1}$. Plugging these bounds into~\eqref{eqn: deviation_transportation_distance_2} yields
\begin{align}
\int_{\Csig} \int_{\mathcal{R}(x)} \,d\Pbb(y) \,d\Pbb(x)  & \leq \Pbb(\Csig) \Lambda_{\sigma} \nu_d \bigl( 2d(r + \Delta)^{d - 1}\varDelta \bigr) \tag{1st-order Taylor expansion of $(r + \Delta)^d$ } \\
& \leq \Pbb(\Csig) \Lambda_{\sigma} \nu_d  2^{d}d r^{d-1}\varDelta =: c_5 \varDelta. \label{eqn: deviation_transportation_distance_1}
\end{align}
where the second line follows from the condition $\Delta \leq r$. Using~\eqref{eqn: deviation_transportation_distance_1}, we proceed to obtain each of the four bounds stated in Lemma~\ref{lem: deviation_transportation_distance}. First, as $\ell_{\Pbb,r^{+}}(x) \leq 1$ for all $x \in \Csig$, we have
\begin{align*}
\int_{\Csig} \ell_{\Pbb,r^+}(x) \,d\Pbb(x) & \leq \int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x) + \int_{\Csig} \int_{\mathcal{R}(x)} \,d\Pbb(y) \,d\Pbb(x) \\
& \leq \int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x) + c_5\varDelta.
\end{align*}
An equivalent bound for $\int_{\Csig} \ell_{\Pbb,r^-}(x) \,d\Pbb(x)$ is obtained by similar reasoning. We now lower bound $\widetilde{Q}_{\Pbb,r^{-}}(\Sset, \Sset^c)$,
\begin{align*}
\widetilde{Q}_{\Pbb,r^{-}}(\Sset, \Sset^c) & = \int_{\Sset} \widetilde{P}_{\Pbb,r^-}(x; \Sset) \,d\widetilde{\pi}_{\Pbb,r}(x) \\
& = \frac{\int_{\Sset} \Pbb(\Sset^c \cap B(x,r^-)) d\Pbb(x)}{\int_{\Csig} \Pbb(\Csig \cap B(x,r^-)) d\Pbb(x)} \\
& \geq \frac{\int_{\Sset} \Pbb(\Sset^c \cap B(x,r)) d\Pbb(x) - \int_{\Csig} \int_{\mathcal{R}(x)} \,d\Pbb(y) \,d\Pbb(x)}{\int_{\Csig} \Pbb(\Csig \cap B(x,r)) d\Pbb(x)} \\
& \geq \widetilde{Q}_{\Pbb,r}(\Sset, \Sset^c) - \frac{c_5 \varDelta}{\int_{\Csig}\ell_{\Pbb,r}(x) \,d\Pbb(x)} =:\widetilde{Q}_{\Pbb,r}(\Sset, \Sset^c) - c_6 \varDelta.
\end{align*}
Finally, we upper bound $\widetilde{\pi}_{\Pbb,r^{+}}(\Sset)$,
\begin{align*}
\widetilde{\pi}_{\Pbb,r^{+}}(\Sset) & = \frac{\int_{\Sset} \ell_{\Pbb,r^+}(x) \,d\Pbb(x)}{\int_{\Csig} \ell_{\Pbb,r^+}(x) \,d\Pbb(x)} \\
& \leq \frac{\int_{\Sset} \ell_{\Pbb,r}(x) \,d\Pbb(x) + \int_{\Sset} \int_{\mathcal{R}(x)} \,d\Pbb(y) \,d\Pbb(x)}{\int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x)} \\
& \leq \widetilde{\pi}_{\Pbb,r}(\Sset) + c_6\varDelta.
\end{align*}

\paragraph{Proof of Lemma~\ref{lem: continuous_pi_lb}: }
The proof of Lemma~\ref{lem: continuous_pi_lb} will not be too different from the proof of Lemma~\ref{lem: deviation_transportation_distance}. From the change of variables formula, we have
\begin{align*}
\widetilde{\pi}_{n,r}(S) & = \frac{\int_{\Sset}\int_{\Csig} \1(\norm{x - y} \leq r) \,d\widetilde{\Pbb}_n(y) \,d\widetilde{\Pbb}_n(x)}{\int_{\Csig}\int_{\Csig} \1(\norm{x - y} \leq r) \,d\widetilde{\Pbb}_n(y) \,d\widetilde{\Pbb}_n(x)} \\
& \leq \frac{\int_{\Sset} \ell_{\Pbb,r^+}(x) \,d\Pbb(x) }{\int_{\Csig}\ell_{\Pbb,r^-}(x) \,d\Pbb(x)} \\
& \leq \frac{\int_{\Sset} \ell_{\Pbb,r}(x) \,d\Pbb(x) + c_5\varDelta}{\int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x) - c_5\varDelta} \\
& \leq \widetilde{\pi}_{\Pbb,r}(\Sset) + \frac{2c_5\varDelta}{\int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x) - c_5\varDelta} \\
& \leq \widetilde{\pi}_{\Pbb,r}(\Sset) + \frac{4c_5\varDelta}{\int_{\Csig} \ell_{\Pbb,r}(x) \,d\Pbb(x)} =: \widetilde{\pi}_{\Pbb,r}(\Sset) + c_7 \varDelta.
\end{align*}
where the last inequality follows from the assumption $\varDelta \leq \lambda_{\sigma}/(2^{d + 1} dr\Lambda_{\sigma})$, which implies $c_5\varDelta \leq \int_{\Csig}\ell_{\Pbb,r}(x) \,d\Pbb(x) / 2$.

\subsection{Proof of Lemma \ref{lem: nonuniform_continuous_conductance}. }
We recall that our goal is to lower bound the continuous conductance $\widetilde{\Phi}_{\Pbb,r}$. Results along these lines are already known (see e.g. \cite{kannan04}) when the density $f$ is uniform (or log-concave) and $\Csig$ is itself a convex set -- indeed, in this case stronger versions of it exist, though we will not require them. In \cite{abbasi-yadkori2016a}, a statement of this sort is made with respect to uniform density $f$ and $\Csig$ a Lipschitz deformation of a convex set, but for completeness we produce all proofs here.

\paragraph{Population-level conductance with uniform density.}
We first state an analogous result for the special case where the density $f$ is uniform everywhere on $\Csig$. For $x \in \Csig$ and measurable $\Sset \subseteq \Csig$, let 
\begin{equation*}
\widetilde{P}_{\nu,r}(x; \Sset) := \frac{\nu(\Sset \cap B(x,r))}{\nu(\Csig \cap B(x,r))}, \quad \widetilde{\pi}_{\nu,r}(\Sset) = \frac{\int_{\Sset} \ell_{\nu,r}(x) \dx}{\int_{\Csig} \ell_{\nu,r}(x) \dx}, \quad  \widetilde{Q}_{\nu,r}(\Sset_1,\Sset_2) := \int_{\Sset_1} \widetilde{P}_{\nu,r}(x;\Sset_2) \,d\widetilde{\pi}_{\nu,r}(x).
\end{equation*}
The uniform continuous normalized cut and conductance are then defined analogously to the $\Pbb$-weighted case,
\begin{equation*}
\widetilde{\Phi}_{\nu,r}(\Sset) := \frac{\widetilde{Q}_{\nu,r}(\Sset, \Sset^c)}{\min\set{\widetilde{\pi}_{\nu,r}(\Sset),\widetilde{\pi}_{\nu,r}(\Sset^c)}}, \quad \widetilde{\Phi}_{\nu,r} := \min_{\Sset \subseteq \Csig} \widetilde{\Phi}_{\nu,r}(\Sset).
\end{equation*}
where the minimum is over measurable sets $\Sset \subseteq \Csig$.

\begin{lemma}
	\label{lem: uniform_continuous_conductance}
	Let $\Csig$ satisfy Assumption \ref{asmp: embedding} for some convex set $\mathcal{K}$ with diameter $\rho$, and measure-preserving mapping $g: \mathcal{K} \to \Csig$ with Lipschitz constant $L$. Then, for any $0 < r < \frac{\sigma}{2\sqrt{d}}$, the uniform conductance $\widetilde{\Phi}_{\nu,r}$ satisfies
	\begin{equation*}
	\widetilde{\Phi}_{\nu,r} > \frac{r}{2^{13} \rho L \sqrt{d}}.
	\end{equation*}
\end{lemma}

Most of the technical work needed to show Lemma~\ref{lem: nonuniform_continuous_conductance} involves proving Lemma~\ref{lem: uniform_continuous_conductance}. We first show that  Lemma~\ref{lem: nonuniform_continuous_conductance} is a simple consequence of Lemma~\ref{lem: uniform_continuous_conductance} -- and \ref{asmp: bounded_density} -- before turning to prove Lemma~\ref{lem: uniform_continuous_conductance}.

To relate the uniform- and $\Pbb$-continuous conudctances, observe that by \ref{asmp: bounded_density} we obtain
\begin{align*}
\frac{\widetilde{Q}_{\Pbb,r}(\Sset,\Sset^c)}{\piwt_{\Pbb,r}(\Sset)} & = \frac{\int_{\Sset} \Pbb(\Sset^c \cap B(x,r)) \,d\Pbb(x)}{\int_{\Sset} \Pbb(\Csig \cap B(x,r)) \,d\Pbb(x)} \\
& \geq \frac{\lambda_{\sigma}^2 \int_{\Sset} \nu(\Sset^c \cap B(x,r)) \dx}{\Lambda_{\sigma}^2 \int_{\Sset} \nu(\Csig \cap B(x,r)) \dx} \\
& \geq \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2} \frac{\widetilde{Q}_{\nu,r}(\Sset,\Sset^c)}{\pi_{\nu,r}(\Sset)}
\end{align*}
with equivalent reasoning leading to bound $\widetilde{Q}_{\Pbb,r}(\Sset,\Sset^c)/\piwt_{\Pbb,r}(\Sset) \leq (\lambda_{\sigma}^2/\Lambda_{\sigma}^2) \widetilde{Q}_{\nu,r}(\Sset,\Sset^c)/\piwt_{\nu,r}(\Sset)$. Lemma~\ref{lem: nonuniform_continuous_conductance} therefore follows from Lemma~\ref{lem: uniform_continuous_conductance}, and it remains to show the latter Lemma.

\paragraph{Proof of Lemma \ref{lem: uniform_continuous_conductance}:}
Lower bounds on the conductance of geometric random walks follow a typical pattern. The first step is to establish that an isoperimetric inequality holds uniformly over the domain of interest, in this case $\Csig$. The following result contains the isoperimetric inequality we will use. It is stated and proved in \cite{abbasi-yadkori2016a}; for completeness, we reproduce the proof here.
\begin{lemma}[Isoperimetry of Lipschitz embeddings of convex sets.]
	\label{lem: nonconvex_isoperimetry}
	Let $\Csig$ satisfy Assumption \ref{asmp: embedding} for some convex set $\mathcal{K}$ with diameter $\rho$, and measure-preserving mapping $g: \mathcal{K} \to \Csig$ with Lipschitz constant $L$. Then, for any partition $(\Omega_1,\Omega_2,\Omega_3)$ of $\Csig$, 
	\begin{equation*}
	\nu(\Omega_3) \geq 2\frac{\dist(\Omega_1, \Omega_2)}{\rho L} \min(\nu(\Omega_1), \nu(\Omega_2))
	\end{equation*}
\end{lemma}
The proof of Lemma \ref{lem: nonconvex_isoperimetry} from first principles is non-trivial, even when the domain $\Csig$ is itself convex, and is a primary technical contribution of the seminal work \cite{lovasz1990}, extended by \cite{dyer1991} among (many) others. Once the result is shown in the convex setting, however, it is not hard to show that it applies to Lipschitz transformations of convex sets as well.
\begin{proof}[Proof of Lemma \ref{lem: nonconvex_isoperimetry}]
	For $\Omega_i, i = 1,2,3$, denote the preimage
	\begin{equation*}
	R_i = \set{x \in \mathcal{K}: g(x) \in \Omega_i}
	\end{equation*}
	For any $x \in R_1, y \in R_2$, 
	\begin{equation*}
	\norm{x - y} \geq \frac{1}{L}\norm{g(x) - g(y)} \geq \frac{1}{L} \dist(\Omega_1, \Omega_2). 
	\end{equation*}
	Since $x \in \Omega_1$ and $y \in \Omega_2$ were arbitrary, we have
	\begin{equation*}
	\dist(R_1, R_2) \geq \frac{1}{L} \dist(\Omega_1, \Omega_2).
	\end{equation*}
	By Theorem 2.2 of \cite{lovasz1990},
	\begin{align*}
	\nu(R_3) & \geq 2\frac{\dist(R_1, R_2)}{\rho} \min \{\nu(R_1), \nu(R_2)\} \\
	& \geq \frac{2}{\rho L} \dist(\Omega_1, \Omega_2) \min\{\nu(R_1), \vol(R_2)\}
	\end{align*}
	and by the measure-preserving property of $g$, this implies
	\begin{equation*}
	\nu(\Omega_3) \geq\frac{2}{\rho L} \dist(\Omega_1, \Omega_2) \min\{\nu(\Omega_1), \nu(\Omega_2)\}.
	\end{equation*}
\end{proof}
With an isoperimetric inequality in hand, the next step towards showing a lower bound on conductance is to prove that for any pair of sufficiently close points $u,v \in \Csig$, the total variation distance between the distributions $\widetilde{P}_{\nu,r}(u;\cdot)$ and $\widetilde{P}_{\nu,r}(v;\cdot)$ is small. 
\begin{lemma}[One-step distributions]
	\label{lem: one_step_distributions}
	Let $u,v \in \Csig$ be such that 
	\begin{equation*}
	\norm{u - v} \leq \frac{r t}{2\sqrt{d}}
	\end{equation*}
	for some $0 < t < 1$, and further assume there exists $\ell > 0$ such that $\ell_{\nu,r}(u), \ell_{\nu,r}(v) \geq \ell \nu_d r^d$. Then,
	\begin{equation*}
	\norm{\widetilde{P}_{\nu,r}(u; \cdot) - \widetilde{P}_{\nu,r}(v; \cdot)}_{\mathrm{TV}} \leq 1 + \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} - \ell
	\end{equation*}
	where $\norm{P - Q}_{\mathrm{TV}}$ is the total variation distance between probabilities $P$ and $Q$. 
\end{lemma}
For the moment, we will take Lemma \ref{lem: one_step_distributions} as given and prove Lemma \ref{lem: uniform_continuous_conductance}, before returning to prove Lemma \ref{lem: one_step_distributions}. Deriving a lower bound on conductance from a isoperimetric inequality (i.e Lemma \ref{lem: nonconvex_isoperimetry}) along with a bound on the one-step distributions is standard. Let $\Sset_1 \cup \Sset_2 = \Csig$ be an arbitrary partition of $\Csig$, and let $\ell = 6/25$. Note by Lemma \ref{lem: local_conductance}, as $r < \sigma/(2\sqrt{d})$ $\ell \nu_d r^d  \leq \ell_{\nu,r}(x)$ for all $x \in \Csig$. We will show that 
\begin{equation}
\label{eqn: uniform_continuous_conductance_1}
\int_{\Sset_1} \widetilde{P}_{\nu,r}(x; \Sset_2) d \widetilde{\pi}_{\nu,r}(x) \geq \frac{\sqrt{2} r \ell^4}{48 \rho L \sqrt{d}} \min\set{\widetilde{\pi}_{\nu,r}(\Sset_1), \widetilde{\pi}_{\nu,r}(\Sset_2)}.
\end{equation}
Then, dividing both sides by $\min\set{\widetilde{\pi}_{\nu,r}(\Sset_1), \widetilde{\pi}_{\nu,r}(\Sset_2)}$ yields the desired result~\eqref{eqn: uniform_continuous_conductance_1}, as $\Sset_1,\Sset_2$ was an arbitrary partition.

Now, consider the sets
\begin{align*}
\Sset_1' & = \set{x \in \Sset_1: \widetilde{P}_{\nu,r}(x; \Sset_2) < \frac{\ell}{4}} \\
\Sset_2' & = \set{x \in \Sset_2: \widetilde{P}_{\nu,r}(x; \Sset_1) < \frac{\ell}{4}}
\end{align*}
and $\Sset_3' = \Csig \setminus (\Sset_1' \cup \Sset_2')$. Suppose $\widetilde{\pi}_{\nu,r}(\Sset_1') < \widetilde{\pi}_{\nu,r}(\Sset_1)/2$. Then,
\begin{equation*}
\int_{\Sset_1} \widetilde{P}_{\nu,r}(x; \Sset_2) \,d\widetilde{\pi}_{\nu,r}(x) \geq \frac{\ell \widetilde{\pi}_{\nu,r}(\Sset_1)}{8}
\end{equation*}
Similarly, if $\widetilde{\pi}_{\nu,r}(\Sset_1') < \widetilde{\pi}_{\nu,r}(\Sset_1)/2$, then since
\begin{equation*}
\int_{\Sset_1} \widetilde{P}_{\nu,r}(x; \Sset_2) \,d\widetilde{\pi}_{\nu,r}(x) = \int_{\Sset_2} \widetilde{P}_{\nu,r}(x; \Sset_1) \,d\widetilde{\pi}_{\nu,r}(x)
\end{equation*}
a symmetric result holds. In either case, \eqref{eqn: uniform_continuous_conductance_1} follows.

Now, suppose $\pi_{\nu,r}(\Sset_1') \geq \pi_{\nu,r}(\Sset_1)/2$, and likewise for $\Sset_2'$ and $\Sset_2$. For every $u \in \Sset_1'$ and $v \in \Sset_2'$ we have that
\begin{equation*}
\norm{\widetilde{P}_{\nu,r}(u;\cdot) - \widetilde{P}_{\nu,r}(v;\cdot)}_{\mathrm{TV}} \geq 1 - \widetilde{P}_{\nu,r}(u;\Sset_1) - \widetilde{P}_{\nu,r}(v;\Sset_2) > 1 - \frac{\ell}{2}.
\end{equation*}
and thus by Lemma \ref{lem: one_step_distributions},
\begin{equation*}
\norm{u - v} \geq \frac{\sqrt{2 \pi } r \ell}{3 \sqrt{3 d}} \geq \frac{\sqrt{2}r \ell}{3\sqrt{d}}.
\end{equation*}
Since $u \in \Sset_1', v \in \Sset_2'$ were arbitrary, the same inequality holds for $\mathrm{dist}(\Sset_1', \Sset_2')$. Therefore by Lemma \ref{lem: nonconvex_isoperimetry}
\begin{equation}
\label{eqn: uniform_continuous_conductance_2}
\nu(\Sset_3') \geq  \frac{\sqrt{2} r \ell}{3 \rho L \sqrt{d}} \min \set{\nu(\Sset_1'), \nu(\Sset_2')}.
\end{equation}
Let $\Sset \subseteq \Csig$ be arbitrary, and note that an upper (lower) bound on $\nu(\Sset)$ implies an upper (lower) bound on $\widetilde{\pi}_{\nu,r}(\Sset)$, as follows:
\begin{equation}
\label{eqn: uniform_continuous_conductance_3}
\ell \nu_d r^d \nu(\Sset) \leq \widetilde{\pi}_{\nu,r}(\Sset) \leq \nu_d r^d \nu(\Sset).
\end{equation}
We now prove~\eqref{eqn: uniform_continuous_conductance_1}, the desired result:
\begin{align*}
\int_{\Sset_1} \widetilde{P}_{\nu,r}(x; \Sset_2) \dx & = \frac{1}{2} \biggl(\int_{\Sset_1} \widetilde{P}_{\nu,r}(x; \Sset_2)  \,d\widetilde{\pi}_{\nu,r}(x) + \int_{\Sset_2} \widetilde{P}_{\nu,r}(x; \Sset_1) \,d\widetilde{\pi}_{\nu,r}(x)\biggr) \\
& \geq \frac{\ell}{8} \Bigl( \widetilde{\pi}_{\nu,r}(\Sset_1 \setminus \Sset_1') + \widetilde{\pi}_{\nu,r}(\Sset_2 \setminus \Sset_2') \Bigr) \\
& \geq \frac{\ell}{8} \widetilde{\pi}_{\nu,r}(\Sset_3') \\
& \geq \frac{\ell^2}{8 \nu(\Csig)} \nu(\Sset_3') \tag{by \eqref{eqn: uniform_continuous_conductance_3}}\\
& \geq \frac{\sqrt{2} r \ell^3}{24 \rho L \sqrt{d} \nu(\Csig)} \min \set{\nu(\Sset_1'), \nu(\Sset_2')} \tag{by \eqref{eqn: uniform_continuous_conductance_2}}\\
& \geq \frac{\sqrt{2} r \ell^4}{24 \rho L \sqrt{d}} \min \set {\widetilde{\pi}_{\nu,r}(\Sset_1'), \widetilde{\pi}_{\nu,r}(\Sset_2')} \tag{by \eqref{eqn: uniform_continuous_conductance_3}}\\
& \geq \frac{\sqrt{2} r \ell^4}{48 \rho L \sqrt{d}} \min \set {\widetilde{\pi}_{\nu,r}(\Sset_1), \widetilde{\pi}_{\nu,r}(\Sset_2)}
\end{align*}
where the last line follows since $\pi_{\nu,r}(\Sset_1') \geq \pi_{\nu,r}(\Sset_1)/2$, and likewise for $\Sset_2'$ and $\Sset_2$. To prove Lemma \ref{lem: uniform_continuous_conductance}, it remains to prove Lemma \ref{lem: one_step_distributions}, and we do this next.

\paragraph{Proof of Lemma \ref{lem: one_step_distributions}:}

The key result needed to show Lemma \ref{lem: one_step_distributions} deals with volume of the overlap $B(u,r) \cap B(v,r)$. We state this result and immediately prove it.
\begin{lemma}
	\label{lem: overlap_balls_1}
	Let $u,v \in \Rd$ be points such that $\norm{u - v} \leq t\frac{r}{2\sqrt{d}}$ for some $0 < t < 1$. Then,
	\begin{equation*}
	\nu(B(u,r) \cap B(v,r)) \geq \nu_d r^d \left(1 - \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} \right)
	\end{equation*}
\end{lemma}
\begin{proof}
	We will treat only the case where $\norm{u - v} = t\frac{r}{2\sqrt{d}}$; if they are closer together the overlap of the volume will only increase. Then, it is not hard to see that $\mathcal{I} := B(u,r) \cap B(v,r)$ consists of two symmetric spherical caps, each with height
	\begin{equation*}
	h = r\left(1 - \frac{t}{4\sqrt{d}}\right)
	\end{equation*} 
	As a result, by Lemma \ref{lem: volume_of_spherical_cap} we have
	\begin{equation*}
	\nu\bigl(\mathcal{I}\bigr) = \nu_d r^d I_{1 - \alpha}(\frac{d + 1}{2}; \frac{1}{2})
	\end{equation*}
	where
	\begin{equation*}
	\alpha = 1 - \frac{2r^2(1 - \frac{t}{4 \sqrt{d}}) - r^2(1 - \frac{t}{4 \sqrt{d}})^2}{r^2} = \frac{t^2}{16d}.
	\end{equation*}
	Expanding the incomplete beta function in integral form, we therefore have
	\begin{align*}
	\nu\bigl(\mathcal{I}\bigr) & = \nu_d r^d \frac{\Gamma\bigl(\frac{d}{2}+ 1\bigr)}{\Gamma\bigl(\frac{d + 1}{2}\bigr) \Gamma\bigl(\frac{1}{2}\bigr)} \int_{0}^{1 - \alpha}u^{(d-1)/2}(1 - u)^{-1/2}du \\
	& \overset{\text{(i)}}{\geq} \nu_d r^d \left(1 - \frac{\Gamma\bigl(\frac{d}{2}+ 1\bigr)}{\Gamma\bigl(\frac{d + 1}{2}\bigr) \Gamma\bigl(\frac{1}{2}\bigr)} \frac{3 t}{4\sqrt{d}} \right) \\
	& \overset{\text{(ii)}}{\geq} \nu_d r^d \left(1 -  \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} \right)
	\end{align*}
	where $\text{(i)}$ follows from Lemma \ref{lem: beta_integral} (which we may validly apply since $\alpha \leq \frac{t^2}{16d}$), and $\text{(ii)}$ from Lemma \ref{lem: beta_function}.
\end{proof}

We now complete the proof of Lemma \ref{lem: one_step_distributions}. Let $\Sset_1 \cup \Sset_2 = \Csig$ be an arbitrary partition of $\Csig$. We will show that 
\begin{equation*}
\widetilde{P}_{\nu,r}(u; \Sset_1) - \widetilde{P}_{\nu,r}(v; \Sset_1) \leq 1 + \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} - \ell.
\end{equation*}
Since this will hold for arbitrary $\Sset_1 \subset \Csig$, it will hold for the supremum over all such $\Sset_1$ as well, and therefore the same upper bound will hold for $\norm{\widetilde{P}_{\nu,r}(u; \cdot) - \widetilde{P}_{\nu,r}(v; \cdot)}_{\mathrm{TV}}$. Now, note that
\begin{equation*}
\widetilde{P}_{\nu,r}(u; \Sset_1) - \widetilde{P}_{\nu,r}(v; \Sset_1) = 1 - \widetilde{P}_{\nu,r}(u; \Sset_2) - \widetilde{P}_{\nu,r}(v; \Sset_1)
\end{equation*}
As before, let $\mathcal{I} = B(u,r) \cap B(v,r)$. Then we have
\begin{equation*}
\widetilde{P}_{\nu,r}(u; \Sset_2) \geq \frac{1}{\nu(B(u,r))} \nu(\Sset_2 \cap (B(u,r)) \geq \frac{1}{\nu(B(u,r))} \nu(\Sset_2 \cap \mathcal{I})
\end{equation*}
with a symmetric inequality holding for $\widetilde{P}_{\nu,r}(v; \Sset_1)$. As a result,
\begin{equation}
1 - \widetilde{P}_{\nu,r}(u; \Sset_2) - \widetilde{P}_{\nu,r}(v; \Sset_1) \leq 1 - \frac{1}{\nu_d r^d} \nu(\Csig \cap \mathcal{I}) \label{eqn: one_step_1}
\end{equation}
From here, some simple manipulations yield
\begin{align}
\nu(\Csig \cap \mathcal{I})  & = \nu(\mathcal{I}) - \nu(\mathcal{I} \setminus \Csig) \nonumber \\
& \geq \nu\bigl(\mathcal{I}\bigr) - \max \set{\nu\bigl(B(u,r) \setminus \Csig\bigr), \nu\bigl(B(v,r) \setminus \Csig\bigr)} \nonumber \\
& \geq \nu_d r^d \left(1 - \frac{3 \sqrt{3} t}{4\sqrt{2\pi}}\right) - \max \set{\nu\bigl(B(u,r) \setminus \Csig\bigr), \nu\bigl(B(v,r) \setminus \Csig\bigr)} \tag{Lemma \ref{lem: overlap_balls_1}} \\
& \geq \nu_d r^d \left(1 - \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} - (1 -\ell) \right) = \nu_d r^d\left(\ell - \frac{3 \sqrt{3} t}{4\sqrt{2\pi}}\right) \label{eqn: one_step_2} 
\end{align}
where the last inequality follows by the hypothesis $\ell_{\nu,r}(u),\ell_{\nu,r}(v) \geq \ell \nu_d r^d$. Then~\eqref{eqn: one_step_2} along with~\eqref{eqn: one_step_1} give the desired result
\begin{align*}
\widetilde{P}_{\nu,r}(u; \Sset_1) - \widetilde{P}_{\nu,r}(v; \Sset_1) & \leq  1 - \frac{1}{\nu_d r^d} \nu(\Csig \cap \mathcal{I}) \\
& \leq 1 - \ell + \frac{3 \sqrt{3} t}{4\sqrt{2\pi}}.
\end{align*}
We have completed our proof of Lemma \ref{lem: one_step_distributions}, and therefore (in turn), Lemmas~\ref{lem: uniform_continuous_conductance}, \ref{lem: nonuniform_continuous_conductance}, and \ref{lem: graph_conductance_profile_lb}. As a result, Proposition \ref{prop: local_spread_conductance} is proved, and the proof of Theorem \ref{thm: mixing_time_upper_bound} is complete.

\section{Probabilistic bounds}
\label{sec: concentration}

Given a symmetric kernel function $K: \Rd \times \Rd \to \Reals$ with $\sup_{x,y \in \Rd} \abs{K(x,y)} \leq 1$, and independent and identically distributed data $\set{x_1, \ldots, x_n}$, we define the \textit{order-$2$ $U$-statistic} to be 
\begin{equation*}
U := \frac{1}{ {n \choose 2} } \sum_{1 \leq i < j \leq n} K(x_i,x_j).
\end{equation*}

\begin{lemma}[Hoeffding's inequality for $U$-statistics.]
	\label{lem: bounded_difference}
	For any $t > 0$,
	\begin{equation*}
	\mathbb{P}(\abs{U - \mathbb{E}U} \geq t) \leq 2 \exp\left\{-nt^2\right\}
	\end{equation*}
	Further, for any $\delta > 0$, we have
	\begin{align*}
	U & \leq \mathbb{E}U + \sqrt{\frac{\log(1 / \delta)}{n} }, \\
	U & \geq \mathbb{E}U - \sqrt{\frac{\log(1 / \delta)}{n} }
	\end{align*}
	each with probability at least $1 - \delta$. 
\end{lemma}

We will employ a sharper inequality to bound the sum of independent Bernoulli random variables.
\begin{lemma}
	\label{lem: multiplicative_Hoeffding}
	Let $X_i \in \set{0,1}$ for $i = 1, \ldots, n$ and let $\mu = \sum_{i = 1}^n \Ebb(X_i)$. Then,
	\begin{align*}
	\Pbb\left(\sum_{i=1}^{n}X_i > (1 + \epsilon) \mu \right) \leq \exp \left(\frac{-\epsilon^2 \mu}{3}\right) \\
	\Pbb\left(\sum_{i=1}^{n}X_i < (1 - \epsilon) \mu \right) \leq \exp \left(\frac{-\epsilon^2 \mu}{2}\right)
	\end{align*}
\end{lemma}

We will require large-sample and asymptotic bounds on several different degree and volume functionals, which follow from Lemmas \ref{lem: bounded_difference} and \ref{lem: multiplicative_Hoeffding}. Let $\Cset' \in \Cbb_f(\lambda)$ be a high-density cluster, with $\sigma$-expansion $\Csig' = \Cset' + \sigma B$, and recall the notation $\widetilde{\deg}_{n,r}(x) = \deg_{n,r}(x,G_{n,r}[\Csig[\Xbf]])$ for $x \in \Csig[\Xbf]$. Then, let
\begin{alignat*}{2}
\degminwt & := \min_{u \in \Csig[\Xbf]} \deg(u; G_{n,r}\left[\Csig[\Xbf]\right]),~ && \degmaxwt := \max_{u \in \Csig[\Xbf]} \deg(u; G_{n,r}\left[\Csig[\Xbf]\right]) \\ 
\degminpr & := \min_{u \in \Csig'[\Xbf]} \deg(u; G_{n,r}),~ && \degmax := \max_{u \in \Csig[\Xbf]} \deg(u; G_{n,r}) \\
\degmin & := \min_{u \in \Xbf} \deg(u; G_{n,r}).
\end{alignat*}
We collect the bounds we need in Lemma \ref{lem: ball_bounds_in_probability}.
\begin{lemma}
	\label{lem: ball_bounds_in_probability}
	Under the geometric conditions \ref{asmp: bounded_density}-\ref{asmp: bounded_volume}, for any $\epsilon > 0$, each of the following bounds hold with probability tending to one as $n \to \infty$:
	\begin{align}
	\frac{\degminwt}{n - 1} & \geq (1 - \epsilon)\frac{6}{25} \lambda_{\sigma} r^d \nu_d \label{eqn: min_cluster_degree_bound} \\
	\frac{\degmax}{n} & \leq (1 + \epsilon)\Lambda_{\sigma} r^d \nu_d \label{eqn: max_degree_bound}\\
	(1 - \epsilon)\Pbb(\Csig) & \leq \frac{\abs{\Csig[\Xbf]}}{n} \leq (1 + \epsilon)\Pbb(\Csig) \label{eqn: Csig_cardinality_bound}
	\end{align}

	Additionally, with probability one,
	\begin{align}
	\liminf_{n \to \infty} \frac{\degminwt}{n - 1} & \geq \frac{6}{25} \lambda_{\sigma} r^d \nu_d \label{eqn: min_cluster_degree_bound_asymp}\\
	\limsup_{n \to \infty} \frac{\widetilde{\vol}_{n,r}(\Csig[\Xbf])}{n(n-1)} & \leq \Pbb(\Csig) \Lambda_{\sigma} \nu_d r^d \label{eqn: max_vol_bound_asymp}
	\end{align}

	If additionally \ref{asmp: C'_bounded_density} holds, then for any $\epsilon > 0$, the following bound holds with probability tending to one as $n \to \infty$:
	\begin{equation}
	\label{eqn: min_cluster_degree_bound_2}
	\frac{\degminpr}{n} \geq (1 - \epsilon)\frac{6}{25} \lambda_{\sigma} r^d \nu_d
	\end{equation}
\end{lemma}
\begin{proof}
	We first note that \eqref{eqn: Csig_cardinality_bound} follows directly from the law of large numbers. Now, we examine~\eqref{eqn: min_cluster_degree_bound}. Note that for any $x \in \Csig$, by assumption \ref{asmp: bounded_density} and Lemma~\ref{lem: local_conductance},
	\begin{equation*}
	\Ebb(\widetilde{\deg}_{n,r}(x_i)|x_i = x) \geq \frac{6}{25} \lambda_{\sigma}r^d\nu_d,
	\end{equation*}
	Conditional on $x_i = x$, the functional $\widetilde{\deg}_{n,r}(x_i)$ is a sum of $n - 1$ independent Bernoulli random variables. Therefore by Lemma \ref{lem: multiplicative_Hoeffding}
	\begin{equation*}
	\Pbb\left(\widetilde{\deg}_{n,r}(x_i) \leq (n-1) (1 - \epsilon) \frac{6}{25}  \lambda_{\sigma} r^d \nu_d \mid x_i = x \right) \leq \exp\set{- \frac{6 n \epsilon^2  \lambda_{\sigma} r^d \nu_d}{50} }
	\end{equation*}
	Since this holds for all $x \in \Csig$, by the law of total probability it holds unconditionally as well, i.e.
	\begin{equation*}
	\Pbb\left(\widetilde{\deg}_{n,r}(x_i) \leq (n - 1) (1 - \epsilon) \frac{6}{25}  \lambda_{\sigma} r^d \nu_d \right) \leq \exp\set{- \frac{6 n \epsilon^2  \lambda_{\sigma} r^d \nu_d}{50} }
	\end{equation*}
	Taking a union bound over all $x_i \in \Csig[\Xbf]$ (of which, of course, there are no more $n$), we have
	\begin{equation*}
	\Pbb \left(\degminwt \leq (n - 1) (1 - \epsilon) \frac{6}{25}  \lambda_{\sigma} r^d \nu_d \right) \leq n \exp\set{- \frac{6 n \epsilon^2  \lambda_{\sigma} r^d \nu_d}{50} } \overset{n \to \infty}{\longrightarrow} 0.
	\end{equation*}
	and we have shown \eqref{eqn: min_cluster_degree_bound}. The equivalent bound \eqref{eqn: max_degree_bound} on maximum degree of vertices $x \in \Csig[\Xbf]$ follows from similar reasoning. Under~\ref{asmp: C'_bounded_density}, equivalent reasoning leads to \eqref{eqn: min_cluster_degree_bound_2}. Additionally, taking $\epsilon_n \downarrow 0$ sufficiently slowly so that 
	\begin{equation*}
	\sum_{n = 1}^{\infty }n \exp\set{- \frac{6 n \epsilon_n^2  \lambda_{\sigma} r^d \nu_d}{50} } < \infty
	\end{equation*}
	application of the Borel-Cantelli Lemma yields~\eqref{eqn: min_cluster_degree_bound_asymp}.

	Now, for any $\epsilon > 0$, letting $\mu = \int_{\Csig} \Pbb(B(x,r)) \,d\Pbb(x) = \Ebb\left(\frac{\widetilde{\vol}_{n,r}(\Csig[\Xbf])}{n(n-1)}\right)$ we have
	\begin{equation*}
	\Pbb\left(\frac{\widetilde{\vol}_{n,r}(\Csig[\Xbf])}{n(n-1)} \geq (1 + \epsilon)\mu \right) \leq \exp(-n\epsilon^2\mu).
	\end{equation*}
	Therefore, taking a sequence $\epsilon_n$ such that $\epsilon_n \downarrow 0$ sufficiently slowly so that
	\begin{equation*}
	\sum_{n = 1}^{\infty}\exp(-n\epsilon_n^2\mu) < \infty,
	\end{equation*}
	the almost-sure bound~\eqref{eqn: max_vol_bound_asymp} follows by the Borel-Cantelli Lemma.
\end{proof}

\section{Proof of Theorem \ref{thm: misclassification_rate}}
\label{sec: proof_of_misclassification_rate}
In this section, we prove Theorem \ref{thm: misclassification_rate}.
The result is in essence a consequence of i) a result of \citet{zhu2013} relating the quality of the PPR output to normalized cut and mixing time, and ii) the upper bounds in Theorem~\ref{thm: conductance_upper_bound} 
and Theorem~\ref{thm: mixing_time_upper_bound} on these functionals.

We begin with Lemma \ref{lem: zhu}, which is a restatement of Corollary 3.3 of \citet{zhu2013}, translated into notation consistent with the rest of this paper. It is stated with respect to an arbitrary $S \subseteq G = (V,E)$, and a good set $S^g \subseteq S$ with $\vol(S^g; G) \geq \vol(S; G)/2$.

\begin{lemma}
	\label{lem: zhu}
	Let $\varepsilon \geq 0$, and suppose $v \in S^g$ and $\alpha \leq \frac{1}{9\tau_{\infty}(G[S])}$. Consider the approximate PPR vector $p^{(\varepsilon)}_v := p^{(\varepsilon)}(v,\alpha;G)$. Then,
	\begin{equation}
	\label{eqn:zhu_1}
	p^{(\varepsilon)}_v(V \setminus S) \leq 2\frac{\Phi(S;G)}{\alpha},
	\end{equation}
	and additionally there exists a residual vector $p_{\ell} \in [0,1]^V$ with $\norm{p_{\ell}}_1 \leq 2\Phi(S;G)/\alpha$ such that
	\begin{equation}
	\label{eqn:zhu_2}
	\textrm{for all $u \in S$}, \quad p^{(\varepsilon)}_v(u) \geq \frac{4}{5} \frac{\deg(u;G[S])}{\vol(S;G)} - \varepsilon \deg(u;G) - p_{\ell}(u).
	\end{equation}
\end{lemma}

\noindent Now, consider the sweep cut set
\begin{equation*}
S_{\beta} := \set{u \in V: p^{(\varepsilon)} \geq \beta \pi_0},
\end{equation*}
and suppose the conditions (on seed node $v$ and teleportation parameter $\alpha$) of Lemma~\ref{lem: zhu} are met. Some straightforward calculations show the implications of Lemma~\ref{lem: zhu} on the symmetric set difference metric $\Delta(S_{\beta},S)$. In particular,
\begin{equation}
\abs{S_{\beta} \setminus S} \leq \frac{p_v(V \setminus S; G)}{\min_{u \in S_{\beta}}p^{(\varepsilon)}_v(u)} \leq \frac{2 \Phi(S;G)/\alpha}{\beta \pi_0}. \label{eqn:zhu_3}
\end{equation}
To upper bound $\abs{S \setminus S_{\beta}}$, we first observe that by definition, for every node $u \in S \setminus S_{\beta}$ $p^{(\varepsilon)}(u) \leq \beta \pi_0$. Therefore by \eqref{eqn:zhu_2} we have that the residual vector is lower bounded 
\begin{equation}
\label{eqn:zhu_5}
p_{\ell}(u) \geq \frac{4\deg(u;G[S])}{5\vol(S;G)} - \varepsilon \deg(u;G) - \beta\pi_0  = \underbrace{\left(\frac{4}{5} - \varepsilon \vol(S;G) \frac{\deg(u;G)}{\deg(u;G[S])}\right)}_{T_1(u,\varepsilon)}\frac{\deg(u;G[S])}{\vol(S;G)} - \beta\pi_0
\end{equation}

Since $\norm{p_{\ell}}_1 \leq 2\Phi(S;G)/\alpha$, this implies an upper bound on the number of vertices in $S \setminus S_{\beta}$. In particular,
\begin{align}
\label{eqn:zhu_4}
\abs{S \setminus S_{\beta}} \leq \frac{2 \Phi(S;G)/\alpha}{\min_{u \in S} \{T_1(u,\varepsilon)(\deg(u;G[S])/\vol(S;G))\} - \beta\pi_0}. 
\end{align}

\noindent Together \eqref{eqn:zhu_3} and \eqref{eqn:zhu_4} establish an upper bound on the symmetric set difference $\Delta(S_{\beta},S)$ for a general graph $G$, in terms of several graph functionals and input parameters to Algorithm \ref{alg: ppr}. We will move on to establishing bounds on these functionals and parameters in the specific case of $G = G_{n,r}$ and $S = \Csig[\Xbf]$, and the ranges for parameters specified in \eqref{eqn: initialization}. First, however, we verify that the conditions of Lemma \ref{lem: zhu} are satisfied under the setup of Theorem~\ref{thm: misclassification_rate}. Indeed,
\begin{enumerate}
	\item The initialization condition of \eqref{eqn: initialization} explicitly restricts $v \in \Csig[\Xbf]^g$,
	\item By Theorem \ref{thm: mixing_time_upper_bound} along with the initialization condition of \eqref{eqn: initialization}, we have that with probability tending to 1 as $n \to \infty$, $\alpha \leq 1/(9\tau_{\infty}(G_{n,r}[\Csig[\Xbf]]))$,
	\item the \pprspace vector $\pbf(v,\alpha;G_{n,r})$ is simply an approximate \pprspace vector with $\varepsilon = 0$.
\end{enumerate}

All conditions of Lemma~\ref{lem: zhu} are therefore satisfied. We now collect the bounds on graph functionals and algorithm parameters we will use to complete the proof of Theorem~\ref{thm: misclassification_rate}.
\begin{enumerate}
	\item By Theorem 3, with probability tending to $1$ as $n \to \infty$, $\Phi(\Csig[\Xbf]; G_{n,r}) \leq \Phibf(\theta)$. By \eqref{eqn: initialization}, $\alpha \geq 1/(10 \taubf(\theta))$. Therefore, recalling the definition of $\kappa(\Cset)$ in Definition~\eqref{def:well_conditioned_density_cluster}, we have $2 \Phi(\Csig[\Xbf]; G_{n,r})/\alpha \leq 20 \kappa(\Cset)$.
	\item Since $\pbf(v,\alpha;G_{n,r})$ is an exact PPR vector (i.e. an approximate PPR vector with $\varepsilon = 0$), the expression $T_1(u,\varepsilon)$ defined in \eqref{eqn:zhu_5} reduces to $4/5$.
	\item From Lemma~\ref{lem: ball_bounds_in_probability} we have that with probability tending to $1$ as $n \to \infty$, for any $\epsilon > 0$, the following statements hold:
	\begin{equation*}
	\min_{u \in \Csig[\Xbf]} \deg(u; \widetilde{G}_{n,r}) = \degminwt \geq 6/25 n (1 - \epsilon) \lambda_{\sigma} \nu_d r^d
	\end{equation*}
	and 
	\begin{equation*}
	\vol(\Csig[\Xbf];\widetilde{G}_{n,r}) \leq \abs{\Csig[\Xbf]} \degmaxwt \leq \abs{\Csig[\Xbf]} n (1 + \epsilon) \Lambda_{\sigma} \nu_d r^d.
	\end{equation*}
	\item Recalling Algorithm \eqref{alg: ppr}, we have that $\widehat{C} = S_{\beta}$ for some $1/40 \leq \beta \leq 1/12$. Moreover, by the initialization condition of \eqref{eqn: initialization}, $(3/4)(\lambda_{\sigma}/(\Lambda_{\sigma} n \Pbb(\Csig)) \leq \pi_0 \leq (5/4)(\lambda_{\sigma}/(\Lambda_{\sigma} n \Pbb(\Csig))$. Therefore, by \eqref{eqn: Csig_cardinality_bound}, with probability tending to $1$ as $n \to \infty$, $\pi_0$ is contained in the random interval
	\begin{equation*}
	(1/2)(\lambda_{\sigma}/(\Lambda_{\sigma} \abs{\Csig{\Xbf}}) \leq \pi_0 \leq (3/2)(\lambda_{\sigma}/(\Lambda_{\sigma} \abs{\Csig{\Xbf}}).
	\end{equation*}
\end{enumerate}
The latter two points imply that with probability tending to $1$ as $n \to \infty$,
\begin{equation}
\label{eqn:misclassification_rate_proof_1}
4 \degminwt/(5\vol_{n,r}(\Csig[\Xbf])) - \beta\pi_0  \geq \frac{1}{24} \frac{\lambda_{\sigma}}{\Lambda_{\sigma}\abs{\Csig[\Xbf]}}.
\end{equation}
Therefore,
\begin{align}
\Delta(\Cest, \Csig[\Xbf]) & \leq \frac{2 \Phi_{n,r}(\Csig[\Xbf])/\alpha}{\beta \pi_0} + \frac{2 \Phi_{n,r}(\Csig[\Xbf])/\alpha}{4 \degminwt/(5\vol_{n,r}(\Csig[\Xbf])) - \beta\pi_0} \nonumber \\
& \leq \frac{1200 \kappa(\Cset)\Lambda_{\sigma} \abs{\Csig[\Xbf]}}{\lambda_{\sigma}} + \frac{48 \kappa(\Cset) \Lambda_{\sigma} \abs{\Csig{\Xbf}}}{\lambda_{\sigma}}. \label{eqn:misclassification_rate_proof_2}
\end{align}
and the proof of Theorem \ref{thm: misclassification_rate} is complete.

\section{Proof of Theorem \ref{thm: consistent_recovery_of_density_clusters}}
\label{sec: proof_of_consistent_cluster_recovery}

As a reminder, to prove Theorem~\ref{thm: consistent_recovery_of_density_clusters} it will be sufficient to show that under the conditions of that theorem, we have that
\begin{equation}
\label{eqn: PPR_gap}
\min_{u' \in \Cset'[\Xbf]} p_v(u') \leq \beta \pi_0 \leq \min_{u \in \Cset[\Xbf]} p_v(u), \quad \textrm{for every $\beta \in [1/40,1/12]$},
\end{equation}
holds with probability tending to $1$ as $n \to \infty$.

To show this, we first establish in Lemma~\ref{lem: setup} that when Algorithm \ref{alg: ppr} is properly initialized, the resulting PPR vector is large for every vertex $u \in \Cset$, and small for every vertex $u' \in \Cset'$. Let $\overline{\pibf}_{n,r}: \Csig[\Xbf] \to [0,1]$ be the vector given by
\begin{equation*}
\overline{\pibf}_{n,r}(x) := \frac{\widetilde{\deg}_{n,r}(x)}{\vol_{n,r}(\Csig[\Xbf])}.
\end{equation*}
(Note that $\overline{\pibf}_{n,r}$ is distinct from $\widetilde{\pibf}_{n,r}$, as we normalize by $\vol_{n,r}(\Csig[\Xbf])$ rather than $\widetilde{\vol}_{n,r}(\Csig[\Xbf])$.)

\begin{lemma} 
	\label{lem: setup}
	Let $0 < r < \sigma$ and $\alpha \leq \frac{1}{9 \tau_{\infty}(\widetilde{G}_{n,r})}$. Then the following statement holds: there exists a good set $\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$ with $\vol_{n,r}(\Csig[\Xbf]^g) \geq \vol_{n,r}(\Csig[\Xbf])/2$ such that the following bounds hold with respect to $\pbf_v := \pbf(v,\alpha;G_{n,r})$ for any $v \in \Csig[\Xbf]^g$:
	\begin{itemize}
		\item For each $u \in \Cset[\Xbf]$,
		\begin{equation}
		\label{eqn: lower_bound_PPR_in_cluster}
		\pbf_v(u) \geq \frac{4}{5} \overline{\pibf}_{n,r}(u) - \frac{20 \Phi_{n,r}(\Csig[\Xbf])/\alpha}{\degminwt}
		\end{equation}
		\item Let $\Cset' \neq \Cset \in \Cbb_f(\lambda)$ satisfy the cluster separation condition \ref{asmp: cluster_separation} with respect to $\Cset$. Then for each $u' \in \Cset'[\Xbf]$,
		\begin{equation}
		\label{eqn: upper_bound_PPR_in_other_cluster}
		\pbf_v(u') \leq \frac{20 \Phi_{n,r}(\Csig[\Xbf])/\alpha}{\degminpr}.
		\end{equation}
	\end{itemize}
\end{lemma}

We immediately note that by assumption, Algorithm \ref{alg: ppr} is well-initialized, and therefore the seed node $v$ is chosen in $\Csig[\Xbf]^g$. Since we additionally assume $\Cset$ is a $\kappa$-well-conditioned density cluster, we have that $r < \sigma/(2d) < \sigma$, and the upper bound $\alpha \leq 1/(9 \tau_{\infty}(G_{n,r}[\Csig[\Xbf]]))$ holds by Theorem \ref{thm: mixing_time_upper_bound}, with probability tending to $1$ as $n \to \infty$.
Therefore, all the conditions of Lemma \ref{lem: setup} are met.

We now collect the estimates on graph functionals and PPR parameters we will use to complete the proof of Theorem~\ref{thm: consistent_recovery_of_density_clusters}. 
\begin{itemize}
	\item By Theorem 3, with probability tending to $1$ as $n \to \infty$, $\Phi(\Csig[\Xbf]; G_{n,r}) \leq \Phibf(\theta)$. By \eqref{eqn: initialization}, $\alpha \geq 1/(10 \taubf(\theta))$. Therefore, recalling the definition of $\kappa(\Cset)$ in Definition~\eqref{def:well_conditioned_density_cluster}, we have $2 \Phi(\Csig[\Xbf]; G_{n,r})/\alpha \leq 20 \kappa(\Cset)$.
	\item By Lemma \ref{lem: ball_bounds_in_probability}, for some $\epsilon > 0$ to be chosen later, 
	\begin{equation*}
	\degminwt, \degminpr \geq (6/25)(1 - \epsilon)nr^d \nu_d \lambda_{\sigma}
	\end{equation*}
	and additionally
	\begin{equation*}
	\widetilde{\vol}_{n,r}(\Csig[\Xbf]) \leq n^2 \Pbb(\Csig) \nu_d r^d \Lambda_{\sigma}
	\end{equation*}
	each with probability tending to $1$ as $n \to \infty$.

	\item By the initialization condition of \eqref{eqn: initialization}, $\pi_0 \in [3/4,5/4] \lambda_{\sigma}/(n\Pbb(\Csig)\Lambda_{\sigma})$. 
\end{itemize}

Assuming these bounds hold, for sufficiently small choices of $\epsilon > 0$ and the constant $c$ in the upper bound \eqref{eqn: kappa_ub}, we have that for all $u \in \Cset[\Xbf]$,
\begin{equation}
\label{eqn:proof_consistent_cluster_recovery_2}
p_v(u) \geq \frac{24(1-\epsilon) \lambda_{\sigma}}{125 n \Pbb(\Csig) \Lambda_{\sigma}} - \frac{250 \kappa(\Cset)}{3(1-\epsilon)nr^d\nu_d\lambda_{\sigma}} \geq \frac{5\lambda_{\sigma}}{48\Lambda_{\sigma} \Pbb(\Csig)} \geq \pi_0/12.
\end{equation}
and for all $u' \in \Cset'[\Xbf]$,
\begin{equation}
\label{eqn:proof_consistent_cluster_recovery_1}
p_v(u') \leq \frac{250 \kappa(\Cset)}{3(1-\epsilon)nr^d\nu_d\lambda_{\sigma}} < \frac{3\lambda_{\sigma}}{160\Lambda_{\sigma} \Pbb(\Csig)} \leq \pi_0/40. 
\end{equation}
and \eqref{eqn: PPR_gap} follows.

We defer the proof of Lemma \ref{lem: setup}, and first extend Theorems~\ref{thm: misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters} to apply with respect to the aPPR vector.

\subsection{The Approximate PPR Vector}
\label{sec:appr}
We are now in a position to formally state and prove our claims regarding the a\pprspace vector. 
\begin{corollary}
	\label{cor: appr}
	Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ be a $\kappa$-well-conditioned cluster (with respect to some $\theta$). Choose input parameters $\alpha, r, \pi_0, v$ to be well-initialized in the sense of \eqref{eqn: initialization}, additionally set 
	\begin{equation}
	\label{eqn: appr_parameter}
	\varepsilon = \frac{1}{25 n^2 \Pbb(\Csig) \Lambda_{\sigma} \nu_d r^d}
	\end{equation}
	and modify Algorithm \ref{alg: ppr} to compute the a\pprspace vector $\pbf^{(\varepsilon)}_v$ rather than the exact \pprspace vector $\pbf_v$, with resulting output $\Cest$. Then
	\begin{enumerate}
		\item The upper bound on symmetric set difference metric $\Delta(\Cest,\Csig[\Xbf])$ given by \eqref{eqn: misclassification_rate_ub} is still valid.
		\item If \ref{asmp: C'_bounded_density} and \eqref{eqn: kappa_ub} hold, 
		then $\Cest \subseteq \Xbf$ is a consistent estimator for $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}.
	\end{enumerate}
\end{corollary}
\begin{proof}
To prove the first claim, note that in the proof of Theorem~\ref{thm: misclassification_rate}, the only time we use the fact that $p(v,\alpha; G_{n,r})$ is an exact PPR vector is to show $T_1(u,\varepsilon) \geq 4/5$. Now, we will show that the choice of $\varepsilon$ given by \eqref{eqn: appr_parameter} results in the upper bound $T_1(u,\varepsilon) \geq 3/5$. As a result, \eqref{eqn:misclassification_rate_proof_1} and therefore \eqref{eqn:misclassification_rate_proof_2} will still hold, up to constants.

To show the desired upper bound on $T_1(u,\varepsilon)$, observe that for any $\epsilon > 0$, with probability tending to $1$ as $n \to \infty$, 
\begin{equation}
\label{eqn: appr_parameter_lb}
\varepsilon \leq \frac{1}{25(1 - \epsilon)\vol_{n,r}(\Csig[\Xbf])},
\end{equation}
and
\begin{equation*}
\max_{u \in \Csig[\Xbf]} \bigl\{ \deg_{n,r}(u) - \widetilde{\deg}_{n,r}(u)\bigr\} \leq (19/25)(1 + \epsilon) \lambda \nu_d r^d.
\end{equation*}
Assuming these bounds hold, we can upper bound $T_1(u,\varepsilon)$ when $G = G_{n,r}$, as follows:
\begin{align*}
T_1(u,\varepsilon) & = \frac{4}{5} - \varepsilon \vol(\Csig[\Xbf];G_{n,r}) \frac{\deg(u;G_{n,r})}{\deg(u; \widetilde{G}_{n,r})} \\
& \geq \frac{4}{5} - \frac{(1 + \epsilon) \lambda \nu_d r^d + \widetilde{\deg}_{n,r}(u)}{25(1 - \epsilon)\widetilde{\deg}_{n,r}(u)} \\
& \geq \frac{4}{5} - \frac{(1 + \epsilon) \lambda \nu_d r^d + \degminwt}{25(1 - \epsilon)\degminwt} \\
& \geq \frac{4}{5} - \frac{(19/25)(1 + \epsilon) + 6/25}{6(1 - \epsilon)}
\end{align*}
and so, for $\epsilon > 0$ sufficiently small, we have $T_1(u,\varepsilon) \geq 3/5$.

%\paragraph{Proof of Corollary \ref{cor: appr}: consistent cluster estimation.}

We now prove the second claim, by showing that both bounds in \eqref{eqn: PPR_gap} hold with respect to the aPPR vector. The upper bound follows immediately. Since $p^{(\varepsilon)}_v(x) \leq p_v(x)$ for all $x \in \Xbf$, the upper bound on $\max_{u' \in \Cset'[\Xbf]} p_v(u')$ also applies to $\max_{u' \in \Cset'[\Xbf]} p^{(\varepsilon)}_v(u')$.

To show the desired lower bound on $\min_{u \in \Cset[\Xbf]} p^{(\varepsilon)}_v(u)$, we use the uniform approximation guarantee in \eqref{eqn: appr_error}. Note that for every $u \in \Cset[\Xbf]$, since $r < \sigma$, every point $x \in \Xbf$ connected to $u$ must be contained in $\Csig$. Therefore $\widetilde{\deg}_{n,r}(u) = \deg_{n,r}(u)$. Along with \eqref{eqn: appr_parameter_lb}, this means
\begin{equation*}
\varepsilon \deg_{n,r}(u) \leq \frac{\overline{\pi}_{n,r}(u)}{25(1 - \epsilon)}
\end{equation*}
and therefore by \eqref{eqn: lower_bound_PPR_in_cluster}, for every $u \in \Cset[\Xbf]$,
\begin{equation*}
p^{(\varepsilon)}_v(u) \geq \left\{\frac{4}{5} - \frac{1}{25(1 - \epsilon)}\right\} \overline{\pi}_{n,r}(u) - \frac{20 \Phi_{n,r}(\Csig[\Xbf])/\alpha}{\widetilde{\deg}_{\min}}.
\end{equation*}
\noindent Using the arguments given in the proof of Theorem~\ref{thm: consistent_recovery_of_density_clusters}, and choosing the universal constant in \eqref{eqn: kappa_ub} to be sufficiently small, the upper bound in \eqref{eqn:proof_consistent_cluster_recovery_2} follows.

\end{proof}

\subsection{Proof of Lemma~\ref{lem: setup}}

To state Lemma~\ref{lem: setup}, we introduce some notation. Let $\Dbf$ and $\Wbf$ be the degree and lazy random walk matrices over $G_{n,r}$, and $\widetilde{\Dbf}$ and $\widetilde{\Wbf}$ be the degree and lazy random walk matrices over the induced subgraph $\widetilde{G}_{n,r}$. Given a starting distribution $q$ with $\mathrm{supp}(q) \subseteq \Csig[\Xbf]$, we let
	\begin{equation}
	\widetilde{\pbf}_q = \alpha q + (1 - \alpha) \widetilde{\pbf}_{q} \widetilde{\Wbf}
	\end{equation}
	be the \pprspace vector originating from $q$ over $\widetilde{G}_{n,r}$. (When the starting distribution $q = e_v$ is a point mass at a seed node $v \in \Csig[\Xbf]$, we write $\widetilde{\pbf}_v := \widetilde{\pbf}_{e_v}$ in a slight abuse of notation).

	Our analysis will involve \emph{leakage} and \emph{soakage} vectors, defined by
	\begin{align}
	\label{eqn: leakage_soakage}
	\ell_t & := e_v (\Wbf \widetilde{I} )^t (I - \Dbf^{-1} \wDbf),~ \ell := \sum_{t = 0}^{\infty} (1 - \alpha)^t \ell_t, \nonumber \\
	s_t & := e_v (\Wbf \widetilde{I} )^t (\Wbf \widetilde{I}^c),~ s := \sum_{t = 0}^{\infty} (1 - \alpha)^{t} s_t.
	\end{align}
	where $I$ is the $n \times n$ identity matrix, $\widetilde{I}$ is an $n \times n$ diagonal matrix with $\widetilde{I}_{uu} = 1$ if $u \in \Csig[\Xbf]$ and $0$ otherwise, and $\widetilde{I}^c = I - \widetilde{I}$.

	In words, for $u \in \Csig[\Xbf]$, $\ell_t(u)$ is the probability that a random walk over $G_{n,r}$ originating from $v \in \Csig[\Xbf]$ stays within $\widetilde{G}_{n,r}$ for $t$ steps, arriving at $u$ on the $t$th step, and then ``leaks out'' of $\Csig[\Xbf]$ on the $t + 1$th step. For $w \in \Xbf \setminus \Csig[\Xbf]$, $\ell_t(w) = 0$. By contrast, for $w$ again in $\Xbf \setminus \Csig[\Xbf]$, $s_t(w)$ is the probability that a random walk originating from $v$ stays within $\Csig[\Xbf]$ for $t$ steps, and then is ``soaked up'' into $w$ on the $t + 1$ step, while $s_t(u) = 0$ for all $u \in \Csig[\Xbf]$. The vectors $\ell$ and $s$ then give the total mass leaked and soaked, respectively, by the \pprspace vector.

	We first prove \eqref{eqn: lower_bound_PPR_in_cluster}, and begin by restating some results of \citep{zhu2013}, adapted to our notation. By Lemma 3.1 of \citep{zhu2013}, there exists a good set $\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$ with $\vol(\Csig[\Xbf]^g; G_{n,r}) \geq \vol(\Csig[\Xbf]; G_{n,r})/2$ such that for every $v \in \Csig[\Xbf]^g$
	\begin{equation}
	\label{eqn: zhulemma3.1}
	\pbf_v(u) \geq \widetilde{\pbf}_v(u) - \widetilde{\pbf}_{\ell}(u), \quad \textrm{and} \quad \norm{\ell}_1 \leq \frac{2 \Phi_{n,r}(\Csig[\Xbf])}{\alpha}.
	\end{equation}
	(The result $\norm{\ell}_1 \leq \frac{2 \Phi_{n,r}(\Csig[\Xbf])}{\alpha}$ is the only result in the proof of Theorem \ref{thm: consistent_recovery_of_density_clusters} which relies on the restriction $v \in \Csig[\Xbf]^g$.)

	If additionally $\alpha \leq \frac{1}{9\tau_{\infty}(\widetilde{G}_{n,r})}$, then by Corollary 3.3 of \citep{zhu2013}, for every $u \in \Csig[\Xbf]$
	\begin{equation*}
	\widetilde{\pbf}_v(u) \geq \frac{4}{5} \overline{\pibf}_{n,r}(u)
	\end{equation*}
	and along with \eqref{eqn: zhulemma3.1}, we obtain
	\begin{equation*}
	\pbf_v(u) \geq \frac{4}{5} \overline{\pibf}_{n,r}(u) - \widetilde{\pbf}_{\ell}(u).
	\end{equation*}

	We proceed to show the upper bound $\widetilde{\pbf}_{\ell}(u) \leq \norm{\ell}_1 / \degminwt$, whence \eqref{eqn: lower_bound_PPR_in_cluster} follows by \eqref{eqn: zhulemma3.1}. We note two facts regarding $\widetilde{\pbf}_{\ell}(u)$, which hold for all $u \in \Cset[\Xbf]$. 
	\begin{enumerate}
		\item Since $r < \sigma$, $(u,w) \not\in G_{n,r}$ for any $w \not\in \Csig$. As a result, for all $t \geq 1$, $\ell_t(u) = 0$ and by extension, $\ell(u) = 0$ as well.
		\item For any $q$ such that $\sum_{w \in \Csig[\Xbf]} q(w) \leq 1$ and $u \not\in \mathrm{supp}(q)$, and any $t \geq 1$,
		\begin{align}
		\label{eqn: one_step_bound}
		q \widetilde{\Wbf}^t (u) & \leq \norm{q}_1 \max_{v \neq u} W_{vu} \nonumber \\
		& \leq \frac{1}{2\degminwt} 
		\end{align}
		where last inequality follows from the fact $(u,w) \in \widetilde{G}_{n,r}$ implies $w \in \Csig$, and therefore $\deg(w; \widetilde{G}_{n,r}) \geq \degminwt$.
	\end{enumerate}

	These facts, along with some basic algebra, lead to the desired lower bound on $\widetilde{\pbf}_{\ell}(u)$ for every $u \in \Cset[\Xbf]$:
	\begin{align*}
	\widetilde{\pbf}_{\ell}(u) & = \alpha \sum_{t = 0}^{\infty} (1 - \alpha)^t \left(\ell \widetilde{\Wbf}^t \right)(u)  \nonumber \\
	& = \norm{\ell}_1 \alpha \sum_{t = 0}^{\infty} (1 - \alpha)^t \left(\frac{\ell}{\norm{\ell}_1}  \widetilde{\Wbf}^t \right)(u)\nonumber \\
	& = \norm{\ell}_1 \alpha \sum_{t = 1}^{\infty} (1 - \alpha)^t \left(\frac{\ell}{\norm{\ell}_1}  \widetilde{\Wbf}^t \right)(u)\nonumber \\
	& \leq \frac{\norm{\ell}_1 }{\degminwt}. \tag{since $u \not\in \mathrm{supp}(\ell)$}
	\end{align*}
	and \eqref{eqn: lower_bound_PPR_in_cluster} is proved.

	We turn to showing \eqref{eqn: upper_bound_PPR_in_other_cluster}. By Lemma \ref{lem: soakage_ppr_lb}, for all $u' \not\in \Csig[\Xbf]$,
	\begin{equation*}
	\pbf_v(u') \leq \pbf_s(u').
	\end{equation*}

	Note that by \ref{asmp: cluster_separation}, for every $u \in \Csig[\Xbf], (u',u) \not\in E$ and therefore $s(u') = 0$. Some manipulations, similar to those in the preceding part of the proof, yield a lower bound on $\pbf_v(u')$ in terms of $\norm{s}_1$:
	\begin{align*}
	\pbf_s(u') & = \alpha \sum_{t = 0}^{\infty} (1 - \alpha)^t \left(s \Wbf^t\right)(u') \\
	& = \norm{s}_1 \alpha \sum_{t = 0}^{\infty} (1 - \alpha)^t \left(\frac{s}{\norm{s}_1} \Wbf^t\right)(u') \\
	& = \norm{s}_1 \alpha \sum_{t = 1}^{\infty} (1 - \alpha)^t \left(\frac{s}{\norm{s}_1} \Wbf^t\right)(u') \\
	& \leq \frac{\norm{s}_1}{2\degminpr}
	\end{align*}
	where the last inequality follows from precisely the same reasoning as \eqref{eqn: one_step_bound}. The claim follows in light of Lemma \ref{lem: total_leakage_equal_total_soakage}, along with \eqref{eqn: zhulemma3.1}.

\subsection{Linear Algebra Facts}
\label{sec: linalg}

We state here a number of basic facts which follow from matrix manipulations, which are used in the proof of Theorem \ref{thm: consistent_recovery_of_density_clusters}.

\begin{lemma}
	\label{lem: soakage_ppr_lb}
	For any $v \in \Csig[\Xbf]$ and $u \not\in \Csig[\Xbf]$,
	\begin{equation*}
	\pbf_{v}(u) \leq \pbf_{s}(u)
	\end{equation*}
	where $s$ is defined as in \eqref{eqn: leakage_soakage} and depends implicitly upon $v$.
\end{lemma}
\begin{proof}
	The statement follows from Lemma \ref{lem: sum_of_soakages} along with a series of algebraic manipulations,
	\begin{align*}
	\pbf_v(u) & = \alpha\sum_{T = 0}^{\infty} (1 - \alpha)^t \left(e_v \Wbf^T\right)(u) \\
	& = \alpha\sum_{T = 1}^{\infty} (1 - \alpha)^t \left(e_v \Wbf^T\right)(u) \\
	& \leq \alpha \sum_{T=1}^{\infty} (1 - \alpha)^T \left( \sum_{t = 0}^{T - 1} s_t \Wbf^{T - t - 1} \right)(u) \tag{Lemma \ref{lem: sum_of_soakages}}\\
	& = \alpha \sum_{t=0}^{\infty} \sum_{T = t + 1}^{\infty} (1 - \alpha)^T \left( s_t \Wbf^{T - t - 1} \right)(u)\\
	& = \alpha \sum_{t=0}^{\infty} \sum_{\Delta = 0}^{\infty} (1 - \alpha)^{\Delta + t + 1} \left( s_t \Wbf^{\Delta} \right)(u) \\
	& \leq \alpha \sum_{t=0}^{\infty} \sum_{\Delta = 0}^{\infty} (1 - \alpha)^{\Delta + t } \left( s_t \Wbf^{\Delta} \right)(u) \\
	& = \alpha \sum_{\Delta = 0}^{\infty} (1 - \alpha)^{\Delta} \left(s \Wbf^{\Delta}\right)(u) \\
	& = \pbf_s(u)
	\end{align*}
\end{proof}

Let $s_t := q(\Wbf \Ibf{S})^t(\Wbf(\Ibf{S^c}))$ be the soakage vector out of $S \subseteq V$, where $\Ibf{S}$ is a $\abs{V} \times \abs{V}$ diagonal matrix with $(\Ibf{S})_{uu} = 1$ if $u \in S$ and 0 otherwise, and $\Ibf{S^c} := \Ibf{} - \Ibf{S}$. 
\begin{lemma}
	\label{lem: sum_of_soakages}
	Let $G = (V,E)$ be an unweighted, undirected graph with associated random walk matrix $\Wbf$. For any $T \in \mathbb{N}, T \geq 1$, $q \in \Reals^{\abs{V}}$, and $S \subseteq V$
	\begin{equation}
	\label{eqn: sum_of_soakages}
	q\Wbf^T = \sum_{t = 0}^{T - 1} s_t \Wbf^{T - t - 1} + q(\Wbf \Ibf{S})^T
	\end{equation}
	In particular, if $u \in V\setminus S$, then
	\begin{equation}
	\label{eqn: sum_of_soakages_2}
	q\Wbf^T(u) = \sum_{t = 0}^{T - 1} \left(s_t \Wbf^{T - t - 1}\right)(u)
	\end{equation}
\end{lemma}
\begin{proof}	
	We show \eqref{eqn: sum_of_soakages}, from which \eqref{eqn: sum_of_soakages_2} is an immediate consequence.
	To show \eqref{eqn: sum_of_soakages}, we proceed by induction on $T$. When $T = 1$,
	\begin{equation*}
	q\Wbf = q\Wbf\Ibf{S} + q\Wbf\Ibf{S^c} =  q\Wbf\Ibf{S} + s_0. 
	\end{equation*}

	Then, for $T \in \mathbb{N},~ T \geq 2$,
	\begin{align*}
	q\Wbf^{T} & = q\Wbf^{T - 1}{\Wbf} \\
	& = \left\{\sum_{t = 0}^{T - 2} s_t \Wbf^{T - 2 - t} + q(\Wbf \Ibf{S})^{T - 1} \right\} \Wbf \tag{by the inductive hypothesis}\\
	& = \sum_{t = 0}^{T - 2} s_t \Wbf^{T - 1 - t} + q(\Wbf \Ibf{S})^{T - 1} (\Wbf \Ibf{S} + \Wbf \Ibf{S^c}) \\
	& = \sum_{t = 0}^{T - 1} s_t \Wbf^{T - 1 - t} + q(\Wbf \Ibf{S})^{T - 1} (\Wbf \Ibf{S})
	\end{align*}
	and the proof is complete.
\end{proof}

\begin{lemma}
	\label{lem: total_leakage_equal_total_soakage}
	Letting $s_t$, $\ell_t$ $s$ and $\ell$ be as in \eqref{eqn: leakage_soakage},
	\begin{equation*}
	\norm{s_t}_1 = \norm{\ell_t}_1,~ \textrm{for each $t \geq 0$}
	\end{equation*}
	and therefore $\norm{s}_1 = \norm{\ell}_1$. 
\end{lemma}
\begin{proof}
	By the definition of $s_t$ and $\ell_t$, we have
	\begin{align*}
	\norm{s_t}_1 & = \norm{q_t (\Wbf \widetilde{I}^c)}_1 \\
	& = \sum_{u \in \Xbf} \sum_{u' \in \Xbf} q_t(u) (\Wbf \widetilde{I}^c)(u, u') \\
	& = \sum_{u \in\Csig[\Xbf]} \sum_{u' \in \Csig[\Xbf]^c} \frac{q_t(u)}{(\Dbf)_{uu}} \1((u,u') \in G_{n,r}) \\
	& = \sum_{u \in \Csig[\Xbf]} \frac{q_t(u) \left((\Dbf)_{uu} - (\widetilde{\Dbf})_{uu} \right)}{(\Dbf)_{uu}} \\
	& = \norm{q_t (\Ibf{} - \Dbf^{-1} \widetilde{\Dbf})}_1 = \norm{\ell_t}_1.
	\end{align*}
\end{proof}

\section{Experiments}
\label{sec: experimental_setting}

In Section \ref{sec: experimental_setting}, we detail the experimental settings of Section \ref{sec: experiments} in the main text, and include an additional figure.

\subsection{Experimental settings for Figure \ref{fig:fig1}}
We sample points according to the density function $q$, where for $x \in \Rd$
\begin{equation}
q(x) :=
\begin{cases}
\lambda,~ & x \in [0,\sigma] \times \rho^{d-1} =: \Cset, \\
\lambda - \dist(x,\Cset)\eta,~ & x \in \Csig \setminus \Cset, \\
(\lambda - \sigma \eta) - \dist(x,\Csig)^{\gamma}, & x \in (\Csig + \sigma B) \setminus \Csig, \\
0,~ & \textrm{otherwise},
\end{cases}
\end{equation}
where $\lambda = \frac{150}{81} \sigma^{\gamma}$ and $\eta = \frac{15}{81} \sigma^{\gamma - 1}$.

In the top-left and top-middle, we show draws of $n = 20000$ samples from two different density functions. In the top-left panel, $\sigma = \rho = 3.2$, while in the top-middle panel $\sigma = .1$ and $\rho = 3.2$. (For both, $d = 2$).

The remaining four panels (top-right and the bottom row) in Figure \ref{fig:fig1} show the change in normalized cut and mixing time, respectively, as the parameters $\sigma$ (top-right and bottom-left) and $\rho$ (bottom-middle and bottom-right) are varied. In the top-right and bottom-left panels $\sigma = .1 \cdot \sqrt{2}^j, j = 1,\ldots,10$, and $\rho$ is fixed at $3.2$. In the bottom-middle and bottom-right panels, $\rho = .1 \cdot \sqrt{2}^j, j = 1,\ldots,10$ and $\sigma$ is fixed at $.1$.
For each panel, the solid lines show, up to constants, the theoretical upper bound, given by Theorem \ref{thm: conductance_upper_bound} for the top-right and bottom-left panels and Theorem \ref{thm: mixing_time_upper_bound} for the bottom-middle and bottom-right panels. The dashed lines show the computed empirical value, averaged over $m$ trials ($m = 100$ for the normalized cut, dashed lines in the top-right and bottom-left panels, and $m = 20$ for the mixing time, dashed lines in the bottom-middle and bottom-right panels). For each trial across all parameters, $r$, the neighborhood graph radius, is set throughout to be as small as possible such that the resulting graph is connected, for computational efficiency. Green lines correspond to dimension $d = 2$, whereas purple/pink lines correspond to $d = 3$.

\subsection{Experimental settings for Figure \ref{fig:fig2}}

To form each of the three rows in Figure \ref{fig:fig2}, 800 points are independently sampled following a 'two moons plus Gaussian noise model'. Formally, the (respective) generative models for the data are
\begin{align}
Z & \sim \textrm{Bern}(1/2), \theta \sim \textrm{Unif}(0, \pi) \\
X(Z,\theta) & = 
\begin{cases}
\mu_1 + (r \cos(\theta), r \sin(\theta)) + \sigma \epsilon,~ & \text{if}~ Z = 1 \\
\mu_2 + (r \cos(\theta), - r \sin(\theta)) + \sigma \epsilon,~ & \text{if}~ Z = 0
\end{cases}
\end{align}

where 
\begin{align*}
\mu_1 & = (-.5, 0),~ \mu_2 = (0,0),~ \epsilon \sim N(0, I_2) \tag{row 1} \\
\mu_1 & = (-.5, -.07),~ \mu_2 = (0,.07),~ \epsilon \sim N(0, I_2) \tag{row 2} \\
\mu_1 & = (-.5, -.125),~ \mu_2 = (0,.125),~ \epsilon \sim N(0, I_2) \tag{row 3} 
\end{align*}
for $I_d$ the $d \times d$ identity matrix. The first column consists of the empirical density clusters $C[\Xbf]$ and $C'[\Xbf]$ for a particular threshold $\lambda$ of the density function; the second column shows the ~\pprspace plus minimum normalized sweep cut cluster, with hyperparameter $\alpha$ and all sweep cuts considered; the third column shows the global minimum normalized cut, computed according to the algorithm of \cite{szlam2010}; and the last column shows a cut of the density cluster tree estimator of \cite{chaudhuri2010}.

\subsection{Performance of \pprspace with high-dimensional noise.}

Figure \ref{fig:fig3} is similar to Figure \ref{fig:fig2} of the main text, with parameters
\begin{equation*}
\mu_1 = (-.5, -.025),~ \mu_2 = (0,.025),~ \epsilon \sim N(0, I_{10}).
\end{equation*}
The gray dots in $(a)$ (as in the left-hand column of Figure \ref{fig:fig2} in the main text) represent observations in low-density regions. While the \pprspace sweep cut $(b)$ has relatively high symmetric set difference with the chosen density cut, it still recovers $C[\Xbf]$ in the sense of Definition \ref{def: consistent_density_cluster_estimation}.

\begin{figure}
	\centering
	\begin{adjustbox}{minipage=\linewidth,scale=0.8}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{example2plots/row4_true_density_cluster}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{example2plots/row4_ppr_cluster}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{example2plots/row4_conductance_cluster}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.24\linewidth}
			\includegraphics[width=\linewidth]{example2plots/row4_density_cluster}
			\caption{}
		\end{subfigure}
		\caption{}
		\label{fig:fig3}
	\end{adjustbox}
\end{figure}

\bibliographystyle{plainnat}
\bibliography{../local_spectral_bibliography}

\end{document}