### 2021-01-24 jmlr/revision/local_spectral_clustering_jmlr_revision_2.tex
Addressed (nearly) all of Ryan's comments on the revised manuscript.

\documentclass[11pt,twoside]{article}
\usepackage[preprint]{jmlr2e}
\usepackage{color}
\usepackage{fancyhdr}
\usepackage{amsfonts,epsfig,graphicx}
\usepackage{afterpage}
\usepackage{amsmath,amssymb} 
% \usepackage{fullpage}
\usepackage{epsf} 
\usepackage{graphics} 
\usepackage{amsfonts,amsmath}
% \usepackage[sort,numbers]{natbib} 
\usepackage{psfrag,xspace}
\usepackage{color,etoolbox}

\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-6cm}
\setlength{\textheight}{\paperheight}
\addtolength{\textheight}{-4cm}
\addtolength{\textheight}{-1.1\headheight}
\addtolength{\textheight}{-\headsep}
\addtolength{\textheight}{-\footskip}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\renewcommand{\floatpagefraction}{.8}%

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{float}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{bm}
\usepackage{mathtools}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ubb}{\mathbb{U}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\cut}{\mathrm{cut}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\ShortHeadings{PPR Clustering on Neighborhood Graphs}{Green, Balakrishnan and Tibshirani}
\firstpageno{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Statistical Guarantees for Local Spectral Clustering on Random Neighborhood Graphs}

\author{\name Alden Green \email ajgreen@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	\name Sivaraman Balakrishnan \email siva@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	Ryan J. Tibshirani \email ryantibs@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213}

\maketitle

\begin{abstract}
	We analyze the Personalized PageRank (PPR) algorithm, a local spectral method
	for clustering, which extracts clusters using locally-biased random walks around
	a given seed node.  In contrast to previous work, we adopt a classical
	statistical learning setup, where we obtain samples from an unknown nonparametric distribution, and aim to identify sufficiently salient clusters.  We introduce a pair of population-level functionals---the \emph{normalized cut} and \emph{conductance}, analogous to graph-based functionals of the same name---and prove that PPR, run on a neighborhood graph, recovers clusters with small population normalized cut and large population conductance. We apply our general theory to establish that PPR identifies connected regions of high density (density clusters) that satisfy a set of natural geometric conditions. We also show a converse result, that PPR can fail to recover
	geometrically poorly-conditioned density clusters, even asymptotically. Finally,
	we provide empirical support for our theory.
\end{abstract}

\begin{keywords}
	graphs, spectral clustering, density clustering, Personalized PageRank, unsupervised learning
\end{keywords}

\section{Introduction}
In this paper, we consider the problem of clustering: splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference.  Our particular focus is on spectral clustering
methods, a family of powerful nonparametric clustering algorithms. Roughly
speaking, a spectral algorithm first constructs a geometric graph $G$, where
vertices correspond to samples, and edges correspond to proximities between
samples. The algorithm then estimates a feature embedding based on a (suitable) Laplacian matrix of $G$, and applies a simple clustering technique
(like $k$-means clustering) in the embedded feature space.

When applied to geometric graphs built from a large number of samples, global
spectral clustering methods can be computationally cumbersome and insensitive to
the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to increased interest in
\emph{local} spectral clustering algorithms, which leverage locally-biased
spectra computed using random walks around some user-specified seed node.  A
popular local clustering algorithm is the Personalized PageRank (PPR) algorithm,
first introduced by \citet{haveliwala2003}, and then further developed by
several others
\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013}.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, which has led
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (such as normalized cut and conductance). In contrast, in this paper we adopt a classical statistical viewpoint, and examine what the output of local clustering on a data set reveals about the underlying density $f$ of the samples. We establish conditions on $f$ under which PPR, when appropriately tuned and initialized inside a candidate cluster $\mc{C}$, will approximately recover this candidate cluster. We pay special attention to the case where $\mc{C}$ is a \emph{density cluster} of $f$---defined as a connected component of the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$---and show precisely how PPR accounts for both geometry and density in estimating a cluster.

Before giving a more detailed overview of our main results, we review some of the aforementioned worst-case guarantees, formally define PPR on a neighborhood graph, and introduce the population-level functionals that govern the behavior of local clustering in our statistical context.

\subsection{PPR clustering}
\label{subsec:ppr_clustering}
We now review the PPR clustering algorithm, and some worst-case guarantees on the quality of the resulting cluster. Let $G = (V,E)$ be an undirected, unweighted, and connected graph. We denote by $A \in \Reals^{n \times n}$ the adjacency
matrix of $G$, with entries $A_{uv} = 1$ if $(u,v) \in E$ and $0$ otherwise.  We
also denote by $D$ the diagonal degree matrix, with entries $D_{uu} :=
\sum_{v \in V} A_{uv}$, and by $I$ the $n \times n$ identity matrix. The PPR vector $p_v = p(v,\alpha;G)$ is defined with respect to a given seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, as the solution of the following linear system:
\begin{equation}
\label{eqn:ppr_vector}
p_v = \alpha e_{v} + (1 - \alpha) p_v W,
\end{equation}
where $W = (I + D^{-1}A)/2$ is the lazy random walk matrix over
$G$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in
position $v$ and 0 elsewhere).

In practice, exactly solving the system of equations~\eqref{eqn:ppr_vector} to compute the PPR vector may be too computationally expensive. To address this limitation, \citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector (aPPR), which we will denote by \smash{$p^{(\varepsilon)}$}. We refer the curious reader to \citet{andersen2006} for a formal algorithmic definition of the aPPR vector, and limit ourselves to highlighting a few salient points: the aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time, while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn:appr_error}
\textrm{for all $u \in V$}, \quad p_v(u) - \varepsilon D_{uu}\leq
p_v^{(\varepsilon)}(u) \leq p_v(u).  
\end{equation}

Once $p_v$ or $p_v^{(\varepsilon)}$ is computed, the cluster estimate $\wh{C}$ is chosen by taking a particular sweep cut of $p_v$. For a given level $\beta > 0$, the \emph{$\beta$-sweep cut} of $p_v = (p_v(u))_{u \in V}$ is 
\begin{equation}
\label{eqn:sweep_cuts}
S_{\beta,v} := \set{u \in V: \frac{p_v(u)}{D_{uu}} > \beta}.
\end{equation}
To determine $\wh{C}$, one computes $S_{\beta,v}$ over all \smash{$\beta \in (L, U)$} (where the range $(L,U)$ is user-specified), and then outputs the cluster estimate
\smash{$\wh{C} = S_{\beta^*,v}$} with minimum normalized cut. For a set $C \subseteq V$ with complement $C^c = V \!\setminus\! C$, the \emph{cut} and \emph{volume} are respectively,
\begin{equation}
\label{eqn:cut_volume}
\cut(C;G) := \sum_{u \in C} \sum_{v \in C^c}
\1\{(u,v) \in E\},~~ \vol(C; G) := \sum_{u \in C}  \sum_{v \in V} \1\{(u,v) \in E\},
\end{equation}
and the \emph{normalized cut} of $C$ is
\begin{equation}
\label{eqn:normalized_cut}
\Phi(C; G) := \frac{\cut(C;G)}{\min \set{\vol(C; G), \vol(C^c; G)}}.
\end{equation}

\subsection{Worst-case guarantees for PPR clustering}
As mentioned, to date most analysis of local clustering has focused on worst-case guarantees, defined with respect to functionals of an a priori fixed graph $G = (V,E)$.  For instance, \cite{andersen2006} analyze the normalized cut of the cluster estimate $\wh{C}$ output by PPR, showing that when PPR is appropriately seeded within a candidate cluster $C \subseteq V$, the normalized cut of $\wh{C}$ is upper bounded $\Phi(\wh{C};G) \lesssim \sqrt{\Phi(C;G)}$.\footnote{We say $a \lesssim b$ when $a \leq $} \cite{zhu2013} build on this: they introduce a second functional, the \emph{conductance} $\Psi(G)$, defined as 
\begin{equation}
\label{eqn:conductance}
\Psi(G) := \min_{S \subseteq V} \Phi(S;G),
\end{equation}
and show that if $\Phi(C;G)$ is much smaller than $\Psi(G[C])^2$---where $G[C] = (C,E \cap (C \times C))$ is the subgraph of $G$ induced by $C$--- then (in addition to having a small normalized cut) the cluster estimate $\wh{C}$ approximately recovers $C$. Our own analysis builds on that of~\cite{zhu2013}, and we give a more detailed summary of their results in Section~\ref{sec:ub_symmetric_set_difference}.
For now, we merely reiterate that the conclusions of \citep{andersen2006,zhu2013} cannot be straightforwardly applied to our statistical setting, where the input data are random samples $\{x_1,\ldots,x_n\}$ drawn from a distribution $\mathbb{P}$, the graph $G$ is a random neighborhood graph formed by the user, and the candidate cluster is a set $\mc{C} \subset \Rd$.\footnote{Throughout, we use calligraphic notation to refer to subsets of $\Rd$.}

\subsection{PPR on a neighborhood graph}
\label{subsec:ppr_neighborhood_graph}
We now formally describe the statistical setting in which we operate, as well as the method we will study: PPR on a neighborhood graph. Let $X = \{x_1,\ldots, x_n\}$ be samples drawn i.i.d.\ from a distribution
$\Pbb$ on $\Rd$. We will assume throughout that $\Pbb$ has a density $f$ with respect to the Lebesgue measure $\nu$ on $\Rd$ . For a radius $r > 0$, we define
$G_{n,r}=(V,E)$ to be the \emph{$r$-neighborhood graph} of $X$, an
unweighted, undirected graph with vertices $V=X$, and an edge $(x_i,x_j) \in
E$ if and only if $\|x_i - x_j\| \leq r$, where $\|\cdot\|$ is the
Euclidean norm. Once the neighborhood graph $G_{n,r}$ is formed, the PPR vector $p_v$ is then computed over $G_{n,r}$, with a resulting cluster estimate $\wh{C} \subseteq X$. The precise PPR algorithm we analyze is summarized in Algorithm~\ref{alg:ppr}.

\begin{algorithm}
	\caption{PPR on a neighborhood graph}
	\label{alg:ppr}	
	{\bfseries Input:} data $X=\{x_1,\ldots,x_n\}$, radius $r > 0$, teleportation
	parameter $\alpha \in [0,1]$, seed $v \in X$, sweep cut range $(L,U)$. \\     
	{\bfseries Output:} cluster estimate $\wh{C} \subseteq V$.
	\begin{algorithmic}[1]
		\STATE Form the neighborhood graph $G_{n,r}$.
		\STATE Compute the PPR vector $p_v=p(v, \alpha; G_{n,r})$ as in
		\eqref{eqn:ppr_vector}.  
		\STATE For \smash{$\beta \in (L,U)$}, compute sweep cuts $S_{\beta}$ as in
		\eqref{eqn:sweep_cuts}. 
		\STATE Return the cluster \smash{$\wh{C} = S_{\beta^*}$}, where  
		$$
		\beta^* = \argmin_{\beta \in (L,U)}~ \Phi(S_{\beta}; G_{n,r}).
		$$
	\end{algorithmic}
\end{algorithm}

\subsection{Cluster accuracy}
We need a metric to assess the accuracy with which $\wh{C}$ estimates the candidate cluster $\mc{C}$. One commonly used metric is the misclassification error, i.e. the size of the symmetric set difference
between $\wh{C}$ and the empirical cluster $\mc{C}[X] = \mc{C} \cap X$ \citep{korostelev1993,polonik1995,rigollet2009}. We
will consider a related metric, the volume of the symmetric set difference,
which weights misclassified points according to their degree in $G_{n,r}$. To keep things simple, for a given set $S \subseteq X$ we write $\vol_{n,r}(S) := \vol(S;G_{n,r})$. 
\begin{definition}
	\label{def:volume_symmetric_set_difference}
	For an estimator \smash{$\wh{C} \subseteq X$} and a set 
	$\mathcal{C} \subseteq \Rd$, their symmetric set difference is 
	\begin{equation*}
	\wh{C} \vartriangle \mathcal{C}[X] :=
	\bigl(\wh{C} \setminus \mathcal{C}[X]\bigr) \cup
	\bigl(\mathcal{C}[X] \setminus \wh{C}\bigr).
	\end{equation*}
	Furthermore, we denote the volume of the symmetric set difference by 
	$$
	\Delta(\wh{C}, \mathcal{C}[X]) := \vol_{n,r}(\wh{C} \vartriangle \mathcal{C}[X]). 
	$$
\end{definition}

\subsection{Population normalized cut and conductance}
Next we define two population-level functionals of $\mc{C}$, the normalized cut $\Phi_{\Pbb,r}(\mc{C})$ and conductance $\Psi_{\Pbb,r}(\mc{C})$, which govern the volume of the symmetric set difference $\Delta(\wh{C},\mc{C}[X])$. Let the population-level \emph{cut} of $\mc{C}$ be the expectation (up to a rescaling) of $\cut_{n,r}(\mc{C}[X]) := \cut(\mc{C}[X]; G_{n,r})$,  and likewise let the population-level \emph{volume} of $\mc{C}$ be the expectation (up to a rescaling) of $\vol_{n,r}(\mc{C}[X]) := \vol(\mc{C}[X]; G_{n,r})$; i.e. let
\begin{equation*}
\mathrm{cut}_{\Pbb,r}(\mc{C}) := \int_{\mc{C}} \int_{\mc{C}^c} \1\{\|x - y\| \leq r\} \,d\Pbb(y) \,d\Pbb(x)~~ \mathrm{vol}_{\Pbb,r}(\mc{C}) := \int_{\mc{C}} \int_{\Rd} \1\{\|x - y\| \leq r\} \,d\Pbb(y) \,d\Pbb(x),
\end{equation*}
where $\mc{C}^c := \Rd \!\setminus\! \mc{C}$.
\begin{definition}[Population normalized cut]
	For a set $\mc{C} \subset \Rd$, distribution $\Pbb$ and radius $r > 0$, the \emph{population normalized cut} is
	\begin{equation}
	\label{eqn:population_normalized_cut}
	\Phi_{\Pbb,r}(\mc{C}) := \frac{\mathrm{cut}_{\Pbb,r}(\mc{C})}{\min\{\mathrm{vol}_{\Pbb,r}(\mc{C}), \mathrm{vol}_{\Pbb,r}(\mc{C}^c)\}}.
	\end{equation}
\end{definition}

Let $\wt{\mathbb{P}}(\cdot) = \mathbb{P}(\cdot|x \in \mc{C})$ be the conditional distribution of $x$, i.e. the distribution with density function
\begin{equation*}
\wt{f}(x) :=
\begin{cases*}
\frac{1}{\mathbb{P}(\mc{C})} f(x),~~ & \textrm{if $x \in \mc{C}$} \\
0,~~ & \textrm{otherwise.}
\end{cases*}
\end{equation*}

\begin{definition}[Population conductance]
	For a set $\mc{C} \subset \Rd$, distribution $\Pbb$ and radius $r > 0$, the \emph{population conductance} is
	\begin{equation}
	\label{eqn:population_conductance}
	\Psi_{\mathbb{P},r}(\mc{C}) = \inf_{\mc{S} \subseteq \mc{C}} \Phi_{\wt{\Pbb},r}(\mc{S}).
	\end{equation}
\end{definition}

It is quite natural that $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$ should help quantify the role geometry plays in local spectral clustering. Indeed, by construction these functionals are quite obviously the population-level analogues of the empirical quantities $\Phi_{n,r}(\mc{C}[X]) := \Phi(\mc{C}[X];G_{n,r})$ and $\Psi_{n,r}(\mc{C}[X]) := \Psi(G_{n,r}\bigl[\mc{C}[X]\bigr])$, and as we have already mentioned, these empirical quantities in turn suffice to upper bound the volume of the symmetric set difference. For this reason, similar population level functionals are used by \citep{shi2009,schiebinger2015,garciatrillos19} in the analysis of \emph{global} spectral clustering in a statistical context. We will comment more on the relationship between these works and our own results in Section~\ref{subsec:related_work}. \\

\textcolor{red}{(TODO): Possibly add a section here \emph{giving}---rather than summarizing---our main results.}

\subsection{Summary of results}
A summary of our results (and an outline for this paper) is as follows.
\begin{itemize}
	\item In Section~\ref{sec:ub_symmetric_set_difference}, we show that under certain conditions the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$ are close to their population level counterparts $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$. As a consequence, we find that $\Delta(\wh{C},\mc{C}[X])$ is small whenever $\Phi_{\Pbb,r}(\mc{C})$ is small relative to $\Psi_{\Pbb,r}(\mc{C})^2$, giving the first population-level guarantees for local clustering in a nonparametric statistical context.
\end{itemize}
In Section~\ref{sec:ppr_density_cluster}, we focus on the special case where the candidate cluster $\mc{C} = \mc{C}_{\lambda}$ is a \emph{$\lambda$-density cluster}---that is, a connected component of the upper level set $\{x: f(x) \geq \lambda\}$---an object of central interest in the statistical clustering literature, dating back to the work of \citet{hartigan1981}.
\begin{itemize}
	\item In Section~\ref{subsec:recovery_well-conditioned_density_clusters}, we derive specific upper bounds on the normalized cut $\Phi_{\Pbb,r}(\mc{C}_{\lambda})$ and conductance $\Psi_{\Pbb,r}(\mc{C}_{\lambda})$. These upper bounds depend on $\lambda$ as well as some other natural parameters, and immediately result in an upper bound on $\Delta(\wh{C},\mc{C}[X])$ in the density cluster setting (Theorem~\ref{thm:density_cluster_volume_ssd_ub}).
	\item In Section~\ref{sec:lower_bound}, we prove a negative result: we give a hard distribution $\Pbb$ with corresponding density cluster $\mc{C}_{\lambda}$ for which the symmetric set difference $\Delta(\wh{C},\mc{C}_{\lambda}[X])$ is provably large.
\end{itemize}
In Section~\ref{sec:experiments} we empirically investigate some of our conclusions, before ending with some discussion in Section~\ref{sec:discussion}.

\subsection{Related Work}
\label{subsec:related_work}
We conclude this section by summarizing some related work (in addition to the background already given above).

\paragraph{Spectral clustering.}In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. In the SBM, the ability of spectral algorithms to perform clustering---or community detection---is well-understood, dating 
back to \citet{mcsherry2001} who gives conditions under which the entire
community structure can be recovered. In more recent work, \citet{rohe2011}
upper bound the fraction of nodes misclassified by a spectral algorithm for the
high-dimensional (large number of blocks) SBM, and \citet{lei2015} extend these
results to the sparse (low average degree) regime. Relatedly,
\citet{clauset08,balakrishnan2011,li2018}, analyze the misclassification rate
when the block model exhibits some hierarchical structure. The framework we
consider, in which nodes correspond to data points sampled from an underlying 
density, and edges between nodes are formed based on geometric proximity, is
quite different than the SBM, and therefore so is our analysis.

In general, the study of spectral algorithms on neighborhood graphs has been
focused on establishing asymptotic convergence of eigenvalues and eigenvectors
of certain sample objects to the eigenvalues and eigenfunctions of corresponding
limiting operators. \citet{koltchinskii2000} establish convergence of spectral
projections of the adjacency matrix to a limiting integral operator, with
similar results obtained using simplified proofs in
\citet{rosasco10}. \citet{vonluxburg2008} studies convergence of eigenvectors of
the Laplacian matrix for a neighborhood graph of fixed radius. \citet{belkin07} and
\citet{garciatrillos18} extend these results to the regime where the radius $r
\to 0$ as $n \to \infty$.

These results are of fundamental importance. However, they remain silent on the following natural question: do the spectra of these continuum operators induce a partition of the sample space which is ``good'' in some sense? \citet{shi2009,schiebinger2015,garciatrillos19} address this question, showing that spectral algorithms will recover the latent labels in certain well-conditioned nonparametric mixture models. These works are probably the most similar to our own: the conditioning of these mixture models depend on population-level functionals resembling the population-level normalized cut and conductance introduced above, and the resulting bounds on the error of spectral clustering are comparable to those we establish in Theorem~\ref{thm:volume_ssd_ub}. However, these results focus on global rather than local methods, and impose global rather than local conditions on $\Pbb$. Moreover, they do not explicitly consider recovery of density clusters, which is an important concern of our work. We comment further on the relationship between our results and these works after Theorem~\ref{thm:volume_ssd_ub}.

\paragraph{Density clustering.} For a given threshold $\lambda \in (0,\infty)$, let \smash{$\mathbb{C}_f(\lambda)$} denote the connected components of the density upper level set $\{x \in \Rd: f(x) > \lambda\}$. In the density clustering problem, initiated by~\cite{hartigan1975}, the goal is to recover $\mathbb{C}_{f}(\lambda)$. By now, density clustering (and the related problem of level-set estimation) are relatively well-understood. \citet{polonik1995,rigollet2009} study density clustering under the symmetric set difference
metric, \citet{tsybakov1997,singh2009} describe minimax optimal level-set
estimators under Hausdorff loss and
\citet{hartigan1981,chaudhuri2010,balakrishnan2013,kpotufe11} consider
consistent estimation of the cluster tree $\{\mathbb{C}_f(\lambda): \lambda \in (0,\infty)\}$. \textcolor{red}{Add, at a minimum, references to Larry + Ale and Ingo's papers.}

We emphasize that our goal is
not to improve on these results, nor to offer a better algorithm for level set
estimation, but rather to better understand the implications of our general theory by applying it within an already well-studied framework. Indeed, seen as a density clustering algorithm, PPR has none 
of the optimality guarantees found in the aforementioned works. We should also note that since we study a local algorithm, our interest will be in a local version of the density clustering problem, where the goal is to recover a single density cluster $\mc{C}_{\lambda} \in \mathbb{C}_f(\lambda)$.

\section{Recovery of a generic cluster with PPR}
\label{sec:ub_symmetric_set_difference}

In the main result (Theorem~\ref{thm:volume_ssd_ub}) of this section, we give a high probability upper bound on $\Delta(\wh{C}, \mc{C}[X])$, in terms of the population normalized cut $\Phi_{\Pbb,r}(\mc{C})$ and conductance $\Psi_{\Pbb,r}(\mc{C})$. We build to this theorem slowly, giving new structural results in two distinct directions. First, we build on some previous work (mentioned in the introduction) to relate $\Delta(\wh{C}, \mc{C}[X])$ to the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$. Second, we argue that when $n$ is large, $\Phi_{n,r}(\mc{C}[X])$ and $\Psi_{n,r}(\mc{C}[X])$ can be suitably bounded by their population-level analogues $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$.

\subsection{PPR cluster recovery: the fixed graph case}
\label{subsec:ppr_cluster_recovery_fixed_graph}
When PPR is run on a fixed graph $G = (V,E)$ with the goal of recovering a candidate cluster $C \subset V$, \cite{zhu2013} provide the sharpest known bounds on the volume of the symmetric set difference between the cluster estimate $\wh{C}$ and candidate cluster $C$. Since these results will play a major part in our analysis, in Lemma~\ref{lem:zhu} we restate them for the convenience of the reader.\footnote{Lemma~\ref{lem:zhu} improves on Lemma~3.4 of~\cite{zhu2013} by some constant factors, and for completeness we prove Lemma~\ref{lem:zhu} in our appendix. Nevertheless, to be clear the essential idea of Lemma~\ref{lem:zhu} is no different than that of~\cite{zhu2013}, and we do not claim any novelty.}

In their most general form, the results of~\citet{zhu2013} depend on the mixing time of a lazy random walk over the induced subgraph $G[C]$. The \emph{mixing time} of a lazy random walk over a graph $G$ is
\begin{equation}
\label{eqn:mixing_time}
\tau_{\infty}(G) := \min\set{ t: \frac{{\pi}(u) - {q}_{v}^{(t)}(u)}
	{{\pi}(u)} \leq \frac{1}{4}, \; \text{for all $u,v \in V$}};
\end{equation}
here $q_v^{(t)} := e_v W^t$ is the distribution of a lazy random walk over $G$ that originates at node $v$ and runs for $t$ steps, and $\pi := \lim_{t \to \infty} q_v^{(t)}$ is the limiting distribution of $q_v^{(t)}$.
\begin{lemma}[Lemma~3.4 of \cite{zhu2013}]
	\label{lem:zhu}
	For a set $C \subseteq V$, suppose that
	\begin{equation}
	\label{eqn:zhu_condition}
	\alpha \leq \min\Bigl\{\frac{1}{45}, \frac{1}{2\tau_{\infty}(G[C])}\Bigr\},~~ \beta \leq \frac{1}{5\vol(C;G)}
	\end{equation}
	Then there exists a set $C^g \subset C$ with $\vol(C^g;G) \geq \frac{1}{2}\vol(C^g;G)$ such that for any $v \in C^g$, the sweep cut $S_{\beta,v}$ satisfies
	\begin{equation}
	\label{eqn:zhu_ub}
	\vol(S_{\beta,v} \vartriangle C;G) \leq 6\frac{\Phi(C;G)}{\alpha \beta}.
	\end{equation}
\end{lemma}
The upper bound in~\eqref{eqn:zhu_ub} does not obviously depend on the conductance $\Psi(G[C])$. However, as \cite{zhu2013} point out, letting $\pi_{\min}(G) := \min_{u \in V}\{\pi(u)\}$, it follows from Cheeger's inequality~\citep{chung1997} that 
\begin{equation}
\label{eqn:mixing_time_cheeger}
\tau_{\infty}(G) \leq \frac{\log(1/\pi_{\min}(G))}{\Psi(G)^2}.
\end{equation}
Therefore, setting (for instance) $\alpha = \frac{\Psi(G[C])^2}{2\log(1/\pi_{\min}(G))}$ and $\wh{C} = S_{\beta_0,v}$ for $\beta_0 = \frac{1}{5 \vol(C;G)}$, we obtain from~\eqref{eqn:zhu_ub} that 
\begin{equation}
\label{eqn:zhu_ub2}
\frac{\vol(C \vartriangle \wh{C};G)}{\vol(C; G)} \leq 60\frac{\Phi(C;G) \log\bigl( 1/\pi_{\min}(G[C])\bigr)}{\Psi(G[C])^2}.
\end{equation}

\subsection{Improved bounds on mixing time} 
\label{subsec:mixing_time}
Having reviewed the conclusions of~\cite{zhu2013}, we return now to our own setting, where the data is not a fixed graph $G$ but instead random samples $\{x_1,\ldots,x_n\}$, and our goal is to recover a candidate cluster $\mc{C} \subset \Rd$. Ideally, we would like to apply~\eqref{eqn:zhu_ub2} with $C = \mc{C}[X]$ and $G = G_{n,r}$, replace $\Phi_{n,r}(\mc{C}[X])$ and $\Psi_{n,r}(\mc{C}[X])$ by $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$ inside~\eqref{eqn:zhu_ub2}, and thereby obtain an upper bound on $\Delta(\wh{C};\mc{C}[X])$ that depends only on $\Pbb$ and $\mc{C}$. Unfortunately, however, there is a catch: when the graph $G = G_{n,r}$ and the candidate cluster $C = \mc{C}[X]$, as $n \to \infty$ the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$ each converge to their population-level analogues, but $\pi_{\min}\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr) \asymp 1/n$. Therefore the right hand side of~\eqref{eqn:zhu_ub2} grows at a $\log n$ rate, rendering~\eqref{eqn:zhu_ub2} a vacuous upper bound when the number of samples is sufficiently large.

To fix this, in Proposition~\ref{prop:pointwise_mixing_time} we improve the upper bound on mixing time given in~\eqref{eqn:mixing_time_cheeger}. Specifically, in~\eqref{eqn:pointwise_mixing_time} the ``start penalty'' of $\log(1/\pi_{\min}(G))$ is replaced by $\log(1/s(G))$, where $s(G)$ is the \emph{local spread} of $G$, defined as
\begin{equation*}
s(G) := d_{\min}(G) \cdot \pi_{\min}(G),
\end{equation*}
for $d_{\min}(G) = \min_{u \in V}\bigl\{\deg(u;G)\bigr\}$, and likewise $d_{\max}(G) = \max_{u \in V}\bigl\{\deg(u;G)\bigr\}$. Note that $s(G) \geq \pi_{\min}(G)$.
\begin{proposition}
	\label{prop:pointwise_mixing_time}
	Assume $d_{\max}(G)/d_{\min}(G)^2 \leq 1/16$. Then,
	\begin{equation}
	\label{eqn:pointwise_mixing_time}
	\tau_{\infty}(G) \leq \frac{13}{\ln(2)} \biggl(\frac{\ln\bigl(8/s(G)\bigr)}{\Psi(G)}\biggr)^2
	\end{equation}
\end{proposition}
While Proposition~\ref{prop:pointwise_mixing_time} can be applied to \emph{any} graph $G$ (as long as the ratio of maximum degree to squared minimum degree is at most $1/16$), it is particularly useful when $G$ is a geometric graph. In the case where $G = G_{n,r}\bigl[\mc{C}[X]\bigr]$ for a fixed radius $r$, the minimum degree $d_{\min}\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr) \asymp n$, and therefore $s\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr) \asymp 1$. We give a precise upper bound on $s\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr)$ in Proposition~\ref{prop:sample_to_population_1}, which does not grow with $n$, and in combination with Proposition~\ref{prop:pointwise_mixing_time} this allows us to remove the unwanted $\log n$ factor from the upper bound in~\eqref{eqn:zhu_ub2}.

The local spread $s(G)$ plays an intuitive role in the analysis of mixing time. Indeed, in any graph $G$ sufficiently small sets are expanders---that is, if a set $R \subseteq V$ has cardinality less than the minimum degree, the normalized cut $\Phi(R;G)$ will be much larger than the conductance $\Psi(G)$. As a consequence, a random walk over $G$ will rapidly mix over all small sets $R$, and in our analysis of the mixing time we may therefore ``pretend'' that the random walk was given a warm start over a larger set $S$. The local spread $s(G)$ simply delineates small sets $R$ from larger sets $S$. Of course, the proof of Proposition~\ref{prop:pointwise_mixing_time} requires a substantially more careful analysis, and---as with the proofs of all results in this paper---it is deferred to the appendix.

\subsection{Sample-to-population results}
\label{subsec:sample_to_population}
In Propositions~\ref{prop:sample_to_population_1} and~\ref{prop:sample_to_population_2}, we establish high probability bounds on the sample normalized cut, local spread, and conductance, in terms of their population-level analogues. We have already introduced the population normalized cut~\eqref{eqn:population_normalized_cut} and conductance~\eqref{eqn:population_conductance}; denoting $s_{n,r}(\mc{C}[X]) := s\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr)$, we now define the population local spread to be
\begin{equation}
\label{eqn:local_spread}
s_{\Pbb,r}(\mc{C}) := \min_{x \in \mc{C}} \biggl\{\frac{\bigl(\deg_{\wt{\Pbb},r}(x)\bigr)^2}{\vol_{\wt{\Pbb},r}(\mc{C})} \biggr\},
\end{equation}
where $\deg_{\wt{\Pbb},r}(x) := \int_{\Rd} \1\{\|y - x\| \leq r\} \,d\wt{\Pbb}(y)$ is the expected degree of $x$ in $G_{n,r}\bigl[\mc{C}[X]\bigr]$.

To establish these bounds, we impose the following regularity conditions on $\wt{\Pbb}$ and $\mc{C}$.
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:domain} 
	The candidate cluster $\mc{C} \subseteq \Rd$ for $d \geq 2$, is a bounded, connected, open set with Lipschitz boundary. 
	\item 
	\label{asmp:bounded_density} 
	The distribution $\wt{\Pbb}$ has a density $\wt{f}: \mc{C} \to (0,\infty)$ such that there exist $f_{\min} \leq 1 \leq f_{\max}$ for which
	\begin{equation*}
	(\forall x \in \mc{C})~~ f_{\min} \leq \wt{f}(x) \leq f_{\max}.
	\end{equation*}
\end{enumerate}
In what follows, we use $b_1,b_2,\ldots$ and $B_1,B_2,\ldots$ to refer to constants that may depend on $\Pbb$ and $\mc{C}$, but do not depend on $n$ or $\delta$. We explicitly keep track of all constants in our proofs.
\begin{proposition}
	\label{prop:sample_to_population_1}
	Fix $\delta \in (0,1/3)$. Suppose $\mc{C}$ and $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}. There exist positive constants $b_1$ and $b_2$ such that each of the following statements hold.
	\begin{itemize}
		\item With probability at least $1 - 3\exp\{-b_1\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_normalized_cut}
		\Phi_{n,r}(\mc{C}[X]) \leq (1 + 3\delta) \Phi_{\Pbb,r}(\mc{C}).
		\end{equation}
		\item With probability at least $1 - (n + 2)\exp\{-b_2\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_local_spread}
		s_{n,r}(\mc{C}[X]) \geq (1 - 3\delta) s_{\Pbb,r}(\mc{C}).
		\end{equation}
	\end{itemize}
\end{proposition}
Let $p_d := 3/4$ if $d = 2$, and else $p_d := 1/d$ for $d \geq 3$. 
\begin{proposition}
	\label{prop:sample_to_population_2}
	Fix $\delta \in (0,1/2)$. Suppose $\mc{C}$ and $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}. Then there exist positive constants $B_1$,$B_2$, and $b_3$ such that for any $n \in \mathbb{N}$ for which
	\begin{equation}
	\label{eqn:sample_to_population_conductance_sample_complexity}
	\frac{(\log n)^{p_d}}{n^{1/d}} \leq \delta \cdot B_1,
	\end{equation}
	the following inequality holds with probability at least $1 - B_2/n - (n + 2)\exp\{-b_3n\}$:
	\begin{equation}
	\label{eqn:sample_to_population_conductance}
	\Psi_{n,r}(\mc{C}[X]) \geq (1 - 2\delta) \Psi_{\Pbb,r}(\mc{C}).
	\end{equation}
\end{proposition}

A word on the proof techniques: the upper bound in~\eqref{eqn:sample_to_population_normalized_cut} follows by applying Hoeffding's inequality to control the deviations of $\cut_{n,r}(\mc{C}[X])$, $\vol_{n,r}(\mc{C}[X])$ and $\vol_{n,r}(\mc{C}^c[X])$ around their expectations (noting that each of these is an order-$2$ U-statistic). To prove the lower bound~\eqref{eqn:sample_to_population_local_spread}, we require a union bound to control the minimum degree $d_{\min}\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr)$, but otherwise the proof is similarly straightforward. On the other hand, the proof of~\eqref{eqn:sample_to_population_conductance} is considerably more complicated. Our proof relies on the recent results of \citep{garciatrillos16b}, who upper bound the  $\Leb^{\infty}$-optimal transport distance between $\mathbb{P}_n$ and $\mathbb{P}$. For further details, we refer to Appendix~\textcolor{red}{(?)}, where we prove~\eqref{eqn:sample_to_population_conductance}, as well as~\citep{garciatrillos16}, who establish the asymptotic convergence of the sample conductance.

\subsection{Cluster recovery}
\label{subsec:cluster_recovery}
As is typical in the local clustering literature, our algorithmic results will be stated with respect to specific ranges of each of the user-specified
parameters. In particular, for $\delta \in (0,1/4)$ and a candidate cluster $\mc{C} \in \Rd$, we require that some of the tuning parameters of Algorithm~\ref{alg:ppr} be chosen within specific ranges,

\begin{equation}
\label{eqn:initialization}
\begin{aligned}
& \alpha \in \Bigl[(1 - 4\delta)^2, (1 - 2\delta)^2\Bigr) \cdot
 \frac{\alpha_{\Pbb,r}(\mc{C},\delta)}{2} \\
& (L,U) \subseteq \Bigl(\frac{1}{5(1 + 2\delta)},\frac{1}{5(1 + \delta)}\Bigr) \cdot 
\frac{1}{n(n - 1)\vol_{\Pbb,r}(\mc{C})}.
\end{aligned}  
\end{equation}
where
\begin{equation}\
\label{eqn:alpha_initialization}
\alpha_{\Pbb,r}(\mc{C},\delta) := \frac{\ln(2)}{13} \cdot \frac{\Psi_{\Pbb,r}^2(\mc{C})}{\ln^2\Bigl(\frac{8}{(1 - 3\delta)s_{\Pbb,r}(\mc{C})}\Bigr)} 
\end{equation}

\begin{definition}
	If the input parameters to Algorithm \ref{alg:ppr} satisfy \eqref{eqn:initialization} for some $\mc{C} \subseteq \Rd$ and $\delta \in (0,1/4)$, we say the algorithm is $\delta$-\emph{well-initialized} with respect to $\mc{C}$.
\end{definition}

Of course, in practice it is not feasible to set tuning parameters based on the 
underlying (unknown) distribution $\Pbb$ and candidate cluster $\mc{C}$. Typically, one runs PPR over some range of
tuning parameter values and selects the cluster which has the smallest
normalized cut.

\paragraph{Cluster recovery in symmetric set difference.} By combining Lemma~\ref{lem:zhu} and Propositions~\ref{prop:pointwise_mixing_time}-\ref{prop:sample_to_population_2}, we obtain an upper bound on $\Delta(\wh{C},\mc{C}[X])$ that depends solely on the distribution $\Pbb$ and candidate cluster $\mc{C}$. To ease presentation, we introduce the \emph{condition number}, defined for a given $\mc{C} \subseteq \Rd$ and $\delta \in (0,1/4)$ as
\begin{equation}
\label{eqn:condition_number}
\kappa_{\Pbb,r}(\mc{C},\delta) := \frac{(1 + 3\delta)(1+2\delta)}{(1 - 4\delta)^2(1 - \delta)} \cdot \frac{\Phi_{\Pbb,r}(\mc{C})}{\alpha_{\Pbb,r}(\mc{C},\delta)}
\end{equation}

\begin{theorem}
	\label{thm:volume_ssd_ub} 
	Let $\mc{C} \subseteq \Rd$ and $\delta \in (0,1/4)$, and suppose that
	\begin{equation}
	\label{eqn:volume_ssd_ub_sample_complexity}
	\frac{(\log n)^{p_d}}{n^{1/d}} \leq \delta \cdot B_1,~~\textrm{and}~~\frac{(1 + \delta)}{(1 - \delta)^3} \cdot B_3 \leq n.
	\end{equation} 
	With probability at least $1 - B_2/n - 4\exp\{-b_1\delta^2n\} - 2(n + 2)\exp\{-b_2\delta^2n\} - (n + 2)\exp\{-b_3n\}$, the following statement holds: there exists a set $\mc{C}[X]^g \subseteq \mc{C}[X]$ of large volume, $\vol_{n,r}(\mc{C}[X]^g) \geq \vol_{n,r}(\mc{C}[X])/2$, such that if Algorithm~\ref{alg:ppr} is $\delta$-well-initialized with respect to $\mc{C}$, and run with any seed node $v \in \mc{C}[X]^g$, then the PPR estimated cluster $\wh{C}$ satisfies
	\begin{equation}
	\label{eqn:volume_ssd_ub}
	\begin{aligned}
	\frac{\Delta(\wh{C};\mc{C}[X])}{\vol_{n,r}(\mc{C}[X])} \leq 60 \cdot \kappa_{\Pbb,r}(\mc{C},\delta)
	\end{aligned}
	\end{equation}
\end{theorem}
To summarize: if PPR is well-initialized within a candidate cluster $\mc{C}$, then the volume of the symmetric set difference between $\wh{C}$ and $\mc{C}[X]$ is upper bounded by the following function of the population-level functionals:
\begin{equation*}
\frac{\Delta(\wh{C};\mc{C}[X])}{\vol_{n,r}(\mc{C}[X])} \lesssim \Phi_{\Pbb,r}(\mc{C})  \cdot \frac{\log^2\bigl(1/s_{\Pbb,r}(\mc{C})\bigr)}{\Psi^2_{\Pbb,r}(\mc{C})}.
\end{equation*}
We now make some other remarks:
\begin{itemize}
	\item It is useful to compare Theorem~\ref{thm:volume_ssd_ub} with what is already known regarding \emph{global} spectral clustering in the context of nonparametric statistics. \cite{schiebinger2015} consider the following variant of spectral clustering: first embed the data $X$ into $\Reals^{k}$ using the bottom $k$ eigenvectors of the degree-normalized Laplacian $I - D^{-1/2}AD^{-1/2}$, and then partition the embedded data into estimated clusters $\wh{C}_1,\ldots,\wh{C}_k$  using $k$-means clustering. They derive error bounds on the misclassification error that depend on a difficulty function $\varphi(\Pbb)$. In our context, where the goal is to successfully distinguish $\mc{C}$ and $\mc{C}^c$ and thus $k = 2$, this difficulty function is roughly
	\begin{equation}
	\label{eqn:schiebinger}
	\varphi(\Pbb) \approx \sqrt{\Phi_{\Pbb,r}(\mc{C})} \cdot \max\biggl\{ \frac{1}{\Psi_{\Pbb,r}(\mc{C})^2}; \frac{1}{\Psi_{\Pbb,r}(\mc{C}^c)^2}\biggr\}.
	\end{equation}
	We point out two ways in which~\eqref{eqn:volume_ssd_ub} is a tighter bound than~\eqref{eqn:schiebinger}. First,~\eqref{eqn:schiebinger} depends on $\Psi_{\Pbb,r}(\mc{C}^c)$ in addition to $\Psi_{\Pbb,r}(\mc{C})$, and is thus a useful bound only if $\mc{C}^c$ and $\mc{C}$ are both internally well-connected. In contrast~\eqref{eqn:volume_ssd_ub} depends only on $\Psi_{\Pbb,r}(\mc{C})$, and is thus a useful bound if $\mc{C}$ has small conductance, regardless of the conductance of $\mc{C}^c$. This is intuitive: PPR is a local rather than global algorithm, and as such the analysis requires only local rather than global conditions. Second,~\eqref{eqn:schiebinger} depends on $\sqrt{\Phi_{\Pbb,r}(\mc{C})}$ rather than $\Phi_{\Pbb,r}(\mc{C})$, and since $\Phi_{\Pbb,r}(\mc{C}) \leq 1$ this results in a weaker bound.~\citep{schiebinger2015} provide experiments suggesting that the linear, rather than square-root, dependence is correct, and we theoretically confirm this in the local clustering setup. Of course, on the other hand~\eqref{eqn:volume_ssd_ub} depends on $\log^2(1/s_{\Pbb,r}(\mc{C}))$ which does not appear in~\eqref{eqn:schiebinger}. \\
	\textcolor{red}{(TODO): Add a comment explaining why we get different bounds.}

	\item Although Theorem~\ref{thm:volume_ssd_ub} is stated with respect to the exact PPR vector $p_v$, for a sufficiently small choice of $\varepsilon$, the 
	application of \eqref{eqn:appr_error} within the proof of Theorem
	\ref{thm:volume_ssd_ub} leads to an analogous result which holds for the aPPR vector \smash{$p_v^{(\varepsilon)}$}. We formally state and proof this result in our appendix (Corollary~\textcolor{red}{(?)}).
\end{itemize}

\section{Recovery of a density cluster with PPR}
\label{sec:ppr_density_cluster}
We now apply the general theory established in the last section to the special case where $\mc{C} = \mc{C}_{\lambda}$ is a $\lambda$-density cluster---that is, a connected component of the upper level set $\{x: f(x) \geq \lambda\}$. We also derive a lower bound, giving a ``hard problem'' for which PPR will provably fail to recover a density cluster. Together, these results can be summarized as follows: PPR recovers a density cluster $\mc{C}_{\lambda}$ if and only if $\mc{C}_{\lambda}$ is \emph{well-conditioned}, both geometrically---meaning it is not too long and thin--- and vertically---meaning $f$ is approximately uniform inside $\mc{C}_{\lambda}$ while satisfying a low-noise condition near its boundary.

\subsection{Recovery of well-conditioned density clusters}
\label{subsec:recovery_well-conditioned_density_clusters}

All results on density clustering assume the density $f$ satisfies some regularity conditions. A basic requirement is the need to avoid clusters which contain arbitrarily thin bridges or spikes, or more generally clusters which can be disconnected by removing a subset of (Lebesgue) measure $0$, and thus cannot be resolved by any finite number of samples. This requirement can be enforced in several different ways: for instance, many of the aforementioned works assume that the density $f$ is continuous, while~\cite{rinaldo2010} allow for discontinuous $f$ by defining the level sets with respect to a convolution $k \ast \mathbb{P}$ (where $k$ is a continuous kernel on $\Rd$ with compact support), and \cite{steinwart2015} takes the level set with respect to the lower semi-continuous version of $f$. We instead choose to follow the approach of~\cite{chaudhuri2010}, who assume the density is lower bounded on a thickened version of $\mc{C}_{\lambda}$, defined as $\mc{C}_{\lambda,\sigma} := \{x \in \Reals^d: \dist(x,\mc{C}) \leq \sigma\}$ for some $\sigma > 0$. (Here $\dist(x,\mc{C}) := \inf_{y \in \mc{C}} \|y - x\|$.) The set $\mc{C}_{\lambda,\sigma}$ is full dimensional with a smooth boundary, and thus no continuity assumption on $f$ is required. \\

\textcolor{red}{(TODO): Follow the advice of Siva, don't irritate Ingo.}\\

In summary, our goal is to obtain upper bounds on $\Delta(\wh{C},\mc{C}_{\lambda,\sigma}[X])$, for some fixed $\lambda$ and $\sigma > 0$. We have already derived upper bounds on the volume of the symmetric set difference that depend on some population-level functionals of the candidate cluster. What remains is to analyze these population-level functionals in the specific case where the candidate cluster is $\mc{C}_{\lambda,\sigma}$. To carry out this analysis, we will need to impose some conditions, and for the rest of this section we will assume the following.

\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{2}
	\item
	\label{asmp:lambda_bounded_density}
	\emph{Bounded density within cluster:} There exist constants
	$0<\lambda_{\sigma}< \Lambda_{\sigma}<\infty$ such that 
	$$
	\lambda_{\sigma} \leq \inf_{x \in \mc{C}_{\lambda,\sigma}} f(x) \leq \sup_{x \in \mc{C}_{\lambda,\sigma}} f(x)
	\leq \Lambda_{\sigma}.
	$$

	\item 
	\label{asmp:low_noise_density}
	\emph{Low noise density:} There exist $\theta \in (0,\infty)$ and $\gamma \in
	[0,1]$ such that for any $x \in \Rd$ with $0 < \dist(x, \mc{C}_{\lambda,\sigma}) \leq \sigma$,     
	$$
	\inf_{y \in \mc{C}_{\lambda,\sigma}} f(y) - f(x) \geq  \theta \cdot \dist(x, \mc{C}_{\lambda,\sigma})^{\gamma}.  
	$$
	Roughly, this assumption ensures that the density decays sufficiently quickly
	as we move away from the target cluster $\mc{C}_{\lambda,\sigma}$, and is a standard assumption
	in the level-set estimation literature (see for instance \citet{singh2009}).

	\item
	\label{asmp:embedding}
	\emph{Lipschitz embedding:}
	There exists $g: \Reals^d \to \Reals^d$, $\rho \in (0,\infty)$ and $L \in [1,\infty)$ such that
	\begin{enumerate}
		\item we have $\mc{C}_{\lambda,\sigma} = g(\mathcal{K})$, for a convex set $\mathcal{K}
		\subseteq \Rd$ with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in
			\mathcal{K}}\|x - y\| \leq \rho < \infty$;
		\item $\det(\nabla g (x)) = 1$ for all $x \in \mc{K}$, where $\nabla g(x)$ is
		the Jacobian of $g$ evaluated at $x;$ and 
		\item for some $L \geq 1$,   
		$$
		\|g(x) - g(y)\| \leq L \|x - y\| ~
		\text{for all $x,y \in \mathcal{K}$}. 
		$$
	\end{enumerate}
	Succinctly, we assume that $\mc{C}_{\lambda,\sigma}$ is the image of a convex set with finite
	diameter under a measure preserving, Lipschitz transformation. 
\end{enumerate}

\emph{For convenience only}, we will also make the following assumption, 
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{5}
	\item
	\label{asmp:bounded_volume}
	\emph{Bounded volume:}
	The volume of $\mc{C}_{\lambda,\sigma}$ is no more than half the total volume of $\Rd$:
	$$
	\vol_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \leq \vol_{\Pbb,r}(\mc{C}_{\lambda,\sigma}^c). 
	$$
	Roughly, this assumption ensures that with high probability $\vol_{n,r}(\mc{C}_{\lambda,\sigma}[X]) \leq (1 + \delta) \cdot \vol_{n,r}(\mc{C}_{\lambda,\sigma}[X]^c)$, and thus $\Phi_{n,r}(\mc{C}_{\lambda,\sigma}[X]) \leq (1 + \delta) \cdot \cut_{n,r}(\mc{C}_{\lambda,\sigma}[X])/\vol_{n,r}(\mc{C}_{\lambda,\sigma}[X])$. 
\end{enumerate}

\paragraph{Normalized cut, conductance, and local spread of a density cluster.} In Lemma~\ref{lem:density_cluster_local_spread}, Proposition~\ref{prop:density_cluster_normalized_cut}, and Proposition~\ref{prop:density_cluster_conductance}, we give bounds on the population-level local spread, normalized cut, and conductance of $\mc{C}_{\lambda,\sigma}$. These bounds depend on the various geometric parameters just introduced. 
\begin{lemma}
	\label{lem:density_cluster_local_spread}
	Assume $\mc{C}_{\lambda,\sigma}$ satisfies Assumptions~\ref{asmp:lambda_bounded_density} and~\ref{asmp:embedding} for some $\lambda_{\sigma},\Lambda_{\sigma},\rho$ and $L$. Then,
	\begin{equation}
	\label{eqn:density_cluster_local_spread}
	s_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \geq \frac{1}{4} \cdot \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2} \cdot \biggl(\frac{2r}{\rho}\biggr)^{d} \cdot \biggl(1 - \frac{r}{\sigma} \sqrt{\frac{d + 2}{2\pi}}\biggr)
	\end{equation}
\end{lemma}

\begin{proposition}
	\label{prop:density_cluster_normalized_cut}
	Assume $\mc{C}_{\lambda,\sigma}$ satisfies Assumptions~\ref{asmp:lambda_bounded_density},~\ref{asmp:low_noise_density} and~\ref{asmp:bounded_volume} for some $\lambda_{\sigma}, \Lambda_{\sigma}, \theta$, and $\gamma$, and additionally that $0 < r \leq \frac{\sigma}{4d}$. Then,
	\begin{equation}
	\label{eqn:density_cluster_normalized_cut}
	\Phi_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \leq \frac{16}{9} \cdot \frac{dr}{\sigma} \cdot \frac{\lambda\Bigl(\lambda_{\sigma} - \theta \frac{r^{\gamma}}{\gamma + 1}\Bigr)}{\lambda_{\sigma}^2}
	\end{equation}
\end{proposition}

\begin{proposition}
	\label{prop:density_cluster_conductance}
	Assume $\mc{C}_{\lambda,\sigma}$ satisfies Assumptions~\ref{asmp:lambda_bounded_density} and~\ref{asmp:embedding} for some $\lambda_{\sigma}, \Lambda_{\sigma}, \rho$ and $L$. Then,
	\begin{equation}
	\label{eqn:density_cluster_conductance}
	\Psi_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \geq \Bigl(1 - \frac{r}{4\rho L}\Bigr) \cdot \Bigl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\Bigr)^2 \cdot \frac{\sqrt{2\pi}}{36} \cdot \frac{r}{\rho L \sqrt{d + 2}} \cdot \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2}
	\end{equation}
\end{proposition}
Some remarks are in order:
\begin{itemize}
	\item We prove Proposition~\ref{prop:density_cluster_normalized_cut} by separately upper bounding $\cut_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$ and lower bounding the volume $\vol_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$. Of these two bounds, the trickier to prove is the upper bound on the cut, which involves carefully estimating the probability mass of thin tubes around the boundary of $\mc{C}_{\lambda,\sigma}$. 
	\item Proposition~\ref{prop:density_cluster_conductance} is proved in a completely different way. The proof relies heavily on bounds on the isoperimetric ratio of convex sets (as derived by e.g. \cite{lovasz1990} or \cite{dyer1991}), and thus the embedding assumption \ref{asmp:embedding} and Lipschitz parameter $L$
	play an important role in proving the upper bound in Proposition~\ref{prop:density_cluster_conductance}. 
	\item There is some interdependence between $L$ and $\sigma,\rho$, which might lead one to hope that \ref{asmp:embedding} is
	non-essential. However, it is not possible to eliminate condition \ref{asmp:embedding} without incurring an additional factor of at least
	$(\rho/\sigma)^d$ in \eqref{eqn:density_cluster_conductance}, achieved, for
	instance, when $\mc{C}_{\lambda,\sigma}$ is a dumbbell-like set consisting of two balls of diameter $\rho$ linked by a cylinder of radius $\sigma$. In contrast,~\eqref{eqn:density_cluster_conductance} depends polynomially on $d$, and many reasonably shaped sets---such as star-shaped sets as well as half-moon shapes of the type we consider in Section \ref{sec:experiments}---satisfy \ref{asmp:embedding} for reasonably small values of $L$ \citep{abbasi-yadkori2016a, abbasi-yadkori2016}.
\end{itemize}

\noindent Applying these results along with Theorem~\ref{thm:volume_ssd_ub}, we obtain an upper bound on $\Delta(\wh{C},\mc{C}_{\lambda,\sigma}[X])$. In what follows, $C_{1,\delta},C_{2,\delta},\ldots$ are constants which may depend on $\delta$, but not on $n$, $\Pbb$ or $\mc{C}_{\lambda,\sigma}$, and which we keep track of in our proofs.
\begin{theorem}
	\label{thm:density_cluster_volume_ssd_ub}
	Let $C_{\lambda} \subseteq \Rd$ and $\delta \in (0,1/4)$. Suppose that $C_{\lambda,\sigma}$ satisfies Assumptions~\ref{asmp:lambda_bounded_density}-\ref{asmp:bounded_volume} for some $\lambda_{\sigma}, \Lambda_{\sigma}, \theta, \gamma, \rho$ and $L$, that $0 < r \leq \sigma/4d$, and that $n$ satisfies~\eqref{eqn:volume_ssd_ub_sample_complexity}. Then with probability at least $1 - B_2/n - 4\exp\{-b_1\delta^2n\} - 2(n + 2)\exp\{-b_2\delta^2n\} - (n + 2)\exp\{-b_3n\}$, the following statement holds: there exists a set $\mc{C}[X]^g \subseteq \mc{C}[X]$ of large volume, $\vol_{n,r}(\mc{C}[X]) \geq \vol_{n,r}(\mc{C}[X])/2$, such that if Algorithm~\ref{alg:ppr} is $\delta$-well-initialized with respect to $\mc{C}$, and run with any seed node $v \in \mc{C}[X]^g$, then the PPR estimated cluster $\wh{C}$ satisfies
	\begin{equation}
	\label{eqn:density_cluster_volume_ssd_ub}
	\frac{\Delta(\wh{C};\mc{C}_{\lambda,\sigma}[X])}{\vol_{n,r}(\mc{C}[X])} \leq C_{1,\delta} \cdot d^3(d + 2) \cdot \frac{L^2\rho^2}{\sigma r} \cdot \frac{\Lambda_{\sigma}^2 \lambda (\lambda - \theta \frac{r^{\gamma}}{\gamma + 1})}{\lambda_{\sigma}^4} \cdot \log^2\biggl(C_{2,\delta}^{1/d} \frac{\Lambda_{\sigma}^{2/d} L\rho}{\lambda_{\sigma}^{2/d}2r}\biggr)
	\end{equation}
\end{theorem}
Inequality~\eqref{eqn:density_cluster_volume_ssd_ub} shows the separate roles played by geometry and density in the ability of PPR to recover a density cluster. For instance, if the thickened density cluster $\mc{C}_{\lambda,\sigma}$ is both geometrically well-conditioned---i.e. $L^2\rho^2/(\sigma r) \approx 1$---and vertically well-conditioned---i.e. $\Lambda_{\sigma} \approx \lambda \approx \lambda_{\sigma}$ and $(\lambda - \theta r^{\gamma}/(\gamma + 1)) \ll \lambda$---then PPR will approximately recover it.  Some other remarks:
\begin{itemize}
	\item Observe that while the diameter $\rho$ is absent from our upper bound on normalized cut in Proposition \ref{prop:density_cluster_normalized_cut}, it enters the ultimate bound in Theorem~\ref{thm:density_cluster_volume_ssd_ub} through the conductance. This reflects
	(what may be regarded as) established wisdom regarding spectral partitioning
	algorithms more generally \citep{guattery1995, hein2010}, but newly applied
	to the density clustering setting: if the diameter $\rho$ is large, then PPR
	may fail to recover $\mc{C}_{\lambda,\sigma}[X]$ even when $\mc{C}_{\lambda}$ is sufficiently well-conditioned to ensure that $\mc{C}_{\lambda,\sigma}[X]$ has a small normalized cut in $G_{n,r}$. This will be supported by simulations in Section \ref{sec:experiments}.   
	\item Several modifications of global spectral clustering have been proposed with the intent of making such procedures essentially independent of the geometry of the density cluster $\mc{C}_{\lambda}$. For instance,~\citet{pelletier2011,arias-castro2009} introduce a cleaning step to remove low-degree vertices, whereas~\cite{little2020} use a weighted geometric graph, where the weights are computed with respect to a density-dependent distance. The resulting procedures come with stronger density cluster recovery guarantees. However, the key ingredient in such procedures is the explicitly density-dependent part of the algorithm, with spectral clustering functioning as more of a post-processing step. These methods are thus very different in spirit to PPR, which is a bona fide (local) spectral clustering algorithm. 
	\item As mentioned in the discussion after Theorem~\ref{thm:volume_ssd_ub}, the population-level normalized cut and conductance also play a leading role in the analysis of global spectral clustering algorithms. It therefore seems likely that similar bounds to~\eqref{eqn:density_cluster_volume_ssd_ub} would apply to the output of global spectral clustering methods as well, but formalizing this is outside the scope of our work.
	\item The symmetric set difference does not measure whether \smash{$\wh{C}$}
	can (perfectly) distinguish any two distinct clusters \smash{$\mc{C}_{\lambda},\mc{C}_{\lambda}' \in \mathbb{C}_f(\lambda)$}. In Appendix~\textcolor{red}{(?)}, we show in Theorem~\textcolor{red}{(?)} that the PPR estimate $\wh{C}$ can in fact distinguish two distinct clusters $\mc{C}_{\lambda}$ and $\mc{C}_{\lambda}'$, but the result holds only under relatively restrictive conditions.
\end{itemize}

\section{Negative result}
\label{sec:lower_bound}
We now exhibit a hard case for density clustering using PPR, that is, a distribution $\Pbb$ for which PPR is unlikely to recover a density
cluster. Let \smash{$\mc{C}^{(0)}, \mc{C}^{(1)}, \mc{C}^{(2)}$} be rectangles in
$\Reals^2$,    
$$
\mc{C}^{(0)} = \biggl[-\frac{\sigma}{2}, \frac{\sigma}{2}\biggr] \times 
\biggl[-\frac{\rho}{2}, \frac{\rho}{2}\biggr], \quad 
\mc{C}^{(1)} = \mc{C}^{(0)} - \set{(\sigma,0)}, \quad
\mc{C}^{(2)} = \mc{C}^{(0)} + \set{(\sigma,0)},
$$
where $0 < \sigma < \rho$, and let $\Pbb$ be the mixture distribution over
\smash{$\mathcal{X} = \mc{C}^{(0)} \cup \mc{C}^{(1)} \cup \mc{C}^{(2)}$} given by   
$$
\Pbb = \frac{1 - \epsilon}{2} \Pbb_1 + \frac{1 - \epsilon}{2} \Pbb_2 +
\frac{\epsilon}{2} \Pbb_0, 
$$
where $\Pbb_k$ is the uniform distribution over $\mc{C}^{(k)}$ for $k = 0,1,2$.  
The density function $f$ of $\Pbb$ is simply
\begin{equation}
\label{eqn:lb_density}
f(x) = \frac{1}{\rho\sigma}\left(\frac{1 - \epsilon}{2}\1(x \in
\mc{C}^{(1)}) + \frac{1 - \epsilon}{2}\1(x \in \mc{C}^{(2)}) +
\epsilon\1(x \in \mc{C}^{(0)})  \right), 
\end{equation}
so that for any $\epsilon < \lambda < (1 - \epsilon)/2$, we have 
\smash{$\mathbb{C}_f(\lambda) = \set{\mc{C}^{(1)}, \mc{C}^{(2)}}$}. Figure~\ref{fig:hard_case} visualizes the density $f$ for two different choices of $\epsilon,
\sigma, \rho$.

\begin{figure}[tb]
	\centering
	\includegraphics[width=0.495\textwidth]{../lower_bound_1}
	\includegraphics[width=0.495\textwidth]{../lower_bound_2}
	\caption{\it\small The density $f$ in \eqref{eqn:lb_density}, for
		$\rho=1$, and two different choices of $\epsilon$ and $\sigma$. Left:
		$\epsilon = 0.3$ and $\sigma = 0.1$; right: $\epsilon = 0.2$ and 
		$\sigma = 0.2$.} 
	\label{fig:hard_case}
\end{figure}

\paragraph{Lower bound on symmetric set difference.}
As the following theorem demonstrates, even when Algorithm~\ref{alg:ppr} is
reasonably initialized, if the density cluster \smash{$\mc{C}^{(1)}$} is 
sufficiently geometrically ill-conditioned (in words, tall and thin) the cluster 
estimator $\wh{C}$ will fail to recover \smash{$\mc{C}^{(1)}$}. Let
\begin{equation}
\label{eqn:lower_set}
\mathcal{L} = \set{(x_1,x_2) \in \mathcal{X}: x_2 < 0}.
\end{equation}
In the following Theorem, the samples $x_1,\ldots,x_n$ are drawn from the distribution $\Pbb$ with density $f$ given by~\eqref{eqn:lb_density}. \textcolor{red}{(TODO): Clean this up.}
\begin{theorem}
	\label{thm:ppr_lb}
	Fix $\delta \in (0,1/7)$. Assume the neighborhood graph radius $r < \sigma/2$, that
	\begin{equation}
	\label{eqn:ppr_lb_condition}
	\max\biggl\{C_{3,\delta} \cdot \frac{r}{\rho} \cdot \ln\Bigl(C_{4,\delta} \frac{\rho\sigma}{r^2 \epsilon^2} \Bigr), \frac{C_{5,\delta}}{n} \cdot \frac{\rho \sigma}{r^2 \epsilon^2} \biggr\} < \frac{1}{18}~~\textrm{and}~~n \geq \frac{6 \sigma \rho (1 + \delta)}{(\sigma - 2r)(\rho - r) (1 - \delta)},
	\end{equation} 
	and that Algorithm~\ref{alg:ppr} is initialized using inputs $\alpha = 36 \cdot   
	\Phi_{n,r}(\mathcal{L}[X])$, and $(L,U) = (0,1)$.  Then the following statement holds with probability at least $1 - B_3 \exp\set{- \delta^2 b_4 n/(3 + \delta)} - n \exp\{-\delta^2 b_5 n/(3 + \delta)\}$: there exists a set $\mc{C}[X]^g$ of large volume, \\ \smash{$\vol_{n,r}(\mc{C}[X]^g \cap 
		\mc{C}^{(1)}[X]) \geq 6\vol_{n,r}(\mc{C}^{(1)}[X];G_{n,r})/49$}, such
	that for any seed node $v \in \mc{C}[X]^g$, the PPR estimated cluster
	\smash{$\wh{C}$} satisfies    
	\begin{equation}
	\label{eqn:ppr_lb}
	\frac{\sigma \rho}{r^2 n^2} \cdot \vol_{n,r}(\wh{C} \vartriangle
	\mc{C}^{(1)}[X]) \geq \frac{1 - \delta}{2} -  C_{6,\delta} \cdot 
	\frac{\sqrt{\sigma/\rho}}{\epsilon^2} \cdot \sqrt{ \log\left(C_{4,\delta} \cdot \frac{\rho \sigma}
		{\epsilon^2 r^2}\right) \frac{\sigma}{r}},   
	\end{equation} 
\end{theorem}

\textcolor{red}{(TODO): Check that the assumptions of the Theorem match those used in the proof.}

We make a few remarks:
\begin{itemize}
	\item Theorem~\ref{thm:ppr_lb} is stated with respect to a particular hard case, where the density clusters are rectangular subsets of $\Reals^2$.  We chose this
	setting to make the theorem simple to state, and our results are generalizable
	to $\Reals^d$ and to non-rectangular clusters.

	\item Although we state our lower bound with respect to PPR run on a neighborhood graph, the conclusion is likely to hold for a much broader class of spectral clustering algorithms. In the proof of Theorem~\ref{thm:ppr_lb}, we rely heavily on the fact that when $\epsilon^2$ is sufficiently greater than $\sigma/\rho$, the normalized cut of $\mc{C}^{(1)}$ will be much larger than that of $\mathcal{L}$. In this case, not merely PPR but any algorithm that approximates the minimum normalized cut is unlikely to recover $\mc{C}^{(1)}$. In particular, local spectral clustering
	algorithms based on truncated random walks \citep{spielman2013}, global spectral
	clustering algorithms \citep{shi00}, and $p$-Laplacian based spectral embeddings
	\citep{hein2010} all have provable upper bounds on the normalized cut of cluster
	they output, and thus we expect that they would all fail to estimate $\mc{C}^{(1)}$.

	\item Technically, the rectangles $\mc{C}^{(0)},\mc{C}^{(1)},\mc{C}^{(2)}$ are not
	$\sigma$-expansions due to their sharp corners. To fix this, one can   
	simply modify these sets to have appropriately rounded corners, and our lower
	bound arguments do not need to change significantly, subject to some
	additional bookkeeping.  Thus we ignore this technicality in our subsequent
	discussion. 
\end{itemize}

\subsection{Comparison between upper and lower bounds}
\label{subsec:comparison_upper_lower_bounds}

To better digest the implications of Theorem~\ref{thm:ppr_lb}, we translate the
results of our upper bound in Theorem~\ref{thm:density_cluster_volume_ssd_ub} to the density $f$ given in \eqref{eqn:lb_density}. Observe that $\mc{C}^{(1)}$ satisfies each of the Assumptions~\ref{asmp:lambda_bounded_density}--\ref{asmp:bounded_volume}:

\begin{enumerate}[label=(A\arabic*)]
	\item The density $f(x) = \frac{1 - \epsilon}{2 \rho \sigma}$ for all $x \in
	\mc{C}^{(1)}$.  
	\item The density $f(x) = \frac{\epsilon}{\rho\sigma}$ for all $x$ such
	that $0 < \dist(x,\mc{C}^{(1)}) \leq \sigma$. Therefore for all such $x$, 
	$$
	\inf_{x' \in \mc{C}^{(1)}} f(x') - f(x)  = \left\{\frac{1 - \epsilon}{2} -
	\epsilon \right\} \frac{1}{\rho \sigma},
	$$
	which meets the decay requirement with exponent $\gamma=0$.
	\item The set $\mc{C}^{(1)}$ is itself convex, and has diameter $\rho$.
	\item By symmetry, \smash{$\vol_{\Pbb,r}(\mc{C}^{(1)}) =
		\vol_{\Pbb,r}(\mc{C}^{(2)})$}, and therefore
	\smash{$\vol_{\Pbb,r}(\mc{C}^{(1)}) \leq \frac{1}{2}\vol_{\Pbb,r}(\Reals^d)$}.   
\end{enumerate}

If the user-specified parameters are initialized according to~\eqref{eqn:initialization}, we may apply Theorem~\ref{thm:density_cluster_volume_ssd_ub}. This implies that 
there exists a set $\mc{C}^{(1)}[X] \subseteq \mc{C}^{(1)}$ with
\smash{$\vol_{n,r}(\mc{C}[X]^g) \geq \frac{1}{2}\vol_{n,r}(\mc{C}[X])$} such
that for any seed node $v \in \mc{C}^{(1)}[X]$, and for large enough $n$, the
PPR estimated cluster $\wh{C}$ satisfies with high probability
\begin{equation*}
\frac{\vol_{n,r}(\wh{C} \vartriangle \mc{C}^{(1)}[X])}{\vol_{n,r}(\mc{C}^{(1)}[X])} \leq 32 C_{1,\delta} \cdot \frac{\rho^2}{\sigma r} \cdot \frac{\epsilon}{1 - \epsilon} \cdot \log^2\biggl(\sqrt{C_{2,\delta}} \frac{\rho}{2r}\biggr)
\end{equation*}
To facilitate comparisons between our upper
and lower bounds set $r = \sigma/8$.  Then
the following statements each hold with high probability. 
\begin{itemize}
	\item If the user-specified parameters satisfy~\eqref{eqn:initialization}, and
	\begin{equation*}
	\frac{\epsilon}{1 - \epsilon} \leq \frac{a}{256 C_{1,\delta}} \left(\frac{\sigma}{\rho \log(\rho/\sigma \sqrt{C_{2,\delta}})}\right)^2,
	\end{equation*}
	then \smash{$\Delta(\wh{C}, \mc{C}^{(1)}[X]) \leq a \cdot
		\vol_{n,r}(\mc{C}^{(1)}[X])$}.

	\item The population-level volume $\vol_{\Pbb,r}(\mc{C}^{(1)}) \leq (1 - \epsilon)/2 \cdot \pi r^2/(\rho\sigma)$, and
	\begin{equation*}
	\vol_{n,r}(\mc{C}^{(1)}[X]) \leq (1 + \delta) \cdot n(n - 1) \vol_{\Pbb,r}(\mc{C}^{(1)}). 
	\end{equation*}
	Therefore, if the user-specified parameters are as in Theorem~\ref{thm:ppr_lb}, and
	\begin{equation*}
	\epsilon \geq \sqrt{8 C_{6,\delta}} \left({\frac{\sigma}{\rho}} \log \left(64C_{4,\delta} \cdot \frac{\rho}
	{\epsilon^2 \sigma}\right)\right)^{1/4},
	\end{equation*}
	then \smash{$\Delta(\wh{C}, \mc{C}^{(1)}[X]) \geq \frac{1}{8}
		\vol_{n,r}(\mc{C}^{(1)}[X])$}. 
\end{itemize}

Jointly, these upper and lower bounds give a relatively precise characterization
of what it means for a density cluster to be well- or poorly-conditioned for recovery using PPR.

\begin{remark}
	It is worth pointing out that the above conclusions are reliant on specific
	(albeit reasonable) ranges and choices of input parameters, which in some
	instances differ between the upper and lower bounds. We suspect that our lower
	bound continues to hold even when choosing input parameters as dictated by our
	upper bound, but do not pursue the details.
\end{remark}

Finally, it is not hard to show that in the example under consideration, classical
plug-in density cluster estimators can consistently recover the
$\sigma$-expansion $\mc{C}_{\lambda,\sigma}$ of a density cluster $\mc{C}_{\lambda}$, even if $\epsilon$ is
large compared to $\sigma/\rho$. That PPR has trouble recovering density
clusters here (where standard plug-in approaches do not) is not meant to
be a knock on PPR. Rather, it simply reflects that while classical density
clustering approaches are specifically designed to identify high-density regions
regardless of their geometry, PPR relies on geometry as well as density when
forming the output cluster.

\section{Experiments}
\label{sec:experiments}

We provide numerical experiments to investigate the tightness of our theoretical results in Section~\ref{sec:ppr_density_cluster}, and examine the performance of PPR on the ``two moons'' dataset. We defer details of the experimental settings to the appendix.

\paragraph{Validating theoretical bounds.}
We investigate the tightness of Lemma~\ref{lem:density_cluster_local_spread} and Propositions~\ref{prop:density_cluster_normalized_cut} and \ref{prop:density_cluster_conductance} via simulation. Figure \ref{fig:bounds} compares our bounds on normalized cut, conductance, and local spread of a density cluster with the actual empirically-computed quantities. In the first row we vary the diameter $\rho$ of the candidate cluster, in the second row we vary the thickness $\sigma$ of the candidate cluster, and in the last row we vary the magnitude $\theta$ of the density drop. In each case, we see that our bounds track closely with their empirical counterparts, and are loose by roughly an order of magnitude or less. The looseness in each of these bounds will propagate to our eventual upper bound on $\Delta(\wh{C},\mc{C}_{\sigma}[X])/\vol_{n,r}(\mc{C}_{\sigma}[X])$, which as a result is loose by several orders of magnitude. Nevertheless, for sufficiently well-conditioned enough densities our upper bound will be $\ll 1$; for instance, when $\sigma = \textcolor{red}{(?)}$, $\rho = \textcolor{red}{(?)}$ and $\theta = \textcolor{red}{(?)}$.\\

\textcolor{red}{Alden for Ryan: the last couple of sentences are an attempt at what we discussed in our meeting. I am not satisfied with it, but could not do better.}

\begin{figure}[htb]
	\centering
	\includegraphics[width=0.32\textwidth]{plots/experiment_1/normalized_cut.pdf}
	\includegraphics[width=0.32\textwidth]{plots/experiment_1/conductance.pdf}
	\includegraphics[width=0.32\textwidth]{plots/experiment_1/local_spread.pdf}
	\caption{\it \small Empirical normalized cut, conductance, and local spread (in blue), versus their theoretical upper bounds (in red). In the top row we vary the diameter $\rho$, in the middle row we vary the thickness $\sigma$, and in the bottom row we vary the density drop $\theta$. \textcolor{red}{(TODO): Add middle and bottom row.}}
	\label{fig:bounds}
\end{figure}

\paragraph{Empirical behavior of PPR.} In Figure \ref{fig:moons}, to drive home 
the implications of Section~\ref{sec:ppr_density_cluster}, we show the behavior of PPR,
normalized cut, and the density clustering algorithm of \citet{chaudhuri2010} on
the well-known ``two moons'' dataset (with added 2d Gaussian noise), considered
a prototypical success story for spectral clustering algorithms. The first
column shows the empirical density clusters $\mc{C}[X]$ and $\mc{C}'[X]$ for
a particular threshold $\lambda$ of the density function; the second column
shows the cluster recovered by PPR; the third column shows the global minimum
normalized cut, computed according to the algorithm of \citet{szlam2010}; and
the last column shows a cut of the density cluster tree estimator of
\citet{chaudhuri2010}.  We can see the degrading ability of PPR to recover
density clusters as the two moons become less well-separated. Of particular
interest is the fact that PPR fails to recover one of the moons even when
normalized cut still succeeds in doing so. Additionally, we note that the
Chaudhuri-Dasgupta algorithm succeeds even when both PPR and normalized cut
fail.  This supports our main message, which is that PPR recovers only
geometrically well-conditioned density clusters.

\begin{figure}
	\centering
	\includegraphics[width=0.24\textwidth,scale = .5]{../row1_true_density_cluster}
	\includegraphics[width=0.24\textwidth]{../row1_ppr_cluster}
	\includegraphics[width=0.24\textwidth]{../row1_conductance_cluster}
	\includegraphics[width=0.24\textwidth]{../row1_density_cluster}
	\includegraphics[width=0.24\textwidth]{../row2_true_density_cluster}
	\includegraphics[width=0.24\textwidth]{../row2_ppr_cluster}
	\includegraphics[width=0.24\textwidth]{../row2_conductance_cluster}
	\includegraphics[width=0.24\textwidth]{../row2_density_cluster}
	\includegraphics[width=0.24\textwidth]{../row3_true_density_cluster}
	\includegraphics[width=0.24\textwidth]{../row3_ppr_cluster}
	\includegraphics[width=0.24\textwidth]{../row3_conductance_cluster}
	\includegraphics[width=0.24\textwidth]{../row3_density_cluster}
	\caption{\it\small True density (column 1), PPR (column 2), normalized
		cut (column 3) and estimated density (column 4) clusters for 3 different 
		simulated data sets. Seed node for PPR denoted by a black cross.} 
	\label{fig:moons}
\end{figure}

\section{Discussion}
\label{sec:discussion}
In this work, we have analyzed the behavior of PPR in the classical setup of nonparametric statistics. We have shown how PPR depends on the distribution $\Pbb$ through the population-level normalized cut, conductance, and local spread, and established upper bounds on the error with which PPR recovers an arbitrary candidate cluster $\mc{C} \subseteq \Rd$.  In the particularly important case where $\mc{C} = \mc{C}_{\lambda}$ is a $\lambda$-density cluster, we have shown that PPR recovers $\mc{C}_{\lambda}$ if and only if it is geometrically and vertically well-conditioned. We now conclude by summarizing several interesting directions for future work.

Letting the radius of the neighborhood graph shrink, $r \to 0$ as $n \to  
\infty$, would be computationally attractive, as it would ensure that the graph 
$G_{n,r}$ is sparse. However, the bounds~\eqref{eqn:volume_ssd_ub} and~\eqref{eqn:density_cluster_volume_ssd_ub} will blow up as the radius $r$ goes to $0$, preventing us from making claims about the behavior of PPR in
this regime. Although the restriction to a kernel function fixed in $n$ is
common in spectral clustering theory \citep{vonluxburg2008, schiebinger2015, singer2017},
recent works~\citep{shi2015, calder2019, garciatrillos18, garciatrillos2020, yuan2020} have demonstrated that spectral methods have meaningful continuum limits when $r \to 0$ as $n \to \infty$, and given precise rates of convergence.~\cite{garciatrillos19} have applied these results to analyze global spectral clustering in the nonparametric mixture model, obtaining asymptotic upper bounds that do not depend on $r$; it seems plausible that similar bounds could be obtained for local spectral clustering with PPR, although the arguments would necessarily be quite different.

In another direction, it would be very useful to find reasonable conditions under which the ratio $\Delta(\wh{C},\mc{C}[X])/\vol_{n,r}(\mc{C}[X])$ would tend to $0$ as $n \to \infty$. It seems likely that such a strong result would entail bounds on the $\Leb^{\infty}$-error of PPR. Although most results thus far derive bounds on the $\Leb^1$- or $\Leb^2$-error of spectral clustering methods, some recent works~\citep{dunson2020,calder2020} have established $\Leb^{\infty}$-bounds on the error with which the eigenvectors of a graph Laplacian matrix approximate the eigenvectors of a weighted Laplace-Beltrami operator. It is not clear whether the techniques used in these works can be applied to PPR.

Finally, after reading our Section~\ref{sec:ppr_density_cluster}, the practitioner may be tempted to ask why they should estimate a density cluster using PPR, when other methods are available with superior statistical guarantees. A undoubtedly correct, if perhaps unsatisfying, answer is that the ``ideal'' cluster is entirely in the eye of the beholder. For instance, it is quite common to define an ideal cluster as one with a small normalized cut, in which case the cluster output by PPR may be quite preferable to standard density clustering procedures. On the other hand, it would also be interesting to know whether a (local) spectral method such as PPR converges more rapidly (as a function of $n$) than standard density clustering procedures. Such a question could be answered by comparing the sample complexity of spectral clustering methods with that of density clustering procedures, but to our knowledge, the sample complexity of spectral methods remains unknown.

\section*{Acknowledgements}

SB is grateful to Peter Bickel, Martin Wainwright, and Larry Wasserman for
helpful and inspiring conversations. This work was supported in part by the NSF grant DMS-1713003.

\clearpage

\bibliography{../../local_spectral_bibliography}

\end{document}