### 2021-01-31 jmlr/revision/cover_letter.tex

\documentclass{article}
\usepackage{enumerate}

\begin{document}

Please find our attached revised submission titled: Statistical Guarantees for Local Spectral Clustering on Random Neighborhood Graphs. We analyze the Personalized PageRank (PPR) algorithm, a local spectral method for clustering, run over a neighborhood graph, and derive upper bounds with which PPR will recover a generic cluster. We also provide geometric conditions under which the PPR algorithm provably recovers density clusters, and show accompanying lower bounds. No previous publications by any of the authors significantly overlap with the submission. All of the co-authors are aware of the current submission and consent to its review. Thank you for your consideration of this manuscript.

\vspace{.2 in}

Sincerely,

\vspace{.1 in}
Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani

\end{document}