### 2021-09-01 jmlr/cover_letter.tex
Added some old stuff from original JMLR submission.

\documentclass{article}
\usepackage{enumerate}

\begin{document}

Dear Editors,

\vspace{.2 in}

Please find our attached submission titled: Local Spectral Clustering of Density Upper Level Sets. We analyze the Personalized PageRank (PPR) algorithm, a local spectral method for clustering, and provide geometric conditions under which the PPR algorithm provably recovers density clusters; we also show accompanying lower bounds. No previous publications by any of the authors significantly overlap with the submission. All of the co-authors are aware of the current submission and consent to its review. Sivaraman Balakrishnan has recently collaborated with Arthur Gretton; otherwise there are no potential conflicts of interest between action editors and the authors of this submission. We suggest Ingo Steinwart and Michael Mahoney as action editors particularly well-suited to handle the submitted manuscript. We would also suggest Vahab Mirrokni as a reasonable choice to serve as action editor for this submission. Thank you for your consideration of this manuscript.

\vspace{.2 in}

Sincerely,

\vspace{.1 in}
Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani

\end{document}