### 2021-12-23 notes/weekly_notes/9.20.19.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}

\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Data
\newcommand{\Xbf}{X}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Cset}{\mathcal{C}}
\newcommand{\Csig}{\Cset_{\sigma}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Vector spaces
\newcommand{\Zbb}{\mathbb{Z}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\nabla}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\spl}{\textrm{sp}}
\newcommand{\dist}{\textrm{dist}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for 9/20/19 - 9/27/19}
\author{Alden Green}
\date{\today}
\maketitle

\section{Consistency of spectral clustering in the $\epsilon$-outlier model.}

Let $X = x_1,\ldots,x_n \in \Rd$ be sampled independently from the nonparametric mixture distribution
\begin{equation*}
P = \sum_{m = 1}^{M} \pi_k \Psi_k + \epsilon \Psi_0,
\end{equation*}
where $0 < \epsilon < 1$ is the fraction of ``outliers'', and $\pi_1,\cdots,\pi_M$ are mixture weights such that $\pi_1 + \cdots + \pi_M = 1 - \epsilon$. $\Psi_1, \ldots, \Psi_k$ will each be uniform distributions over sets $\mathcal{C}_1, \ldots, \mathcal{C}_m \subset \Rd$, and $\Psi_0$ will be the outlier distribution, a uniform distribution on $[-1,1]^d$.

We wish to examine the consistency properties of kernel-based spectral clustering algorithms in the $\epsilon$-outlier model. Since there are many variants of kernel-based spectral clustering algorithms, we will begin by formalizing the specific method we analyze. To keep things simple, suppose $M = 2$. Our algorithm requires two inputs from the user: a kernel function $K(x,y)$ and a threshold $t$.

\begin{algorithm}
	\caption{Kernel-based spectral clustering.}
	\label{alg: ppr}	
	{\bfseries Input:} data $X =\{x_1,\ldots,x_n\}$, kernel function $K(x,y)$, threshold $t \in \Reals$. \\     
	{\bfseries Output:} cluster $\widehat{C} \subseteq [n]$.
	\begin{algorithmic}[1]
		\STATE Form the $n \times n$ symmetric kernel matrix $A$ with entries = $A_{ij} = (K(x_i),K(x_j))$, and the normalized (Laplacian) $L = I - D^{-1/2}AD^{-1/2}$. 
		\STATE Compute the second eigenvector of $L$, $v_2 = (v_{2,1},\ldots,v_{2,n})$. 
		\STATE Output $\widehat{C} = \set{i: v_{2,i} > t}$. 
	\end{algorithmic}
\end{algorithm}

Let $C_m = \set{i: x_i \in \mathcal{C}_m}$ be the empirical analogue to the population $\mathcal{C}_m$. We say the estimator $\widehat{C}$ is a consistent cluster estimator of the mixture model $P$ if
\begin{equation*}
\lim_{n \to \infty} \left\{\min_{m = 1,\ldots,M} \frac{1}{n} \abs{\widehat{C} \cap C_m} \right\} = 0
\end{equation*}
with probability one.

\paragraph{Summary of analysis.}

It has been shown that the spectra of the Laplacian $L$ asymptotically converges to that of a population-level integral operator, which we now define. Consider the distribution-dependent convolution operator $\mathcal{K}_P: L^2(P) \to L^2(P)$, given by
\begin{equation*}
\mathcal{K}_Pf(x) = \int K(x,y) f(y) \,dP(y),
\end{equation*}
with associated degree functional $d(x) = \int K(x,y) \,dP(y)$ for $x \in \Rd$. The population-level version of the normalized Laplacian $L$ is given by
\begin{equation}
\label{eqn:normalized_Laplacian_operator}
\overline{L}_Pf(x) = f(x) - \int \overline{K}(x,y) f(y) \,dP(y) =: f(x) - \overline{K}_Pf(x)
\end{equation}
where $\overline{K}(x,y)$ is the normalized kernel function defined as
\begin{equation*}
\overline{K}(x,y) = 
\begin{cases}
\frac{K(x,y)}{\sqrt{d(x)}\sqrt{d(y)}}, & \textrm{if $d(x), d(y) > 0$} \\
0, & \textrm{otherwise.}
\end{cases}
\end{equation*}

The operators $\overline{L}_P$ and $\overline{K}_P$ share all the same eigenfunctions, and if $\lambda_k$ is an eigenvalue of $\overline{K}_P$ then $\mu_k = 1 - \lambda_k$ is an eigenvalue of $\overline{L}_P$. We therefore analyze the properties of the leading eigenvalues of $\overline{K}_P$, which we denote $\lambda_1 \geq \lambda_2 \geq \cdots$.\footnote{In order to ensure that $\overline{K}_P$ is compact and therefore has a discrete spectrum, we assume throughout that $\iint \overline{K}^2(x,y) \,dP(x) \,dP(y) < \infty$, so that $\overline{K}_P$ is a trace-class operator.}

\subsection{Consistency of spectral clustering under the noise-free model.}

In the noise-free model, $\epsilon = 0$ and our mixture model reduces to $P = \pi_1 \Psi_1 + \pi_2 \Psi_2$ and $\pi_1 + \pi_2 = 1$, $\pi_1, \pi_2 > 0$.

\subsubsection{Spectral analysis of the $p$-weighted Laplacian in the noise-free model.}

Let $\mathcal{S}$ be the support of $P$. For $u \in H^1(\mathcal{S})$, the weighted, normalized Laplacian operator $\overline{L}_{P}$ is given by
\begin{equation*}
L_Pu = -\frac{1}{p^{3/2}} \dive(p^2 \gradient(\frac{u}{p^{1/2}}))
\end{equation*}
(where we associate functions which are equal $P$-a.e.) We say $\lambda \in \Reals$ is an eigenfunction of $L_p$ if there exists $u \in H^1(\mathcal{S}), u \neq 0$ such that
\begin{equation}
\label{eqn:weighted_Laplacian_spectra}
\int_{\mathcal{S}} \gradient u \cdot \gradient v p^2(x) \,dx = \lambda \int_{\mathcal{S}} u v p(x) \,dx, \quad \textrm{for all $v \in H^1(\mathcal{S})$.}
\end{equation}
If $u \in H^1(D)$ satisfies \eqref{eqn:weighted_Laplacian_spectra}, then we say $u$ is an eigenvector of $L_P$.

Our first result concerns the null space of $L_P$. 
\begin{lemma}
	Suppose $\mathcal{C}_1$ and $\mathcal{C}_2$ are open, bounded, and connected domains with Lipschitz boundaries. Write $p_1$ and $p_2$ for the uniform densities over $\mathcal{C}_1$ and $\mathcal{C}_2$, and let $\lambda_0$ be the smallest eigenvalue of $L_P$. Then,
	\begin{itemize}
		\item $\lambda_0 = 0$.
		\item If additionally $\dist(\mathcal{C}_1, \mathcal{C}_2) > 0$, then $u_1 \propto \sqrt{p_1}$ and $u_2 \propto \sqrt{p_2}$ are both eigenvectors of $L_P$, with eigenvalue $\lambda_0$. 
		\item If $v \in H^1(\mathcal{S})$ is an eigenvector of $L_p$ and  $v \neq u_1, v \neq u_2$, then $v$ has eigenvalue greater than $0$.
	\end{itemize}
\end{lemma}

\subsubsection{Convergence of $\overline{L}_P$.}

We will require some (very mild) conditions on the internal and external connectivity of $\Cset_1$ and $\Cset_2$ to ensure the consistency of spectral clustering under this model. We assume the clusters $\Cset_m$ are each connected sets, and have interiors of positive probability, i.e. for $\Cset_m^{\sigma} = \set{x \in \Cset_m: \dist(x,\partial \Cset_m)}$, we assume $P(\Cset_m^{\sigma}) > 0$. Further, we assume that the separation between $\mathcal{C}_1$ and $\mathcal{C}_2$ is positive, meaning $s = \dist(\mathcal{C}_1,\mathcal{C}_2) > 0$.

Suppose $K(x,y) = k(\norm{x - y})$ for a kernel function supported on $[0,r)$. The following lemma details some useful properties of the leading eigenfunctions of the associated operator $\overline{K}_P$, when $r$ is chosen to be suitably small.

\begin{lemma}
	Assume $r < s$, and further that the conductance 
	\begin{equation*}
	\Psi(\Sset) = 
	\end{equation*}
\end{lemma}

\subsection{Perturbation analysis.}

In order to demonstrate that

\subsubsection{Cheeger's inequality.}

\subsection{Asymptotic analysis.}

\section{To add.}

\begin{itemize}
	\item Explain why this distribution is interesting.
	\item Detail the various possible generalizations (non-compact kernel function,non-uniform distributions,)
	\item Flesh out the summary of analysis section.
	\item Replace assumption on normalized cut by assumption that the clusters are $\sigma$-expanded sets. 
	\item Define $H^1(\mathcal{S})$.
	\item Background on the spectra of $L_P$ (i.e. how do we know it is countable?)
\end{itemize}

\end{document}